### 2015-01-17 docs/report/title.tex
Added initial docs files

\begin{titlepage}
\begin{center}

\textsc{\LARGE University of Buenos Aires}\\[1.5cm]

\textsc{\Large Professional Assignment}\\[0.5cm]

% Title
{ \huge \bfseries Using algorithmic skeletons in EcmaScript to parallelize computation in browsers through Web Workers \\[0.4cm] }

% Author and supervisor
\noindent

\emph{Authors:}\\
Damian \textsc{Schenkelman}\\
Matias \textsc{Servetto}\\[0.5cm]

\emph{Tutor:}\\
Prof.~Rosa \textsc{Wachenhauzer}

\vfill

% Bottom of the page
{\large \today}

\end{center}
\end{titlepage}

### 2015-01-25 spikes/2.2.1/test-howToSendCode/long/setup.js
[2.2.1] adds tests.

@@ -1,0 +1,80 @@
+// http://jsperf.com/pjs-serialization-long
+
+function setupCopyWorker(){
+  var wCode = function(event){
+    var code = event.data;
+    eval('var __f = ' + code);
+    postMessage(__f());
+  };
+
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var copyWorker = setupCopyWorker();
+copyWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+// ---
+
+function setupTransferrableWorker(){
+  var wCode = function(event){
+    var codeBuffer = event.data;
+    var codeArray = new Uint8Array(codeBuffer);
+    var code = decoder.decode(codeArray);
+    eval('var __f = ' + code);
+    postMessage(__f());
+  };
+
+  var blob = new Blob(["var decoder = new TextDecoder(); onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var transferrableWorker = setupTransferrableWorker();
+transferrableWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+var encoder = new TextEncoder();
+
+// ------
+
+function setupForBlobSerializationWorker(){
+  var wCode = function(event){
+    var blobURL = event.data;
+    importScripts(blobURL);
+    postMessage(__f());
+  };
+  
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+}
+
+var blobWorker = setupForBlobSerializationWorker();
+blobWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+var fStringified = (function () {
+  var n = 1000;
+  var sum = 0;
+  for (var i = 0; i < n; i++) {
+    sum = i + (n * 2 - (i * 3));
+    sum--;
+  }
+  var counter = ((function () { 
+    var safeCounter = 0;
+    return function () { safeCounter++; return safeCounter;};
+  })());
+  while (800 > counter()) {
+    sum--;
+  }
+  return sum;
+}).toString();

### 2015-01-25 spikes/2.2.1/test-howToSendCode/short/setup.js

@@ -1,0 +1,67 @@
+// http://jsperf.com/pjs-serialization/2
+
+function setupCopyWorker(){
+  var wCode = function(event){
+    var code = event.data;
+    eval('var __f = ' + code);
+    __f();
+  };
+
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var copyWorker = setupCopyWorker();
+copyWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+// ---
+
+function setupTransferrableWorker(){
+  var wCode = function(event){
+    var codeBuffer = event.data;
+    var codeArray = new Uint8Array(codeBuffer);
+    var code = decoder.decode(codeArray);
+    eval('var __f = ' + code);
+    __f();
+  };
+
+  var blob = new Blob(["var decoder = new TextDecoder(); onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var transferrableWorker = setupTransferrableWorker();
+transferrableWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+var encoder = new TextEncoder();
+
+// ------
+
+function setupForBlobSerializationWorker(){
+  var wCode = function(event){
+    var blobURL = event.data;
+    importScripts(blobURL);
+    __f();
+  };
+  
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+}
+
+var blobWorker = setupForBlobSerializationWorker();
+blobWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+var fStringified = (function () {
+  postMessage(null);
+}).toString();

### 2015-01-25 spikes/2.2.1/test-howToSendCode/snippet01.js

@@ -1,0 +1,8 @@
+// http://jsperf.com/pjs-serialization/2
+// http://jsperf.com/pjs-serialization-long
+//sending code by blob
+
+var blob = new Blob(['__f = ' + fStringified], { type: 'application/javascript' });
+var blobURL = window.URL.createObjectURL(blob);
+
+blobWorker.postMessage(blobURL);

### 2015-01-25 spikes/2.2.1/test-howToSendCode/snippet02.js

@@ -1,0 +1,5 @@
+// http://jsperf.com/pjs-serialization/2
+// http://jsperf.com/pjs-serialization-long
+// sending code by copy
+
+copyWorker.postMessage(fStringified);

### 2015-01-25 spikes/2.2.1/test-howToSendCode/snippet03.js

@@ -1,0 +1,6 @@
+// http://jsperf.com/pjs-serialization/2
+// http://jsperf.com/pjs-serialization-long
+// sending code by transferrable objects
+
+var b = encoder.encode(fStringified);
+transferrableWorker.postMessage(b, [b.buffer]);

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/setup.js

@@ -1,0 +1,55 @@
+// http://jsperf.com/pjs-serialization-comparison/2
+
+var fStringified = (function () {
+  var n = 1000;
+  var sum = 0;
+  for (var i = 0; i < n; i++) {
+    sum = i + (n * 2 - (i * 3));
+    sum--;
+  }
+  var counter = ((function () { 
+    var safeCounter = 0;
+    return function () { safeCounter++; return safeCounter;};
+  })());
+  while (800 > counter()) {
+    sum--;
+  }
+  return sum;
+}).toString();
+
+var manualEncode = function (str) {
+  var strLength = str.length;
+  var buf = new ArrayBuffer(strLength * 2);
+  var bufView = new Uint16Array(buf);
+  for (; strLength--; ) {
+    bufView[strLength] = str.charCodeAt(strLength);
+  }
+  return bufView;
+};
+var manualDecode = function (ab) {
+  return String.fromCharCode.apply(null, ab);
+};
+
+var encoder = new TextEncoder();
+var decoder = new TextDecoder();
+
+//
+var withManual = function () {
+  var encodedF = manualEncode(fStringified);
+  var decodedF = manualDecode(encodedF);
+  if (decodedF === fStringified) {
+    console.log(decodedF);
+  } else {
+    console.log('error');
+  }
+}
+
+var withEncodeAPI = function () {
+  var encodedF = encoder.encode(fStringified);
+  var decodedF = decoder.decode(encodedF);
+  if (decodedF === fStringified) {
+    console.log(decodedF);
+  } else {
+    console.log('error');
+  }
+};

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/snippet01.js

@@ -1,0 +1,4 @@
+// http://jsperf.com/pjs-serialization-comparison/2
+// manual deco
+
+withManual();

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/snippet02.js

@@ -1,0 +1,4 @@
+// http://jsperf.com/pjs-serialization-comparison/2
+// api deco
+
+withEncodeAPI();

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/setup.js

@@ -1,0 +1,70 @@
+// http://jsperf.com/pjs-encoding
+
+function setupCopyWorker(){
+  var wCode = function(event){
+    var code = event.data.code;
+    var elements = event.data.elements;
+    if (elements.length === 10000) {
+      postMessage(elements, [elements.buffer]);
+    }
+  };
+
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var copyWorker = setupCopyWorker();
+copyWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+// ---
+
+function setupTransferrableWorker(){
+  var wCode = function(event){
+    var elements = event.data.elements;
+    var codeBuffer = event.data.code;
+    var codeArray = new Uint8Array(codeBuffer);
+    var code = decoder.decode(codeArray);
+    if (elements.length === 10000) {
+      postMessage(elements, [elements.buffer]);
+    }
+  };
+
+  var blob = new Blob(["var decoder = new TextDecoder(); onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
+};
+
+var transferrableWorker = setupTransferrableWorker();
+transferrableWorker.onmessage = function(event) {
+  deferred.resolve();
+}
+
+var encoder = new TextEncoder();
+
+// --
+
+var generateElements = function () {
+  var total = 10000;
+  var typed = new Uint8Array(total);
+  for (var i = total; i > 0; i--){
+    typed[i - 1] = i;
+  }
+  return typed;
+};
+var generateElements2 = function () {
+  var total = 10000;
+  var typed = new Uint8Array(total);
+  for (var i = total; i > 0; i--){
+    typed[i - 1] = i;
+  }
+  return typed;
+};
+
+var fStringified = (function () {
+  postMessage(null);
+}).toString();

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/snippet01.js

@@ -1,0 +1,8 @@
+// http://jsperf.com/pjs-encoding
+// sending code by copy
+
+var data = {
+  elements: generateElements(),
+  code: fStringified
+};
+copyWorker.postMessage(data, [data.elements.buffer]);

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/snippet02.js

@@ -1,0 +1,8 @@
+// http://jsperf.com/pjs-encoding
+// sending code by transferrable objects
+
+var data = {
+  elements: generateElements2(),
+  code: encoder.encode(fStringified)
+};
+transferrableWorker.postMessage(data, [data.code.buffer, data.elements.buffer]);

### 2015-01-25 docs/report/intro.tex
Added introduction and bibliography setup

\section*{Introduction}
In 1965 Gordon Moore proposed the popularly known "Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
As a consequence, programs that use a single processor/core will automatically become faster without the need to modify them with that purpose in mind.

In the year 2010 it was predicted \cite{itrs} that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation. This is the cause why newer computer have a greater number of processors instead of processors with more computing power.

To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors. In the case of application executed on a single device (e.g. browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing. For example, the .NET platform has the Task Parallel Library (TPL) and Parallel LINQ (PLINQ) \cite{ms-par} libraries that use algorithmic skeletons modelled with high order functions so developers can perform complex operations through a simple API.

On a different note, one of the tendencies that is beggining to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications that were previously only considered viable in a native environment, such as video games. This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications. Initiatives such as ParallelJS \cite{par-js} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.

The goal of this professional assignment is to provide an alternative for parallel code eection in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce. This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/setup.js
[2.2.1] adds readme. updates tests.

@@ -1,55 +1,70 @@
-// http://jsperf.com/pjs-serialization-comparison/2
+// http://jsperf.com/pjs-encoding
 
-var fStringified = (function () {
-  var n = 1000;
-  var sum = 0;
-  for (var i = 0; i < n; i++) {
-    sum = i + (n * 2 - (i * 3));
-    sum--;
-  }
-  var counter = ((function () { 
-    var safeCounter = 0;
-    return function () { safeCounter++; return safeCounter;};
-  })());
-  while (800 > counter()) {
-    sum--;
-  }
-  return sum;
-}).toString();
+function setupCopyWorker(){
+  var wCode = function(event){
+    var code = event.data.code;
+    var elements = event.data.elements;
+    if (elements.length === 10000) {
+      postMessage(elements, [elements.buffer]);
+    }
+  };
 
-var manualEncode = function (str) {
-  var strLength = str.length;
-  var buf = new ArrayBuffer(strLength * 2);
-  var bufView = new Uint16Array(buf);
-  for (; strLength--; ) {
-    bufView[strLength] = str.charCodeAt(strLength);
-  }
-  return bufView;
-};
-var manualDecode = function (ab) {
-  return String.fromCharCode.apply(null, ab);
+  var blob = new Blob(["onmessage = " + wCode.toString()]);
+  var blobURL = window.URL.createObjectURL(blob);
+
+  return new Worker(blobURL);
 };
 
-var encoder = new TextEncoder();
-var decoder = new TextDecoder();
-
-//
-var withManual = function () {
-  var encodedF = manualEncode(fStringified);
-  var decodedF = manualDecode(encodedF);
-  if (decodedF === fStringified) {
-    console.log(decodedF);
-  } else {
-    console.log('error');
-  }
+var copyWorker = setupCopyWorker();
+copyWorker.onmessage = function(event) {
+  deferred.resolve();
 }
 
-var withEncodeAPI = function () {
-  var encodedF = encoder.encode(fStringified);
-  var decodedF = decoder.decode(encodedF);
-  if (decodedF === fStringified) {
-    console.log(decodedF);
-  } else {
-    console.log('error');
-  }
+// ---
+
+function setupTransferrableWorker(){
+  var wCode = function(event){
+    var elements = event.data.elements;
+    var codeBuffer = event.data.code;
+    var codeArray = new Uint8Array(codeBuffer);
+    var code = decoder.decode(codeArray);
[…]

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/snippet01.js

@@ -1,4 +1,8 @@
-// http://jsperf.com/pjs-serialization-comparison/2
-// manual deco
+// http://jsperf.com/pjs-encoding
+// sending code by copy
 
-withManual();
+var data = {
+  elements: generateElements(),
+  code: fStringified
+};
+copyWorker.postMessage(data, [data.elements.buffer]);

### 2015-01-25 spikes/2.2.1/test-howToSendCodeWithData/snippet02.js

@@ -1,4 +1,8 @@
-// http://jsperf.com/pjs-serialization-comparison/2
-// api deco
+// http://jsperf.com/pjs-encoding
+// sending code by transferrable objects
 
-withEncodeAPI();
+var data = {
+  elements: generateElements2(),
+  code: encoder.encode(fStringified)
+};
+transferrableWorker.postMessage(data, [data.code.buffer, data.elements.buffer]);

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/setup.js

@@ -1,70 +1,55 @@
-// http://jsperf.com/pjs-encoding
-
-function setupCopyWorker(){
-  var wCode = function(event){
-    var code = event.data.code;
-    var elements = event.data.elements;
-    if (elements.length === 10000) {
-      postMessage(elements, [elements.buffer]);
-    }
-  };
-
-  var blob = new Blob(["onmessage = " + wCode.toString()]);
-  var blobURL = window.URL.createObjectURL(blob);
-
-  return new Worker(blobURL);
-};
-
-var copyWorker = setupCopyWorker();
-copyWorker.onmessage = function(event) {
-  deferred.resolve();
-}
-
-// ---
-
-function setupTransferrableWorker(){
-  var wCode = function(event){
-    var elements = event.data.elements;
-    var codeBuffer = event.data.code;
-    var codeArray = new Uint8Array(codeBuffer);
-    var code = decoder.decode(codeArray);
-    if (elements.length === 10000) {
-      postMessage(elements, [elements.buffer]);
-    }
-  };
-
-  var blob = new Blob(["var decoder = new TextDecoder(); onmessage = " + wCode.toString()]);
-  var blobURL = window.URL.createObjectURL(blob);
-
-  return new Worker(blobURL);
-};
-
-var transferrableWorker = setupTransferrableWorker();
-transferrableWorker.onmessage = function(event) {
-  deferred.resolve();
-}
-
-var encoder = new TextEncoder();
-
-// --
-
-var generateElements = function () {
-  var total = 10000;
-  var typed = new Uint8Array(total);
-  for (var i = total; i > 0; i--){
-    typed[i - 1] = i;
-  }
-  return typed;
-};
-var generateElements2 = function () {
-  var total = 10000;
-  var typed = new Uint8Array(total);
-  for (var i = total; i > 0; i--){
-    typed[i - 1] = i;
-  }
-  return typed;
-};
+// http://jsperf.com/pjs-serialization-comparison/2
 
 var fStringified = (function () {
-  postMessage(null);
+  var n = 1000;
+  var sum = 0;
+  for (var i = 0; i < n; i++) {
+    sum = i + (n * 2 - (i * 3));
+    sum--;
+  }
+  var counter = ((function () { 
+    var safeCounter = 0;
+    return function () { safeCounter++; return safeCounter;};
+  })());
+  while (800 > counter()) {
+    sum--;
+  }
+  return sum;
 }).toString();
+
[…]

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/snippet01.js

@@ -1,8 +1,4 @@
-// http://jsperf.com/pjs-encoding
-// sending code by copy
+// http://jsperf.com/pjs-serialization-comparison/2
+// manual deco
 
-var data = {
-  elements: generateElements(),
-  code: fStringified
-};
-copyWorker.postMessage(data, [data.elements.buffer]);
+withManual();

### 2015-01-25 spikes/2.2.1/test-syncEncodingAPI/snippet02.js

@@ -1,8 +1,4 @@
-// http://jsperf.com/pjs-encoding
-// sending code by transferrable objects
+// http://jsperf.com/pjs-serialization-comparison/2
+// api deco
 
-var data = {
-  elements: generateElements2(),
-  code: encoder.encode(fStringified)
-};
-transferrableWorker.postMessage(data, [data.code.buffer, data.elements.buffer]);
+withEncodeAPI();

### 2015-01-25 docs/report/micro_benchmarks.tex
Added micro benchmarks note

\section*{About Microbenchmmarks}
A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark. Throughout this document, there are multiple references to benchmarks of this kind since their results helped us make decision about how to implement some specific library features.

A micro benchmark usually involves wrapping the code under evaluation inside a loop, to:

\begin{description}
\item[First] Be able to evalute the time it takes to perform the operation multiple times (a single one could either be too small or an outlier and the measurement could have to much variance/error).
\item[Second] Allow JIT optimizations to kick in.
\end{description}

The first point is applicable to all programming languages but the second one is only important in engines/VMs that perform such optimizations. Since this document discussed the performance of JavaScript and in particular v8, the second point does affect our benchmarks.

\subsection{A small example\protect\footnote{The example is based on \cite{mraleph-bc}.}}
Let's assume that we want to profile to the following function:
\begin{lstlisting}[caption=Function to benchmark]
function between(a, b, c){
  return a <= b <= c;
}
\end{lstlisting}

A naive benchmark for it could be:
\begin{lstlisting}[caption=Naive benchmark]
// begin setup
function between(a, b, c){
  return a <= b && b <= c;
}
var a = Math.round(Math.random()*100);
var b = Math.round(Math.random()*100) + a;
var c = Math.round(Math.random()*100) + b;
//end setup
//start benchmark
console.time('between');
for (var i = 0; i < 100000; i++){
  between(a, b, c);
}
console.timeEnd('between');
//end benchmark
\end{lstlisting}

The problem with the above is that it does not consider optimizations. Let's assume that the executing engine is smart enough to:
\begin{enumerate}
  \item Inline functions
  \item Detect loop invariants
\end{enumerate}

Then due to \#1 the above could result in:
\begin{lstlisting}[caption=Naive benchmark after inlining]
// begin setup
var a = Math.round(Math.random()*100);
var b = Math.round(Math.random()*100) + a;
var c = Math.round(Math.random()*100) + b;
//end setup
//start benchmark
console.time('between');
for (var i = 0; i < 100000; i++){
  var j = a <= b && b <= c;
}
console.timeEnd('between');
//end benchmark
\end{lstlisting}

If it applies \#2 it realizes it does not need to perform the comparison in every iteration so it could extract it outside the loop, resulting in
\begin{lstlisting}[caption=Naive benchmark after evaluating loop invariants]
// begin setup
var a = Math.round(Math.random()*100);
var b = Math.round(Math.random()*100) + a;
var c = Math.round(Math.random()*100) + b;
//end setup
//start benchmark
console.time('between');
var j = a <= b && b <= c;
for (var i = 0; i < 100000; i++){
}
console.timeEnd('between');
//end benchmark
\end{lstlisting}

Alternatively it could notice that the variable \textit{j} is not being used so it could just throw the line away. In any case, we would just be measuring how long it takes for JavaScript to execute the loop, which is not what we want. To verify that is not the case the tool IRHydra can be used \cite{ir-hydra}, which \textit{can display intermediate representations used by V8}. As shown in the following figure, v8 inline the function (the blue chevron marks that) but the loop invariant was not detected and optimized:
\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{naive_benchmark_ir_hydra}
  \caption{Optimized benchmark code with inlined function.}
\end{figure}

### 2015-01-31 docs/report/serialization.tex
Initial serialization doc

\section*{Serialization and Transference}

Let's assume that we have an array of \textit{N} elements on which we want to perform a particular transformation through the \textit{map} function. If we run in a single thread (no parallelism) and the average time to process is element is \textit{t} then we could approximate the total time \(T_{ser}\) for the operation as:
\[T_{ser} = \sum_{i = 0}^{N}t = Nt\]

When trying to parallelize this operation using \textit{K} threads the ideal goal is to reach a total time \(T_{par}\) that is:
\[T_{par} = \frac{T_{ser}}{K}\]

Nevertheless, there are additional time consuming tasks other than the main computation that need to be considered when performing the operation in parallel in programs with no shared memory. These are:
\begin{itemize}
  \item Partitioning or joining the elements to transfer.
  \item Serializing/deserializing the elements to transfer.
  \item Transferring the elements between the master and the slaves.
\end{itemize}

All of the above happens twice, once when starting the computation and once when it ends, since the resulting need to be put together.

If we consider \(T_{t}\) as the transfer time, \(T_{s}\) as the serialization/deserialization time and \(T_{p}\) as the partitioning/joining time, it is clear that:
\[T_{sync} = T_{t} + T_{s} + T_{p}\]
\[T_{par} \approx \frac{T_{ser}}{K} + T_{sync}\]

Based on the previous equations we deduce that we can reduce \(T_{sync}\) the closer to the best possible scenario we will be.

In our case we are trying to transfer objects between a browser's JavaScript UI thread and Web Workers so we are constrained by the means of that environment. The Worker interface is the following \cite{w3c-ww}:
\begin{lstlisting}[language=Java, caption=The Worker interface]
[Constructor(DOMString scriptURL)]
interface Worker : EventTarget {
  void terminate();

  void postMessage(any message, optional sequence<Transferable> transfer);
  [TreatNonCallableAsNull] attribute Function? onmessage;
};
Worker implements AbstractWorker;
\end{lstlisting}

As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects. From section \textbf{2.7.4} of the HTML Standard \cite{html}:
\textit{"Some objects support being copied and closed in one operation. This is called transferring the object, and is used in particular to transfer ownership of unsharable or expensive resources across worker boundaries."}

Considering the definition of a \textit{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\). Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.5} of that same standard).

\pagebreak

### 2015-01-31 spikes/1.2/longerTransferrableVsCloning.js
[2.2.1] updates test files. [1.2] updates long trasnferable test.

@@ -2,7 +2,7 @@
 
 // JavaScript setup
 function createElements() {
-  var total = 1024*768; // 786432
+  var total = 10000000;
   var elements = new Uint32Array(total);
 
   for (var i = total; i > 0; i--){

### 2015-01-31 spikes/2.2.1/tets-functionSerialization/funcSerialization.js

@@ -1,4 +1,6 @@
-// Setup
+// Test runned at Chrome javascript console
+
+// JavaScript setup
 function setupFunctionWorker(){
   var wCode = function(event){
     var f = event.data;
@@ -19,6 +21,5 @@ functionWorker.onmessage = function(event) {
   }
 }
 
-// Begin test
+// Test case - Begin test
 functionWorker.postMessage(f);
-// End test

### 2015-01-31 docs/report/serialization_transference.tex
Fixed numbers, created macros for benchmarks

--- 2015-01-31 docs/report/serialization.tex
@@
-\section*{Serialization and Transference}
+\section{Serialization and Transference}
 
@@
 
+\subsection{Benchmarks}
+To verify the above we put together two benchmarks that aim to verify the difference between invoking \textit{postMessage} with and without structured cloning for a \textit{TypedArray}.
+Both transfer a \textit{TypedArray} back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 elements in the \textit{TypedArray} and the other one 1000000.
+
+\subsubsection{With 100000 elements}
+The following is the code for the benchmark:
+\bmcode{http://js-perf}{js_lang.tex}
+
+These were the results:
+\bmcode{Results with 100000 elements}{js_lang.tex}
+
+\subsubsection{With 1000000 elements}
+The following is the code for the benchmark:
+\bmcode{http://js-perf}{js_lang.tex}
+
+These were the results:
+\bmcode{Results with 1000000 elements}{js_lang.tex}
+
+\subsubsection{Conclusion}
+As we can see from the previous results increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \textit{Transferable} objects, it scales.
+On the other hand, the amount of operations that can be performed with structure cloning greatly decreases.
+
+For that reason we will only work with \textit{Typed Arrays} and all of them will be passed to/from Web Workers by transferring their ownership.
+
 \pagebreak

### 2015-01-31 spikes/2.1.1/newVsPreExistent.js
[2.1.1] adds graph results.

@@ -1,4 +1,4 @@
-// http://jsperf.com/worker-cold-start/4
+// http://jsperf.com/worker-cold-start/5
 
 // JavaScript setup
 function setupForPreExistentWorker(){
@@ -31,7 +31,6 @@ function postCode(worker, f){
 var worker = setupForPreExistentWorker();
 
 worker.onmessage = function(event){
-  // console.log('here');
   deferred.resolve();
 }
 

### 2015-01-31 docs/report/serialization_transference.tex
Pointing to benchmark in spikes folder

@@
 The following is the code for the benchmark:
-\bmcode{http://js-perf}{js_lang.tex}
+\bmcode{http://jsperf.com/transferrable-vs-cloning}{../../spikes/1.2/transferrableVsCloning.js}
 
 These were the results:
-\bmcode{Results with 100000 elements}{js_lang.tex}
+\bmchart{Results with 100000 elements}{../../spikes/1.2/transferrableVsCloning}
 
@@
 The following is the code for the benchmark:
-\bmcode{http://js-perf}{js_lang.tex}
+\bmcode{http://jsperf.com/longer-transferrable-vs-cloning}{../../spikes/1.2/longerTransferrableVsCloning.js}
 
 These were the results:
-\bmcode{Results with 1000000 elements}{js_lang.tex}
+\bmchart{Results with 100000 elements}{../../spikes/1.2/longerTransferrableVsCloning}
 

### 2015-01-31 spikes/2.3.1/copyComparison.js
[2.3.1] adds copy comparison benchmark and graph results.

@@ -1,0 +1,96 @@
+// http://jsperf.com/arraybuffer-copy
+
+// JavaScript setup
+function createElements() {
+  var total = 1000000;
+  var elements = new Uint8Array(total);
+
+  for (var i = total; i > 0; i--){
+    elements[i - 1] = Math.floor(Math.random() * 10000);
+  }
+  return elements;
+};
+
+var elements = createElements();
+var elementsCount = elements.length;
+
+var fileReader = new FileReader();
+
+function manualCopy (xs) {
+  var ys = new Uint8Array(xs.length);
+  for (var i = 0; i < xs.length; i++) {
+    ys[i] = xs[i];
+  }
+  return ys;
+};
+
+function bufferSliceCopy (xs) {
+  var ysBuffer = xs.buffer.slice(0);
+  return new Uint8Array(ysBuffer);
+};
+
+function constructorFromBufferCopy(xs) {
+  return new Uint8Array(xs.buffer, 0, xs.length);
+};
+
+function constructorFromArrayLikeCopy(xs) {
+  return new Uint8Array(xs);
+};
+
+function typedSetCopy (xs) {
+  var ys = new Uint8Array(xs.length);
+  ys.set(xs);
+  return ys;
+};
+
+function bufferSubarrayCopy (xs) {
+  var ys = xs.subarray();
+  return new Uint8Array(ys);
+};
+
+function blobCopy (xs, callback) {
+  var b = new Blob([xs.buffer]);
+  fileReader.onload = function() {
+      callback(new Uint8Array(this.result));
+  };
+  fileReader.readAsArrayBuffer(b);
+};
+
+// Test case 1 - Manual
+var r = manualCopy(elements);
+if (elementsCount !== r.length || r.buffer === elements.buffer) {
+  console.log('ups manual');
+}
+
+// Test case 2 - Buffer slice
+var r = bufferSliceCopy(elements);
+if (elementsCount !== r.length || r.buffer === elements.buffer) {
+  console.log('ups slice');
+}
+
+// Test case 3 - Constructor
+var r = constructorFromArrayLikeCopy(elements);
+if (elementsCount !== r.length || r.buffer === elements.buffer) {
+  console.log('ups constructor');
+}
+
+// Test case 4 - TypedArray set
+var r = typedSetCopy(elements);
+if (elementsCount !== r.length || r.buffer === elements.buffer) {
+  console.log('ups set');
+}
+
+// Test case 5 - Buffer subarray
[…]

### 2015-01-31 spikes/2.3.1/notes.js

@@ -1,0 +1,41 @@
+// 2.3.1 Particion de ArrayBuffer
+
+Manual
+  var ab = new Uint8Array([1,2,3,4]);
+  var r = new Uint8Array(ab.length);
+  for (var i = 0; i < ab.length; i++) {
+    r[i] = ab[i];
+  }
+
+ArrayBuffer slice // https://www.khronos.org/registry/typedarray/specs/latest/#5
+var ab = new Uint8Array([1,2,3,4]);
+var b = ab.buffer.slice(0);
+var r  = new Uint8Array(b);
+
+new TypeArray() from buffer
+  var ab = new Uint8Array([1,2,3,4]);
+  var r = new Uint8Array(ab.buffer, 0, ab.length);
+
+new TypeArray() from array like
+  var ab = new Uint8Array([1,2,3,4]);
+  var r = new Uint8Array(ab);
+
+ArrayBuffer set
+  var ab = new Uint8Array([1,2,3,4]);
+  var r = new Uint8Array(ab.length);
+  r.set(ab);
+
+subArray + new TypeArray(typeArray)
+  var ab = new Uint8Array([1,2,3,4]);
+  var s = ab.subarray();
+  var r = new Uint8Array(s.buffer, 0, ab.length);
+
+Blob slice
+  var ab = new Uint8Array([1,2,3,4]);
+  var b = new Blob([ab.buffer]);
+  var r;
+  var fileReader = new FileReader();
+  fileReader.onload = function() {
+      r = this.result;
+  };
+  fileReader.readAsArrayBuffer(b);

### 2015-01-31 docs/report/intro.tex
Initial serialization doc

@@
-\section*{Introduction}
+\section{Introduction}
 In 1965 Gordon Moore proposed the popularly known "Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
@@
 This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.
+
+\pagebreak

### 2015-01-31 docs/report/micro_benchmarks.tex
Initial serialization doc

@@
-\section*{About Microbenchmmarks}
+\section{About Microbenchmmarks}
 A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark.
@@
 \end{figure}
+\pagebreak

### 2015-01-31 docs/report/initialization.tex
Added initialization section to docs

\section{Initialization}

As explained in the standard \cite{w3c-ww} \textit{"Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.

There are two factors to consider from the above sentence regarding the library's implementation:

\begin{description}
\item[First] The high start-up time.
\item[Second] The high per-instance memory cost.
\end{description}

\subsection{High start-up time}
This factor raises the issue of when workers should be instantiated. On relation to this, we performed a benchmark to get an idea of the difference between cold starting a worker against having an already available one process messages.

\subsubsection{Cold start benchmark}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/worker-cold-start/5}{../../spikes/2.1.1/newVsPreExistent.js}

These were the results:
\bmchart{Evaluating the inefficiency of cold starting a Web Worker}{../../spikes/2.1.1/newVsPreExistent}

\subsubsection{Conclusion}
The previous benchmark shows a huge difference between the time required to process a message on a pre-existent Web Worker and that necessary to start Web Worker and process a message. For that reason, we decided to initialize a pool of Web Workers when the library is initialized.

\subsection{High per-instance memory cost}
The first factor makes it important that users can control/limit the amount of web workers that the library instantiates. For that reason the library allows users to specify a maximum amount of workers to be instantiated, and also provide through its API a way to terminate all workers in existance.

\subsection{Maximum amount of workers}
Since the work being done is CPU intensive (i.e.: the workers' goal is not to perform I/O operations), the maximum amount of workers to be created should not be greater than the amount of cores in the machine \(W_{cores}\), otherwise interleaving will start to affect performance. Additionally, considering the necessity of users to control how many workers are created, an upper bound can be specified \(W_{upper bound}\). If no upper bound is specified, the user upper bound for workers is assumed to be infinite (because the user is not providing a restriction). To summarize, the number of workers to be instantiated is:
\[W = min(W_{cores}, W_{upper bound})\]

\pagebreak

### 2015-01-31 spikes/1/PoC/series.js

@@ -1,4 +1,4 @@
-var total = 20000000;
+var total = 1000000;
 var typed = new Uint32Array(total);
 
 for (var i = total; i > 0; i--){

### 2015-01-31 spikes/1/PoC/test.js

@@ -1,4 +1,4 @@
-var total = 20000000;
+var total = 1000000;
 var parts = navigator.hardwareConcurrency || 4;
 var typed = new Uint32Array(total);
 
@@ -7,8 +7,8 @@ for (var i = total; i > 0; i--){
 }
 
 var wCode = function(event){
-  console.log(event);
-  var startDate = new Date();
+  // console.log(event);
+  // var startDate = new Date();
 
   var buffer = event.data;
 
@@ -22,11 +22,11 @@ var wCode = function(event){
     + Math.floor(Math.random() * 10000);
   }
 
-  var end = new Date();
+  // var end = new Date();
 
   postMessage(result.buffer, [result.buffer]);
 
-  postMessage({ start: startDate, end: end });
+  // postMessage({ start: startDate, end: end });
 }
 
 var blob = new Blob([
@@ -56,25 +56,25 @@ setTimeout(function(){
   var bytesPerElement = typed.BYTES_PER_ELEMENT;
   workers.forEach(function(w, index){
     w.onmessage = function(event){
-      if (event.data.start){
-        doneTimes++;
-        times.push(event.data);
-        if (doneTimes === parts){
-          times.forEach(function(t){
-            console.log('worker start ' + t.start.toString() + '.' + t.start.getMilliseconds());
-            console.log('worker end  ' + t.end.toString() + '.' + t.end.getMilliseconds());
-          });
-        }
-        return;
-      }
+      // if (event.data.start){
+      //   doneTimes++;
+      //   times.push(event.data);
+      //   if (doneTimes === parts){
+      //     times.forEach(function(t){
+      //       console.log('worker start ' + t.start.toString() + '.' + t.start.getMilliseconds());
+      //       console.log('worker end  ' + t.end.toString() + '.' + t.end.getMilliseconds());
+      //     });
+      //   }
+      //   return;
+      // }
       // results.push(event.data);
 
       if (++done === parts){
         console.timeEnd('worker time');
-        var endDate = new Date();
-        console.log('overall start ' + startDate.toString() + '.' + startDate.getMilliseconds());
-        console.log('overall end ' + endDate.toString() + '.' + endDate.getMilliseconds());
-        console.log(endDate - startDate);
+        // var endDate = new Date();
+        // console.log('overall start ' + startDate.toString() + '.' + startDate.getMilliseconds());
[…]

### 2015-02-08 docs/report/code_sharing.tex
Initial content on code sharing between ui and workers

\section{Code sharing}

One of the aspects that must be handled by the library is the distribution of the code between the different workers when a skeleton is created.

A naive approach would be trying to send the function to be executed directly to the worker. The following code fragment attempts to do so:
\lstinputlisting[caption=Trying to send a function to a worker, escapechar=]{../../spikes/2.2.1/test-functionSerialization/funcSerialization.js}

When executed in the Google Chrome console the error shown in Figure~\ref{fig:func-serialization} is thrown.
\img{Error when trying to send a Function object to a Web Worker}{../../spikes/2.2.1/test-functionSerialization/funcSerialization.png}{fig:func-serialization}

That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a Function does not meet that requirement as explained in section 2.7.5 of that same standard. It could be assumed that this is because of the challenges of serializing function closures \cite{cloud-haskell} \cite{distributed-scheme}.

\subsection{Serializing functions}
Considering the fact that \tfunction{} objects cannot be directly transferred a different serialization approach is to be considered. Since function closures will not be transferrable a possible alternative is to:
\begin{enumerate}
  \item Serialize the function to a \tstring{}.
  \item Transfer the \tstring{} to the worker.
  \item Create a new function on the worker from that \tstring{}.
\end{enumerate}

\note{An approach to provide contextual data to worker functions is explored in \todo{link to section}.}

\subsubsection{\tstring{} serialization benchmarks}
One possible way of passing serialized \tfunction{} to workers would be to encode them as binary and transfer them as an \tabuffer{}. To encode the strings there are two possible approaches:
\begin{itemize}
  \item Using the Encoding API\cite{encoding-api}
  \item Implementing non native encoding/decoding functions
\end{itemize}

We created a benchmark to understand the difference between each approach:
\bmcode{http://jsperf.com/pjs-serialization-comparison/2}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI.js}

These were the results:
\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI}{fig:string-encode-decode}

\subsubsection{Conclusion}
The native implementation was faster than the non-native one and the difference increased between Chrome versions, which seems to indicate that some extra optimizations could be expected in the future.

\subsection{Transferring strings}
Function's code not only needs to be serialized but also transferred. For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and tranfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.

\pagebreak

### 2015-02-08 docs/report/serialization_transference.tex
Initial content on code sharing between ui and workers

--- 2015-01-31 docs/report/serialization.tex
@@
-\section*{Serialization and Transference}
+\section{Serialization and Transference}
 
-Let's assume that we have an array of \textit{N} elements on which we want to perform a particular transformation through the \textit{map} function.
-If we run in a single thread (no parallelism) and the average time to process is element is \textit{t} then we could approximate the total time \(T_{ser}\) for the operation as:
+Let's assume that we have an array of \code{N} elements on which we want to perform a particular transformation through the \code{map} function.
+If we run in a single thread (no parallelism) and the average time to process is element is \code{t} then we could approximate the total time \(T_{ser}\) for the operation as:
 \[T_{ser} = \sum_{i = 0}^{N}t = Nt\]
 
-When trying to parallelize this operation using \textit{K} threads the ideal goal is to reach a total time \(T_{par}\) that is:
+When trying to parallelize this operation using \code{K} threads the ideal goal is to reach a total time \(T_{par}\) that is:
 \[T_{par} = \frac{T_{ser}}{K}\]
@@
   \item Serializing/deserializing the elements to transfer.
-  \item Transferring the elements between the master and the slaves.
+  \item Transferring the elements between the main thread and the workers.
 \end{itemize}
@@
 
-As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects. From section \textbf{2.7.4} of the HTML Standard \cite{html}:
+As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects.
+From section \textbf{2.7.5} of the HTML Standard \cite{html-whatwg}:
 \textit{"Some objects support being copied and closed in one operation.
@@
 
-Considering the definition of a \textit{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\).
+Considering the definition of a \code{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\).
 Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.5} of that same standard).
 
+\subsection{Benchmarks}
+To verify the above we put together two benchmarks that aim to verify the difference between invoking \code{postMessage} with and without structured cloning for a \ttarray{}.
+Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 elements in the \ttarray{} and the other one 1000000.
+
+\subsubsection{With 100000 elements}
+The following is the code for the benchmark:
+\bmcode{http://jsperf.com/transferrable-vs-cloning}{../../spikes/1.2/transferrableVsCloning.js}
+
+These were the results:
+\img{Results with 100000 elements}{../../spikes/1.2/transferrableVsCloning}{}{fig:transferrable-vs-cloning}
+
+\subsubsection{With 1000000 elements}
+The following is the code for the benchmark:
+\bmcode{http://jsperf.com/longer-transferrable-vs-cloning}{../../spikes/1.2/longerTransferrableVsCloning.js}
+
+These were the results:
+\img{Results with 100000 elements}{../../spikes/1.2/longerTransferrableVsCloning}{fig:long-transferrable-vs-cloning}
+
+\subsubsection{Conclusion}
+As we can see from the previous results increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \code{Transferable} objects, it scales.
+On the other hand, the amount of operations that can be performed with structure cloning greatly decreases.
+
+For that reason we will only work with {\ttarray{}}s and all of them will be passed to/from Web Workers by transferring their ownership.
+
 \pagebreak

### 2015-02-08 spikes/2.2.1/test-howToSendCodeWithData/encoding.js

@@ -1,6 +1,6 @@
 // http://jsperf.com/pjs-encoding
 
-// JsvaScript setup
+// JavaScript setup
 function setupCopyWorker(){
   var wCode = function(event){
     var code = event.data.code;

### 2015-02-08 spikes/2.2.1/test-syncEncodingAPI/encodingAPI.js

@@ -37,7 +37,7 @@ var manualDecode = function (ab) {
 var encoder = new TextEncoder();
 var decoder = new TextDecoder();
 
-//
+// ---
 var withManual = function () {
   var encodedF = manualEncode(fStringified);
   var decodedF = manualDecode(encodedF);
@@ -58,7 +58,7 @@ var withEncodeAPI = function () {
   }
 };
 
-// Test case - Manual deco
+// Test case 1 - Manual deco
 withManual();
 
 // Test case 2 - API deco

### 2015-02-08 docs/report/code_sharing.tex
Completed code_sharing.tex

@@
 That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a Function does not meet that requirement as explained in section 2.7.5 of that same standard.
-It could be assumed that this is because of the challenges of serializing function closures \cite{cloud-haskell} \cite{distributed-scheme}.
+It could be assumed that this is because of the challenges of serializing function closures\cite{cloud-haskell}\cite{distributed-scheme}.
 
@@
 
-These were the results:
+The results for it are displayed in Figure~\ref{fig:string-encode-decode}.
 \img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI}{fig:string-encode-decode}
@@
 
+Additionally, in most scenarios code would have to be sent along with a \ttarray{} so it would also be interesting to see if the transfer time was affected by this fact.
+
+\subsubsection{\tstring{} serialization and tranfer benchmarks}
+We created a benchmark using two functions of different length to see if the function's length affected the serialization and transfer time. We measured three approaches to transfer the string:
+\begin{enumerate}
+  \item Encoding the \tstring{} into a \tabuffer{} and sending it as a \code{Tranferable} and decoding it on the Web Worker.
+  \item Sending the string directly
+  \item Using a \code{Blob} to store the stream and retrieving that in the Worker.
+\end{enumerate}
+
+The benchmark with the short function was the following one:
+\bmcode{http://jsperf.com/pjs-serialization/2}{../../spikes/2.2.1/test-howToSendCode/shortSerialization.js}
+
+The results for it are displayed in Figure~\ref{fig:short-serialization}.
+\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCode/shortSerialization}{fig:short-serialization}
+
+The benchmark with the long function was the following one:
+\bmcode{http://jsperf.com/pjs-serialization-long}{../../spikes/2.2.1/test-howToSendCode/longSerialization.js}
+
+The results for it are displayed in Figure~\ref{fig:long-serialization}.
+\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCode/longSerialization}{fig:long-serialization}
+
+\subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}
+We created a benchmark that tranfers a \ttarray{} and \tstring{} back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.
+
+The benchmark's code is the following one:
+\bmcode{http://jsperf.com/pjs-encoding}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding.js}
+
+The results for it are displayed in Figure~\ref{fig:code-with-data}.
+\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding}{fig:code-with-data}
+
+\subsubsection{Conclusion}
+As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by directly sending it.
+
 \pagebreak

### 2015-02-08 src/typed_array_partitioner.js
[2.3.2] adds new test and implementation for TypedArray partitioner module.

@@ -16,36 +16,23 @@ Partitioner.prototype.partition = function (array) {
 };
 
 Partitioner.prototype.validateTypedArray = function (array) {
-  if (!array) {
-    throw new errors.InvalidArgumentsError(errors.messages.PARTITIONER_ARGUMENT_IS_NOT_TYPED_ARRAY); 
-  }
-  var temp = array.toString();
-  type = temp.substring('[object '.length, temp.length - 1);
-  switch(type){
-    case 'Uint8Array':
-    case 'Int8Array':
-    case 'Uint8ClampedArray':
-    case 'Uint16Array':
-    case 'Int16Array':
-    case 'Uint32Array':
-    case 'Int32Array':
-    case 'Float32Array':
-    case 'Float64Array':
-      return;
-    default:
-      throw new errors.InvalidArgumentsError('Invalid type: ' + type + '. ' + errors.messages.PARTITIONER_ARGUMENT_IS_NOT_TYPED_ARRAY); 
-  }
+  if (!utils.isTypedArray(array)) {
+    throw new errors.InvalidArgumentsError('Invalid Type: ' + array + '. ' + errors.messages.PARTITIONER_ARGUMENT_IS_NOT_TYPED_ARRAY);
+  };
 };
 
 Partitioner.prototype.doPartition = function (array) {
   var parts = this.parts;
-  var subElementsCount = array.length / parts;
+  var subElementsCount = parseInt(array.length / parts);
   var arrays = [];
   for (var i = 0; i < parts; i++) {
     var begin = i * subElementsCount;
     var end = i * subElementsCount + subElementsCount;
+    if (parts - 1 === i) {
+      end = array.length;
+    }
     var subXs = array.subarray(begin, end);
-    arrays.push(new Uint32Array(subXs));
+    arrays.push(new array.constructor(subXs));
   }
   return arrays;
 };

### 2015-02-08 src/utils.js

@@ -9,3 +9,25 @@ utils.getter = function (obj, name, value) {
     }
   });
 };
+
+utils.isTypedArray = function (array) {
+  if (!array) {
+    return false;
+  }
+  var temp = array.toString();
+  type = temp.substring('[object '.length, temp.length - 1);
+  switch(type){
+    case 'Uint8Array':
+    case 'Int8Array':
+    case 'Uint8ClampedArray':
+    case 'Uint16Array':
+    case 'Int16Array':
+    case 'Uint32Array':
+    case 'Int32Array':
+    case 'Float32Array':
+    case 'Float64Array':
+      return true;
+    default:
+      return false;
+  }
+};

### 2015-02-08 test/typed_array_partitioner.tests.js

@@ -2,92 +2,156 @@
 
 describe('array partition', function(){
   
-  var parts = 2;
-  
   var Partitioner = require('../src/typed_array_partitioner.js');
   var errors = require('../src/errors.js');
 
-  it('should be initialized with ' + parts + ' parts.', function () {
-    var partitioner = new Partitioner(parts);
-    expect(partitioner.parts).to.equal(parts);
-  });
-
-  it('should not be possible to override parts property', function () {
-    var partitioner = new Partitioner(parts);
-    expect(function () {
-      partitioner.parts = 2;
-    }).to.throw();
-  });
-
-  it('should initialize parts correctly', function () {
-    [0, 1, 2, 3, 4, 6, 8].forEach(function (p) {
-      var partitioner = new Partitioner(p);
-      expect(partitioner.parts).to.equal(p);
+  [1, 2, 3, 4, 6, 8, 16].forEach(function (parts) {
+    it('should be initialized with ' + parts + ' parts.', function () {
+      var partitioner = new Partitioner(parts);
+      expect(partitioner.parts).to.equal(parts);
     });
-  });
 
-  it('should not be posible to initialize partitioner with not numeric parts argument.', function () {
-    expect(function () {
-      var partitioner = new Partitioner();
-    }).to.throw(errors.InvalidArgumentsError);
-  });
-
-  it ('should throw exception for not TypedArray instance of Partitioner.prototype.partition\'s argument', function () {
-    var partitioner = new Partitioner(parts);
-    expect(function () {
-      partitioner.partition(null);
-    }).to.throw(errors.InvalidArgumentsError);
-    expect(function () {
-      partitioner.partition(undefined);
-    }).to.throw(errors.InvalidArgumentsError);
-    expect(function () {
-      partitioner.partition({});
-    }).to.throw(errors.InvalidArgumentsError);
-    expect(function () {
-      partitioner.partition(function () {});
-    }).to.throw(errors.InvalidArgumentsError);
-    expect(function () {
-      partitioner.partition([1,3,5]);
-    }).to.throw(errors.InvalidArgumentsError);
-  });
-
-  it ('should not throw exception for supported TypedArrays when calling Partitioner.prototype.partition', function () {
-    var partitioner = new Partitioner(parts);
-    [Uint8Array, Int8Array, Uint8ClampedArray, 
-    Uint16Array, Int16Array,
-    Uint32Array, Int32Array,
-    Float32Array, Float64Array].forEach(function (TypedArray) {
[…]

### 2015-02-08 docs/report/serialization_transference.tex

@@
 
-These were the results:
-\img{Results with 100000 elements}{../../spikes/1.2/transferrableVsCloning}{}{fig:transferrable-vs-cloning}
+The results for it are displayed in Figure~\ref{fig:transferrable-vs-cloning}.
+\img{Results with 100000 elements}{../../spikes/1.2/transferrableVsCloning}{fig:transferrable-vs-cloning}
 
@@
 
-These were the results:
+The results for it are displayed in Figure~\ref{fig:long-transferrable-vs-cloning}.
 \img{Results with 100000 elements}{../../spikes/1.2/longerTransferrableVsCloning}{fig:long-transferrable-vs-cloning}

### 2015-02-08 test/typed_array_partitioner.tests.js
[2.3.2] refactors TypedArray partitioner's tests.

@@ -11,41 +11,54 @@ describe('array partition', function(){
     }).to.throw(errors.InvalidArgumentsError);
   });
 
+  it ('should throw exception for not TypedArray instance of Partitioner.prototype.partition\'s argument', function () {
+    var partitioner = new Partitioner(4);
+    expect(function () {
+      partitioner.partition(null);
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      partitioner.partition(undefined);
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      partitioner.partition({});
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      partitioner.partition(function () {});
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      partitioner.partition([1,3,5]);
+    }).to.throw(errors.InvalidArgumentsError);
+  });
+  
+  it ('should not throw exception for supported TypedArrays when calling Partitioner.prototype.partition', function () {
+    var partitioner = new Partitioner(4);
+    [Uint8Array, Int8Array, Uint8ClampedArray, 
+    Uint16Array, Int16Array,
+    Uint32Array, Int32Array,
+    Float32Array, Float64Array].forEach(function (TypedArray) {
+      expect(function () {
+        partitioner.partition(new TypedArray([1, 2, 3, 4]));
+      }).to.not.throw();
+    });
+  });
+
+  it ('should create patitions of the same TypedArray type when calling Partitioner.prototype.partition', function () {
+    var partitioner = new Partitioner(4);
+    [Uint8Array, Int8Array, Uint8ClampedArray, 
+    Uint16Array, Int16Array,
+    Uint32Array, Int32Array,
+    Float32Array, Float64Array].forEach(function (TypedArray) {
+      var arrays = partitioner.partition(new TypedArray([1, 2, 3, 4]));
+      arrays.forEach(function (p) {
+        expect(p instanceof TypedArray).to.equal(true);
+      });
+    });
+  });
+
   [1, 2, 3, 4, 6, 8, 16].forEach(function (parts) {
     it('should be initialized with ' + parts + ' parts.', function () {
       var partitioner = new Partitioner(parts);
       expect(partitioner.parts).to.equal(parts);
-    });
-
-    it ('should throw exception for not TypedArray instance of Partitioner.prototype.partition\'s argument', function () {
-      var partitioner = new Partitioner(parts);
-      expect(function () {
[…]

### 2015-02-08 docs/report/initialization.tex
Initial content on code sharing between ui and workers
Array sharing documentation

@@
 
-As explained in the standard \cite{w3c-ww} \textit{"Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
+As explained in the standard \cite{w3c-ww} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
 
@@
 
-These were the results:
-\bmchart{Evaluating the inefficiency of cold starting a Web Worker}{../../spikes/2.1.1/newVsPreExistent}
+The results for it are displayed in Figure~\ref{fig:new-vs-pre}.
+\img{Evaluating the inefficiency of cold starting a Web Worker}{../../spikes/2.1.1/newVsPreExistent}{fig:new-vs-pre}
 

### 2015-02-08 spikes/2.3.1/copyComparison.js

@@ -55,31 +55,31 @@ function blobCopy (xs, callback) {
 // Test case 1 - Manual
 var r = manualCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {
-  console.log('ups manual');
+  console.log('error manual');
 }
 
 // Test case 2 - Buffer slice
 var r = bufferSliceCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {
-  console.log('ups slice');
+  console.log('error slice');
 }
 
 // Test case 3 - Constructor
 var r = constructorFromArrayLikeCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {
-  console.log('ups constructor');
+  console.log('error constructor');
 }
 
 // Test case 4 - TypedArray set
 var r = typedSetCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {
-  console.log('ups set');
+  console.log('error set');
 }
 
 // Test case 5 - Buffer subarray
 var r = bufferSubarrayCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {
-  console.log('ups subarray');
+  console.log('error subarray');
 }
 
 // Test case 6 - Blob
@@ -87,6 +87,6 @@ blobCopy(elements, function (result) {
   if (elementsCount === result.length) {
     deferred.resolve();
   } else {
-    console.log('ups blob');
+    console.log('error blob');
   }
 });

### 2015-02-08 spikes/2.3.1/splitComparison.js

@@ -62,23 +62,23 @@ function typedSetSplit (xs) {
 // Test case 1 - Manual
 var r = manualSplit(elements);
 if (parts !== r.length) {
-  console.log('ups manual');
+  console.log('error manual');
 }
 
 // Test case 2 - Buffer slice
 var r = bufferSliceSplit(elements);
 if (parts !== r.length) {
-  console.log('ups slice');
+  console.log('error slice');
 }
 
 // Test case 3 - Buffer subarray
 var r = bufferSubarraySplit(elements);
 if (parts !== r.length) {
-  console.log('ups subarray');
+  console.log('error subarray');
 }
 
 // Test case 4 - TypedArray set
 var r = typedSetSplit(elements);
 if (parts !== r.length) {
-  console.log('ups set');
+  console.log('error set');
 }

### 2015-02-08 spikes/2.3.1/copyComparison.js
[2.4.2] adds merging array buffers benchmark

@@ -52,6 +52,15 @@ function blobCopy (xs, callback) {
   fileReader.readAsArrayBuffer(b);
 };
 
+function dataViewCopy(xs) {
+  var buffer = new ArrayBuffer(xs.length);
+  var dv = new DataView(buffer);
+  for (var i = 0; i < xs.length; i++) {
+    dv.setInt8(xs[i]);
+  }
+  return new Uint8Array(buffer));
+}
+
 // Test case 1 - Manual
 var r = manualCopy(elements);
 if (elementsCount !== r.length || r.buffer === elements.buffer) {

### 2015-02-08 spikes/2.4.1/merge.js

@@ -1,0 +1,135 @@
+// http://jsperf.com/typedarray-merge
+
+// JavaScript setup
+var arraySize = 250000;
+var arraysCount = 4;
+
+function generateElements() {
+  var arrays = [];
+
+  for (var i = 0; i < arraysCount; i++) {
+    arrays[i] = new Uint8Array(arraySize);
+  }
+  
+  for (var j = 0; j < arraySize; j++) {
+    for (var i = 0; i < arrays.length; i++) {
+      arrays[i][j] = j + i * arraySize;
+    }
+  }
+  return arrays;
+};
+
+var parts = generateElements();
+
+function caseNonNativeFunction(parts) {
+  var totalSize = 0;
+  for (var j = 0; j < parts.length; j++) {
+    totalSize += parts[j].length;
+  }
+  var result = new Uint8Array(totalSize);
+
+  var from = 0;
+  for (var j = 0; j < parts.length; j++) {
+    var part = parts[j];
+    var partLength = part.length;
+    for (var i = 0; i < partLength; i++) {
+      result[from + i] = part[i];
+    }
+    from += partLength;
+  }
+  return result;
+};
+
+function caseTypedArraySet(parts) {
+  var totalSize = 0;
+  for (var j = 0; j < parts.length; j++) {
+    totalSize += parts[j].length;
+  }
+  var result = new Uint8Array(totalSize);
+
+  var from = 0;
+  for (var j = 0; j < parts.length; j++) {
+    var array = parts[j];
+    result.set(array, from);
+    from += array.length;
+  }
+  return result;
+};
+
+function caseDataViewSet(parts) {
+  var totalSize = 0;
+  for (var j = 0; j < parts.length; j++) {
+    totalSize += parts[j].length;
+  }
+  var buffer = new ArrayBuffer(totalSize);
+  var dv = new DataView(buffer);
+
+  var from = 0;
+  for (var j = 0; j < parts.length; j++) {
+    var part = parts[j];
+    var partLength = part.length;
+    for (var i = 0; i < partLength; i++) {
+      dv.setUint8(from + i,  part[i]);
+    }
+    from += partLength;
+  }
+  return new Uint8Array(buffer);
+};
+
+function caseArrayLikeFunction(parts) {
+  var totalSize = 0;
+  for (var j = 0; j < parts.length; j++) {
+    totalSize += parts[j].length;
+  }
+  var arrayLike = new Array(totalSize);
+  var from = 0;
+  for (var i = 0; i < parts.length; i++) {
[…]

### 2015-02-08 docs/report/array_sharing.tex

\section{Array sharing}

As explained in section~\ref{sec:serialization-and-transference} of this document transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance. Since our library will receive \ttarray{} objects to be partiotioned and handled between multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.

If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s. For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but since there's no shared memory and element ownership is tranferred to the Web Workers the elements have to be copied to the sub-arrays.

\subsection{Copying \ttarray{}s}
Sections 22.2 and 24.1 of the ECMAScript 6 draft \cite{es6} specify the API for \ttarray{} and \tabuffer{} objects respectively. Some of the exposed operations provide a way to copy the contents of an \tabuffer{} into another \tabuffer{}, so they are to be considered as viable alternatives for the copy operation. Other alternatives to consider are:
\begin{itemize}
  \item Implementing a non-native function that copies the array elements one by one
  \item Using a \code{Blob} to store the contents of a \ttarray{} and reading the \code{Blob}'s contents into a new \ttarray{}.
\end{itemize}

\subsubsection{\ttarray{} copy benchmark}
A benchmark was created based on the aforementioned alternatives:
\bmcode{http://jsperf.com/arraybuffer-copy}{../../spikes/2.3.1/copyComparison.js}

The results for it are displayed in Figure~\ref{fig:typed-array-copy}.
\img{Comparing \ttarray{} copy approaches}{../../spikes/2.3.1/copyComparison}{fig:typed-array-copy}

Based on those results we can discard the \code{Blob} approach for splitting \ttarray{}s.

\note{Alternatives \textit{Buffer subarray} and \textit{Constructor} do the same thing in the case of copying an entire \ttarray{}. However, \textit{Constructor} is not a viable alternative for copying part of an array}.

\subsubsection{\ttarray{} split benchmark}
Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array into parts:
\bmcode{http://jsperf.com/arraybuffer-copy}{../../spikes/2.3.1/splitComparison.js}

The results for it are displayed in Figure~\ref{fig:typed-array-split}.
\img{Comparing \ttarray{} split approaches}{../../spikes/2.3.1/splitComparison}{fig:typed-array-split}

\subsubsection{Conclusion}
The benchmark shows that there are three very similar alternatives to split a \ttarray{}. We are going to use \textit{Buffer subarray} as it yields better performance in a newer version of Chrome. The implementation of the slice function could look something like this:
\begin{lstlisting}[caption=Function to benchmark]
function typedArraySlice(array, from, to) {
  return new array.constructor(array.subarray(from, to));
}
\end{lstlisting}

\pagebreak

### 2015-02-08 docs/report/intro.tex

@@
 \section{Introduction}
-In 1965 Gordon Moore proposed the popularly known "Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
+In 1965 Gordon Moore proposed the popularly known ``Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
 As a consequence, programs that use a single processor/core will automatically become faster without the need to modify them with that purpose in mind.

### 2015-02-14 docs/report/array_sharing.tex
Fixes in report

@@
 If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s.
-For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but since there's no shared memory and element ownership is tranferred to the Web Workers the elements have to be copied to the sub-arrays.
+For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but since there's no shared memory and element ownership is tranferred to the Web Workers the elements must be copied to the sub-arrays.
 
@@
 \note{Alternatives \textit{Buffer subarray} and \textit{Constructor} do the same thing in the case of copying an entire \ttarray{}.
-However, \textit{Constructor} is not a viable alternative for copying part of an array}.
+However, \textit{Constructor} is not a viable alternative for copying part of an array}
 

### 2015-02-14 docs/report/initialization.tex
Fixes in report

@@
 
-As explained in the standard \cite{w3c-ww} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
+As explained in the Web Workers standard \cite{w3c-ww} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
 
@@
 The previous benchmark shows a huge difference between the time required to process a message on a pre-existent Web Worker and that necessary to start Web Worker and process a message.
-For that reason, we decided to initialize a pool of Web Workers when the library is initialized.
+For that reason, we decided to create a pool of Web Workers when the library is initialized.
 
 \subsection{High per-instance memory cost}
-The first factor makes it important that users can control/limit the amount of web workers that the library instantiates.
+The second factor makes it important for users to be able to control/limit the amount of web workers that the library instantiates.
 For that reason the library allows users to specify a maximum amount of workers to be instantiated, and also provide through its API a way to terminate all workers in existance.
@@
 \subsection{Maximum amount of workers}
-Since the work being done is CPU intensive (i.e.: the workers' goal is not to perform I/O operations), the maximum amount of workers to be created should not be greater than the amount of cores in the machine \(W_{cores}\), otherwise interleaving will start to affect performance.
+Since the work being done is CPU intensive (i.e.: the workers' goal is not to perform I/O operations), the maximum amount of workers to be created should not be greater than the amount of cores in the machine \(W_{cores}\), otherwise interleaving would start to affect performance.
 Additionally, considering the necessity of users to control how many workers are created, an upper bound can be specified \(W_{upper bound}\).

### 2015-02-14 docs/report/array_sharing.tex
Added array merge section

@@
 \subsection{Copying \ttarray{}s}
+\label{subsec:copying-typed-arrays}
 Sections 22.2 and 24.1 of the ECMAScript 6 draft \cite{es6} specify the API for \ttarray{} and \tabuffer{} objects respectively.

### 2015-02-14 spikes/2.4.1/merge.js

@@ -10,7 +10,7 @@ function generateElements() {
   for (var i = 0; i < arraysCount; i++) {
     arrays[i] = new Uint8Array(arraySize);
   }
-  
+
   for (var j = 0; j < arraySize; j++) {
     for (var i = 0; i < arrays.length; i++) {
       arrays[i][j] = j + i * arraySize;

### 2015-02-14 src/errors.js
[2.4.2] adds initial packages code and tests.

@@ -23,5 +23,8 @@ errors.messages = {
   PARTITIONER_ARGUMENT_IS_NOT_TYPED_ARRAY: 'Expected TypedArray argument.',
   ZERO_ARRAYS_TO_MERGE: 'Zero arrays to merge. Provide at least one.',
   INVALID_PARTS: 'Invalid number of parts.',
-  PART_ALREADY_COLLECTED: 'Tried to collect part {0} more than once'
+  PART_ALREADY_COLLECTED: 'Tried to collect part {0} more than once',
+  INVALID_CODE: 'Invalid code argument to package.',
+  INVALID_ELEMENTS: 'Invalid number of elements to package.',
+  INVALID_PACKAGE_INDEX: 'Package index should be not negative and less than {0}.'
 };

### 2015-02-14 src/job_packager.js

@@ -1,0 +1,47 @@
+var errors = require('./errors.js');
+var utils = require('./utils.js');
+
+var JobPackager = module.exports = function (parts, code, elements) {
+  if (!parts) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_PARTS);
+  }
+  if (!code) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
+  }
+  if (!elements) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_ELEMENTS);
+  }
+  this.parts = parts;
+  this.code = code;
+  this.elements = elements;
+};
+
+JobPackager.prototype.packageForIndex = function (index) {
+  if (0 <= index && index < this.parts) {
+    return this.packages[index];
+  }
+  var message = utils.format(errors.messages.INVALID_PACKAGE_INDEX, this.parts);
+  throw new errors.InvalidArgumentsError(message);
+};
+
+JobPackager.prototype.generatePackages = function () {
+  var parts = this.parts;
+  var elements = this.elements;
+  var packageCode = this.code.toString();
+  var elementsType = utils.typeFromTypedArray(elements);
+
+  var packages = new Array(parts);
+  for (var index = 0; index < parts; index++) {
+    packages[index] = generatePackage(index, packageCode, elements, elementsType);
+  }
+  this.packages = packages;
+};
+
+function generatePackage(index, code, elements, elementsType) {
+  return {
+    index: index,
+    code: code,
+    buffer: elements.buffer,
+    elementsType: elementsType
+  };
+};

### 2015-02-14 src/utils.js

@@ -14,8 +14,7 @@ utils.isTypedArray = function (array) {
   if (!array) {
     return false;
   }
-  var temp = array.toString();
-  type = temp.substring('[object '.length, temp.length - 1);
+  type = typedArrayType(array);
   switch(type){
     case 'Uint8Array':
     case 'Int8Array':
@@ -31,6 +30,18 @@ utils.isTypedArray = function (array) {
       return false;
   }
 };
+
+utils.typeFromTypedArray = function(array) {
+  if (!utils.isTypedArray(array)) {
+    return ;
+  }
+  return typedArrayType(array);
+}
+
+function typedArrayType (array) {
+  var temp = array.toString();
+  return temp.substring('[object '.length, temp.length - 1);
+}
 
 utils.format = function (template) {
   var toReplace = Array.prototype.slice.call(arguments, 1);

### 2015-02-14 test/job_packager.test.js

@@ -1,0 +1,109 @@
+'use strict';
+
+describe.only('job packager', function(){
+  var JobPackager = require('../src/job_packager.js');
+  var errors = require('../src/errors.js');
+  var utils = require('../src/utils.js');
+
+  var parts = 4;
+  var code = function (a) { return a + 1; };
+  var elements = new Uint32Array([1,2,3,4,5,6,7,8]);
+
+  it('should not be initialized without parts, code and elements', function () {
+    expect(function () {
+      var packager = new JobPackager();
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      var packager = new JobPackager(parts);
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      var packager = new JobPackager(parts, code);
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      var packager = new JobPackager(parts, code, elements);
+    }).to.not.throw(errors.InvalidArgumentsError);
+  });
+
+  it('should not generate packages on initialization', function () {
+    var packager = new JobPackager(parts, code, elements);
+
+    expect(packager.packages).to.equal(undefined);
+  });
+
+  it('should generate ' + parts + ' packages', function () {
+    var packager = new JobPackager(parts, code, elements);
+    packager.generatePackages();
+
+    expect(packager.packages).to.not.equal(undefined);
+    expect(packager.packages.length).to.equal(parts);
+  });
+
+  it('should return a package for each valid index', function () {
+    var packager = new JobPackager(parts, code, elements);
+    packager.generatePackages();
+
+    for (var i = 0; i < parts; i++) {
+      expect(packager.packageForIndex(i)).to.not.equal(undefined);
+    }
+  });
+
+  it('should not return a package for invalid index', function () {
+    var invalidIndex = parts;
+    var negativeIndex = -1;
+    var packager = new JobPackager(parts, code, elements);
+    packager.generatePackages();
+
+    expect(function () {
+      packager.packageForIndex(invalidIndex);
+    }).to.throw(errors.InvalidArgumentsError);
+    expect(function () {
+      packager.packageForIndex(negativeIndex);
+    }).to.throw(errors.InvalidArgumentsError);
+  });
+  
+  it('should track order on all packages', function () {
+    var packager = new JobPackager(parts, code, elements);
+    packager.generatePackages();
+
+    packager.packages.forEach(function (jobPackage, index) {
[…]

### 2015-02-14 docs/report/code_sharing.tex

@@
 When executed in the Google Chrome console the error shown in Figure~\ref{fig:func-serialization} is thrown.
-\img{Error when trying to send a Function object to a Web Worker}{../../spikes/2.2.1/test-functionSerialization/funcSerialization.png}{fig:func-serialization}
+\img{Error when trying to send a \code{Function} object to a Web Worker}{../../spikes/2.2.1/test-functionSerialization/funcSerialization.png}{fig:func-serialization}
 
-That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a Function does not meet that requirement as explained in section 2.7.5 of that same standard.
+That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a \code{Function} does not meet that requirement as explained in section 2.7.5 of that same standard.
 It could be assumed that this is because of the challenges of serializing function closures\cite{cloud-haskell}\cite{distributed-scheme}.
@@
 \subsubsection{\tstring{} serialization and tranfer benchmarks}
-We created a benchmark using two functions of different length to see if the function's length affected the serialization and transfer time. We measured three approaches to transfer the string:
+We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time.
+We measured three approaches to transfer the string:
 \begin{enumerate}
-  \item Encoding the \tstring{} into a \tabuffer{} and sending it as a \code{Tranferable} and decoding it on the Web Worker.
+  \item Encoding the \tstring{} into an \tabuffer{}, sending it as a \code{Tranferable} and decoding it on the Web Worker.
   \item Sending the string directly
@@
 \subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}
-We created a benchmark that tranfers a \ttarray{} and \tstring{} back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.
+We created a benchmark that tranfers a \ttarray{} and a \tstring{} that are properties of the same object back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.
 
@@
 \subsubsection{Conclusion}
-As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by directly sending it.
+As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by sending it directly.
 

### 2015-02-14 docs/report/micro_benchmarks.tex
Initial content on code sharing between ui and workers

@@
 A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark.
-Throughout this document, there are multiple references to benchmarks of this kind since their results helped us make decision about how to implement some specific library features.
+Throughout this document, there are multiple references to benchmarks of this kind since their results helped us make decisions about how to best implement some specific library features.
 
@@
 \begin{description}
-\item[First] Be able to evalute the time it takes to perform the operation multiple times (a single one could either be too small or an outlier and the measurement could have to much variance/error).
+\item[First] Be able to evalute the time it takes to perform the operation multiple times (a single execution could either be too small or an outlier and the measurement could have too much variance/error).
 \item[Second] Allow JIT optimizations to kick in.
@@
 
-Alternatively it could notice that the variable \textit{j} is not being used so it could just throw the line away.
-In any case, we would just be measuring how long it takes for JavaScript to execute the loop, which is not what we want.
-To verify that is not the case the tool IRHydra can be used \cite{ir-hydra}, which \textit{can display intermediate representations used by V8}.
+Alternatively it could notice that the variable \code{j} is not being used so it could just throw the line away.
+In any case, we would just be measuring how long it takes for JavaScript to execute the loop, which is not what we originally intended to do.
+To verify that is not the case the tool IRHydra\cite{ir-hydra} can be used , which \textit{can display intermediate representations used by V8}.
 As shown in the following figure, v8 inline the function (the blue chevron marks that) but the loop invariant was not detected and optimized:

### 2015-02-14 src/job_packager.js
[2.4.2] adds packages code and tests.

@@ -1,5 +1,6 @@
 var errors = require('./errors.js');
 var utils = require('./utils.js');
+var Partitioner = require('./typed_array_partitioner.js');
 
 var JobPackager = module.exports = function (parts, code, elements) {
   if (!parts) {
@@ -14,6 +15,7 @@ var JobPackager = module.exports = function (parts, code, elements) {
   this.parts = parts;
   this.code = code;
   this.elements = elements;
+  this.partitioner = new Partitioner(parts);
 };
 
 JobPackager.prototype.packageForIndex = function (index) {
@@ -29,10 +31,11 @@ JobPackager.prototype.generatePackages = function () {
   var elements = this.elements;
   var packageCode = this.code.toString();
   var elementsType = utils.typeFromTypedArray(elements);
+  var partitionedElements = this.partitioner.partition(elements);
 
   var packages = new Array(parts);
   for (var index = 0; index < parts; index++) {
-    packages[index] = generatePackage(index, packageCode, elements, elementsType);
+    packages[index] = generatePackage(index, packageCode, partitionedElements[index], elementsType);
   }
   this.packages = packages;
 };

### 2015-02-14 test/job_packager.test.js

@@ -106,4 +106,17 @@ describe.only('job packager', function(){
     });
   });
 
+  it('should generate correct buffer on al packages', function () {
+    var elementsConstructor = elements.constructor;
+    var packager = new JobPackager(parts, code, elements);
+    packager.generatePackages();
+
+    var index = 0;
+    packager.packages.forEach(function (jobPackage) {
+      var packagedElements = new elementsConstructor(jobPackage.buffer);
+      for (var i = 0; i < packagedElements.length; i++, index++) {
+        expect(packagedElements[i]).to.equal(elements[index]);
+      }
+    });
+  });
 });

### 2015-02-14 test/job_packager.test.js
[2.4.2] refactors get typed array type util

@@ -93,7 +93,7 @@ describe.only('job packager', function(){
     packager.generatePackages();
 
     packager.packages.forEach(function (jobPackage, index) {
-      expect(jobPackage.elementsType).to.equal(utils.typeFromTypedArray(elements));
+      expect(jobPackage.elementsType).to.equal(utils.getTypedArrayType(elements));
     });
   });
 
@@ -106,7 +106,7 @@ describe.only('job packager', function(){
     });
   });
 
-  it('should generate correct buffer on al packages', function () {
+  it('should not lose or duplicate data on packaged buffers', function () {
     var elementsConstructor = elements.constructor;
     var packager = new JobPackager(parts, code, elements);
     packager.generatePackages();
@@ -118,5 +118,6 @@ describe.only('job packager', function(){
         expect(packagedElements[i]).to.equal(elements[index]);
       }
     });
+    expect(index).to.equal(elements.length);
   });
 });

### 2015-02-16 examples/sapiaPicture/default-pjs.js
adds sapia sample start script

@@ -49,5 +49,5 @@ var run;
         });
     };
 
-    source.src = "mop.jpg";
+    source.src = "pic.jpg";
 })();

### 2015-02-16 examples/sapiaPicture/defaultnoworker.js

@@ -41,5 +41,5 @@ var run;
         console.timeEnd('serial-total');
     };
 
-    source.src = "mop.jpg";
+    source.src = "pic.jpg";
 })();

### 2015-02-16 dist/p-j-s-standalone.js
fixes sapia sample start script. removes dist git ignore.

@@ -1,0 +1,440 @@
+!function(e){if("object"==typeof exports&&"undefined"!=typeof module)module.exports=e();else if("function"==typeof define&&define.amd)define([],e);else{var f;"undefined"!=typeof window?f=window:"undefined"!=typeof global?f=global:"undefined"!=typeof self&&(f=self),f.pJS=e()}}(function(){var define,module,exports;return (function e(t,n,r){function s(o,u){if(!n[o]){if(!t[o]){var a=typeof require=="function"&&require;if(!u&&a)return a(o,!0);if(i)return i(o,!0);var f=new Error("Cannot find module '"+o+"'");throw f.code="MODULE_NOT_FOUND",f}var l=n[o]={exports:{}};t[o][0].call(l.exports,function(e){var n=t[o][1][e];return s(n?n:e)},l,l.exports,e,t,n,r)}return n[o].exports}var i=typeof require=="function"&&require;for(var o=0;o<r.length;o++)s(r[o]);return s})({1:[function(require,module,exports){
+var bundleFn = arguments[3];
+var sources = arguments[4];
+var cache = arguments[5];
+
+var stringify = JSON.stringify;
+
+module.exports = function (fn) {
+    var keys = [];
+    var wkey;
+    var cacheKeys = Object.keys(cache);
+    
+    for (var i = 0, l = cacheKeys.length; i < l; i++) {
+        var key = cacheKeys[i];
+        if (cache[key].exports === fn) {
+            wkey = key;
+            break;
+        }
+    }
+    
+    if (!wkey) {
+        wkey = Math.floor(Math.pow(16, 8) * Math.random()).toString(16);
+        var wcache = {};
+        for (var i = 0, l = cacheKeys.length; i < l; i++) {
+            var key = cacheKeys[i];
+            wcache[key] = key;
+        }
+        sources[wkey] = [
+            Function(['require','module','exports'], '(' + fn + ')(self)'),
+            wcache
+        ];
+    }
+    var skey = Math.floor(Math.pow(16, 8) * Math.random()).toString(16);
+    
+    var scache = {}; scache[wkey] = wkey;
+    sources[skey] = [
+        Function(['require'],'require(' + stringify(wkey) + ')(self)'),
+        scache
+    ];
+    
+    var src = '(' + bundleFn + ')({'
+        + Object.keys(sources).map(function (key) {
+            return stringify(key) + ':['
+                + sources[key][0]
[…]

### 2015-02-16 docs/report/array_merging.tex

\section{Array merging}

Once all the workers have executed the function provided to them on each element of their respective \ttarray{}s they are to notify the main thread and provide the resulting \ttarray{}.

At this point the main thread must consolidate the results into a new \ttarray{} that represents the result for the entire operation. This means copying all the elements of each partial result into a larger \ttarray{} while maintaining the original order.

\subsection{\ttarray{} merge benchmark}
Unlike the case for \ttarray{} splitting (\ref{subsec:copying-typed-arrays}) there aren't that many native alternatives to create a \ttarray{} from smaller ones.

The benchmark presented only has one alternative that uses a method with a native implementation:
\bmcode{http://jsperf.com/typedarray-merge}{../../spikes/2.4.1/merge.js}

The results for the benchmark are displayed in Figure~\ref{fig:typed-array-merge}.
\img{Comparing \ttarray{} merge approaches}{../../spikes/2.4.1/merge}{fig:typed-array-merge}

\subsection{Conclusion}
The benchmark shows that the best alternative to merge small \ttarray{}s into a larger one is using \code{TypedArray::Set} method.

For illustrative purposes, a naive implementation (without optimizations or error checking) would be similar to the following one:
\begin{lstlisting}[caption=Simple \ttarray{} merge function]
function merge(arrays){
  var first = arrays[0];
  var total = arrays.reduce(function(c,a) { return c + a.length; }, 0);
  var result = new first.constructor(total);
  var start = 0;

  arrays.forEach(function(a){
    result.set(array, start);
    start += array.length;
  });

  return result;
}
\end{lstlisting}

\pagebreak

### 2015-02-16 docs/report/intro.tex

@@
 
-The goal of this professional assignment is to provide an alternative for parallel code eection in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce.
+The goal of this professional assignment is to provide an alternative for parallel code execution in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce.
 This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

### 2015-02-16 docs/report/serialization_transference.tex
Array sharing documentation

@@
 \section{Serialization and Transference}
+\label{sec:serialization-and-transference}
 
@@
   \item Serializing/deserializing the elements to transfer.
-  \item Transferring the elements between the main thread and the workers.
+  \item Transferring the elements back and forth between the main thread and the workers.
 \end{itemize}
 
-All of the above happens twice, once when starting the computation and once when it ends, since the resulting need to be put together.
+All of the above happens twice, once when starting the computation and once when it ends, since the result from each worker needs to be consolidated before returning the new array with the transformed elements.
 
@@
 From section \textbf{2.7.5} of the HTML Standard \cite{html-whatwg}:
-\textit{"Some objects support being copied and closed in one operation.
+\textit{``Some objects support being copied and closed in one operation.
 This is called transferring the object, and is used in particular to transfer ownership of unsharable or expensive resources across worker boundaries."}
@@
 Considering the definition of a \code{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\).
-Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.5} of that same standard).
+Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.6} of that same standard).
 
@@
 To verify the above we put together two benchmarks that aim to verify the difference between invoking \code{postMessage} with and without structured cloning for a \ttarray{}.
-Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 elements in the \ttarray{} and the other one 1000000.
+Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \ttarray{} and the other one 1000000 (a million).
 
@@
 As we can see from the previous results increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \code{Transferable} objects, it scales.
-On the other hand, the amount of operations that can be performed with structure cloning greatly decreases.
+On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.
 

### 2015-02-22 examples/mapVsSerialBench/setup.js
[3.1.3] adds map vs serial benchmark and sample.

@@ -1,0 +1,76 @@
+var pjs = require('p-j-s');
+pjs.init();
+
+var generateElements = function (total) {
+  var typed = new Uint32Array(total);
+  for (var i = total; i > 0; i--){
+    typed[i - 1] = 0xdddddddd;
+  }
+  return typed;
+};
+var xsLen = 100000;
+var xs = generateElements(xsLen);
+var wrappedXs = pjs(xs);
+
+function mapper(pixel) {
+  var r = pixel & 0xFF;
+  var g = (pixel & 0xFF00) >> 8;
+  var b = (pixel & 0xFF0000) >> 16;
+
+  var noise_r = Math.random() * 0.5 + 0.5;
+  var noise_g = Math.random() * 0.5 + 0.5;
+  var noise_b = Math.random() * 0.5 + 0.5;
+
+  var new_r = Math.max(Math.min(255, noise_r * ((r * 0.393) + (g * 0.769) + (b * 0.189)) + (1 - noise_r) * r), 0);
+  var new_g = Math.max(Math.min(255, noise_g * ((r * 0.349) + (g * 0.686) + (b * 0.168)) + (1 - noise_g) * g), 0);
+  var new_b = Math.max(Math.min(255, noise_b * ((r * 0.272) + (g * 0.534) + (b * 0.131)) + (1 - noise_b) * b), 0);
+
+  return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
+};
+
+function mapCallback (result) {
+  if (0 === result.length) {
+    console.log('error');
+  }
+  deferred.resolve();
+};
+
+function serialMap(pixels, l) {
+  var result = new Uint32Array(l);
+  for (var i = 0; i < l; i += 1) {
+    var pixel = pixels[i];
+    var r = pixel & 0xFF;
+    var g = (pixel & 0xFF00) >> 8;
+    var b = (pixel & 0xFF0000) >> 16;
+
+    var noise_r = Math.random() * 0.5 + 0.5;
+    var noise_g = Math.random() * 0.5 + 0.5;
+    var noise_b = Math.random() * 0.5 + 0.5;
+
+    var new_r = Math.max(Math.min(255, noise_r * ((r * 0.393) + (g * 0.769) + (b * 0.189)) + (1 - noise_r) * r), 0);
+    var new_g = Math.max(Math.min(255, noise_g * ((r * 0.349) + (g * 0.686) + (b * 0.168)) + (1 - noise_g) * g), 0);
[…]

### 2015-02-22 spikes/3.1.3/pjsMapVsSerial/pjsMapVsSerial.js

@@ -1,0 +1,103 @@
+// http://jsperf.com/pjs-map-vs-serial
+
+// HTML setup
+<script src="https://cdn.rawgit.com/pjsteam/pjs/98fe69f822a20496ca740dc4171f0845538adac1/dist/p-j-s.min.js"></script>
+<script>
+  var pjs = require('p-j-s');
+  pjs.init();
+
+  function setupForPreExistentWorker(){
+    var wCode = function(event){
+      var codeBuffer = event.data;
+      var codeArray = new Uint8Array(codeBuffer);
+      var decoder = new TextDecoder();
+      var code = decoder.decode(codeArray);
+
+      for (var i = 0; i < l; i += 1) {
+        var r = pixel & 0xFF;
+        var g = (pixel & 0xFF00) >> 8;
+        var b = (pixel & 0xFF0000) >> 16;
+
+        var noise_r = Math.random() * 0.5 + 0.5;
+        var noise_g = Math.random() * 0.5 + 0.5;
+        var noise_b = Math.random() * 0.5 + 0.5;
+
+        var new_r = Math.max(Math.min(255, noise_r * ((r * 0.393) + (g * 0.769) + (b * 0.189)) + (1 - noise_r) * r), 0);
+        var new_g = Math.max(Math.min(255, noise_g * ((r * 0.349) + (g * 0.686) + (b * 0.168)) + (1 - noise_g) * g), 0);
+        var new_b = Math.max(Math.min(255, noise_b * ((r * 0.272) + (g * 0.534) + (b * 0.131)) + (1 - noise_b) * b), 0);
+
+        return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
+      }
+      
+    };
+
+    var blob = new Blob([
+        "onmessage = " + wCode.toString()]);
+
+    var blobURL = window.URL.createObjectURL(blob);
+
+    return new Worker(blobURL);
+  }
+
+  var generateElements = function (total) {
+    var typed = new Uint32Array(total);
+    for (var i = total; i > 0; i--){
+      typed[i - 1] = 0xdddddddd;
+    }
+    return typed;
+  };
+  var xsLen = 100000;
+  var xs = generateElements(xsLen);
+  var wrappedXs = pjs(xs);
+
+  function mapper(pixel) {
+    var r = pixel & 0xFF;
[…]

### 2015-02-22 docs/report/inlining.tex

\section{The importance of function inlining}
One of the things found while comparing the performance of the parallel and non-parallel implementations of \code{map} is that our parallel implementation performed a lot worse than what the one expected when considering the size of the \ttarray{} and the \code{mapper} function.

\subsection{The scenario}
An image tranformation to apply a sepia tone effect to pictures was used for the comparison. The core linear implementation was:
\begin{lstlisting}[caption=Sepia tone linear implementation]
function noise() {
  return Math.random() * 0.5 + 0.5;
};

function clamp(component) {
  return Math.max(Math.min(255, component), 0);
}

function colorDistance(scale, dest, src) {
  return clamp(scale * dest + (1 - scale) * src);
};

function map(binaryData, len) {
  for (var i = 0; i < len; i += 4) {
      var r = binaryData[i];
      var g = binaryData[i + 1];
      var b = binaryData[i + 2];

      binaryData[i] = colorDistance(noise(), (r * 0.393) + (g * 0.769) + (b * 0.189), r);
      binaryData[i + 1] = colorDistance(noise(), (r * 0.349) + (g * 0.686) + (b * 0.168), g);
      binaryData[i + 2] = colorDistance(noise(), (r * 0.272) + (g * 0.534) + (b * 0.131), b);
  }
};
\end{lstlisting}

\note{The serial works with a \code{Uint8ClampedArray} but accesses three elements at the same time. The parallel implementation uses a \code{Uint32Array} and works on one pixel (A,R,G,B) at a time.}

Which was implemented as shown in the following listing for the parallel case:
\begin{lstlisting}[caption=Sepia tone initial parallel implementation]
pjs(new Uint32Array(canvasData.data.buffer)).map(function(pixel){
  function noise() {
      return Math.random() * 0.5 + 0.5;
  };

  function clamp(component) {
      return Math.max(Math.min(255, component), 0);
  }

  function colorDistance(scale, dest, src) {
      return clamp(scale * dest + (1 - scale) * src);
  };

  var r = pixel & 0xFF;
  var g = (pixel & 0xFF00) >> 8;
  var b = (pixel & 0xFF0000) >> 16;

  var new_r = colorDistance(noise(), (r * 0.393) + (g * 0.769) + (b * 0.189), r);
  var new_g = colorDistance(noise(), (r * 0.349) + (g * 0.686) + (b * 0.168), g);
  var new_b = colorDistance(noise(), (r * 0.272) + (g * 0.534) + (b * 0.131), b);

  return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
}, function(result){
  // work on result
});
\end{lstlisting}

\subsection{The problem}
Taking the experience from previous experiments into account we considered that the performance of the two implementation with a \code{Uint32Array} of about \(8E7\) (eight million) elements should be similar, even favoring the parallel one. The initial measurements showed that the linear implementation fared approximately four times better than the parallel one.

\subsection{The hypothesis}
As the two code samples were similar our hypothesis was that for some reason some v8 optimizations were not taking place. It was a definite possibility considering that the parallel implementation depends on the \tfunction constructor to create \tfunction objects in the Web Workers. Specifically we believed that the internal functions \code{noise}, \code{clamp} and \code{colorDistance} were not being inlined. Inlining could be the cause of both benefits\cite{inline-pro} and drawbacks\cite{inline-cons} but the belief was that performance was being affected because of lack of it

Using IR Hydra it was easy to verify the hypothesis, as Figures \ref{fig:serial-clamp} through \ref{fig:parallel-noise} show.

\note{The blue chevron indicates that a function has been inlined.}

\subsubsection{Serial implementation generated code}
Figures \ref{fig:serial-clamp} through \ref{fig:serial-processSepia} are related to the serial implementation. As it can be seen from them, all functions are being inlined.
\img{\code{clamp} function - serial implementation}{../../spikes/3.1.3/irhydra-noworkers/inlined-clamp}{fig:serial-clamp}
\img{\code{colorDistance} function - serial implementation}{../../spikes/3.1.3/irhydra-noworkers/inlined-colorDistance}{fig:serial-colorDistance}
\img{\code{noise} function - serial implementation}{../../spikes/3.1.3/irhydra-noworkers/inlined-noise}{fig:serial-noise}
\img{\code{processSepia} function - serial implementation}{../../spikes/3.1.3/irhydra-noworkers/inlined-processSepia}{fig:serial-processSepia}

\subsubsection{Serial implementation generated code}
Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the serial implementation. As it can be seen from them, only \code{Math.random} is being inlined.
\img{\code{clamp} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-clamp}{fig:parallel-clamp}
\img{\code{colorDistance} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-colorDistance}{fig:parallel-colorDistance}
\img{\code{mapper} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-mapper}{fig:parallel-mapper}
\img{\code{noise} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-noise}{fig:parallel-noise}

\subsection{The solution}
The way to solve the problem was to manually inline the functions. The resulting parallel implementation was:
\begin{lstlisting}[caption=Sepia tone final parallel implementation]
pjs(buff).map(function(pixel){
  var r = pixel & 0xFF;
  var g = (pixel & 0xFF00) >> 8;
  var b = (pixel & 0xFF0000) >> 16;
  var noiser = Math.random() * 0.5 + 0.5;
  var noiseg = Math.random() * 0.5 + 0.5;
  var noiseb = Math.random() * 0.5 + 0.5;

  var new_r = Math.max(Math.min(255, noiser * ((r * 0.393) + (g * 0.769) + (b * 0.189)) + (1 - noiser) * r), 0);
  var new_g = Math.max(Math.min(255, noiseg * ((r * 0.349) + (g * 0.686) + (b * 0.168)) + (1 - noiseg) * g), 0);
  var new_b = Math.max(Math.min(255, noiseb * ((r * 0.272) + (g * 0.534) + (b * 0.131)) + (1 - noiseb) * b), 0);

  return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
}, function(result){
  // work on result
});
\end{lstlisting}

We created a benchmark to understand the difference between the initial version (the one in which functions were not inlined) and the final one (with inlining in charge of the developer):
\bmcode{http://jsperf.com/pjs-map-inlining}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}

The results for it are displayed in Figure~\ref{fig:pjs-map-inlining}.
\img{Comparing string encoding/decoding approaches}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining}{fig:pjs-map-inlining}

\subsection{Conclusion}
This section shows that in those cases in which the compiler inlines some operations that are executed as part of the skeleton, it is important for developers to explicitly inline those when using the library.

\pagebreak

### 2015-03-01 spikes/2.3.1/copyComparison.js
[2.3.1] updates 2.3.1 benchmarks

@@ -1,65 +1,72 @@
-// http://jsperf.com/arraybuffer-copy
+// http://jsperf.com/arraybuffer-copy/3
 
-// JavaScript setup
-function createElements() {
-  var total = 1000000;
-  var elements = new Uint8Array(total);
-
-  for (var i = total; i > 0; i--){
-    elements[i - 1] = Math.floor(Math.random() * 10000);
-  }
-  return elements;
-};
-
-var elements = createElements();
-var elementsCount = elements.length;
-
-var fileReader = new FileReader();
-
-function manualCopy (xs) {
-  var ys = new Uint8Array(xs.length);
-  for (var i = 0; i < xs.length; i++) {
-    ys[i] = xs[i];
-  }
-  return ys;
-};
-
-function bufferSliceCopy (xs) {
-  var ysBuffer = xs.buffer.slice(0);
-  return new Uint8Array(ysBuffer);
-};
-
-function constructorFromArrayLikeCopy(xs) {
-  return new Uint8Array(xs);
-};
-
-function typedSetCopy (xs) {
-  var ys = new Uint8Array(xs.length);
-  ys.set(xs);
-  return ys;
-};
-
-function bufferSubarrayCopy (xs) {
-  var ys = xs.subarray();
-  return new Uint8Array(ys);
-};
-
-function blobCopy (xs, callback) {
-  var b = new Blob([xs.buffer]);
-  fileReader.onload = function() {
-      callback(new Uint8Array(this.result));
+// HTML setup
+<script src="http://mrale.ph/irhydra/jsperf-renamer.js"></script>
+<script>
+  function createElements() {
+    var total = 1000000;
+    var elements = new Uint8Array(total);
+  
+    for (var i = total; i > 0; i--){
+      elements[i - 1] = Math.floor(Math.random() * 10000);
+    }
+    return elements;
   };
-  fileReader.readAsArrayBuffer(b);
-};
-
-function dataViewCopy(xs) {
-  var buffer = new ArrayBuffer(xs.length);
-  var dv = new DataView(buffer);
-  for (var i = 0; i < xs.length; i++) {
-    dv.setInt8(xs[i]);
-  }
-  return new Uint8Array(buffer));
-}
+  
+  var elements = createElements();
+  var elementsCount = elements.length;
+  
+  var fileReader = new FileReader();
+  
+  function manualCopy (xs) {
+    var ys = new Uint8Array(xs.length);
[…]

### 2015-03-01 spikes/2.3.1/splitComparison.js

@@ -1,63 +1,66 @@
-// http://jsperf.com/arraybuffer-split
+// http://jsperf.com/arraybuffer-split/3
 
-// JavaScript setup
-function createElements() {
-  var total = 1000000;
-  var elements = new Uint8Array(total);
+// HTML setup
+<script src="http://mrale.ph/irhydra/jsperf-renamer.js"></script>
+<script>
+  function createElements() {
+    var total = 1000000;
+    var elements = new Uint8Array(total);
 
-  for (var i = total; i > 0; i--){
-    elements[i - 1] = Math.floor(Math.random() * 10000);
-  }
-  return elements;
-};
-
-var parts = 4;
-var elements = createElements();
-var elementsCount = elements.length;
-var subElementsCount = elementsCount / parts;
-
-function manualSplit (xs) {
-  var ys = [];
-  for (var i = 0; i < parts; i++) {
-    ys.push(new Uint8Array(subElementsCount));
-  }
-
-  for (var i = 0; i < subElementsCount; i++) {
-    for (var p = 0; p < parts; p++) {
-      ys[p][i] = xs[i + p * subElementsCount];
+    for (var i = total; i > 0; i--){
+      elements[i - 1] = Math.floor(Math.random() * 10000);
     }
-  }
-  return ys;
-};
+    return elements;
+  };
 
-function bufferSliceSplit (xs) {
-  var ys = [];
-  for (var i = 0; i < parts; i++) {
-    var b = xs.buffer.slice(i * subElementsCount * xs.BYTES_PER_ELEMENT, (i * subElementsCount + subElementsCount) * xs.BYTES_PER_ELEMENT);
-    ys.push(new Uint8Array(b));
-  }
-  return ys;
-};
+  var parts = 4;
+  var elements = createElements();
+  var elementsCount = elements.length;
+  var subElementsCount = elementsCount / parts;
 
-function bufferSubarraySplit (xs) {
-  var ys = [];
-  for (var i = 0; i < parts; i++) {
-    var subXs = xs.subarray(i * subElementsCount, i * subElementsCount + subElementsCount);
-    ys.push(new Uint8Array(subXs));
-  }
-  return ys;
-};
+  function manualSplit (xs) {
+    var ys = [];
+    for (var i = 0; i < parts; i++) {
+      ys.push(new Uint8Array(subElementsCount));
+    }
 
-function typedSetSplit (xs) {
-  var ys = [];
[…]

### 2015-03-01 docs/report/micro_benchmarks.tex
Notes about benchmarks

@@
 \end{figure}
+
+\note{All benchmarks are performed using two different versions of Google Chrome (40.0.2214 and 41.0.2272) to be able to get an idea of the evolution of each feature.}
+
+\note{Each benchmark is executed 5 (five) times to reduce variance.}
+
+\note{No browser plug-ins are enabled when running the benchmarks as those could introduce additional sources of error.}
 \pagebreak

### 2015-03-01 spikes/2.4.1/merge.js
[2.4.1] updates 2.4.1 benchmarks

@@ -1,126 +1,128 @@
-// http://jsperf.com/typedarray-merge
+// http://jsperf.com/typedarray-merge/2
 
 // JavaScript setup
-var arraySize = 250000;
-var arraysCount = 4;
+<script>
+  var arraySize = 250000;
+  var arraysCount = 4;
 
-function generateElements() {
-  var arrays = [];
+  function generateElements() {
+    var arrays = [];
 
-  for (var i = 0; i < arraysCount; i++) {
-    arrays[i] = new Uint8Array(arraySize);
+    for (var i = 0; i < arraysCount; i++) {
+      arrays[i] = new Uint8Array(arraySize);
+    }
+    
+    for (var j = 0; j < arraySize; j++) {
+      for (var i = 0; i < arrays.length; i++) {
+        arrays[i][j] = j + i * arraySize;
+      }
+    }
+    return arrays;
+  };
+
+  var parts = generateElements();
+
+  function caseNonNativeFunction(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var result = new Uint8Array(totalSize);
+
+    var from = 0;
+    for (var j = 0; j < parts.length; j++) {
+      var part = parts[j];
+      var partLength = part.length;
+      for (var i = 0; i < partLength; i++) {
+        result[from + i] = part[i];
+      }
+      from += partLength;
+    }
+    return result;
+  };
+
+  function caseTypedArraySet(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var result = new Uint8Array(totalSize);
+
+    var from = 0;
+    for (var j = 0; j < parts.length; j++) {
+      var array = parts[j];
+      result.set(array, from);
+      from += array.length;
+    }
+    return result;
+  };
+
+  function caseDataViewSet(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var buffer = new ArrayBuffer(totalSize);
+    var dv = new DataView(buffer);
+
+    var from = 0;
+    for (var j = 0; j < parts.length; j++) {
+      var part = parts[j];
+      var partLength = part.length;
+      for (var i = 0; i < partLength; i++) {
+        dv.setUint8(from + i,  part[i]);
+      }
+      from += partLength;
+    }
[…]

### 2015-03-01 docs/report/function_caching.tex
Added function caching benchmark

\section{Function caching}

Before comparing the sequential implementation of \code{map} against the parallel one we believed that the cost of dynamically instantiating a \tfunction{} from a \tstring{} in the Web Workers could be avoided by introducing a cache for cases where the worker has already seen the code for a particular function.

We decided to introduce a simple key/value map acting as a cache where the keys are the \tstring{}s with the function's code and the values are the \tfunction{} objects.

To verify its performance we created a benchmark that runs with and without the function cache:
\bmcode{http://jsperf.com/p-j-s-with-vs-without-function-cache}{../../spikes/3.1.3/functionCache/functionCache.js}

The results for the benchmark are displayed in Figure~\ref{fig:function-cache-result}.
\img{Comparing p-j-s \code{map} with and without a function cache}{../../spikes/3.1.3/functionCache/functionCache}{fig:function-cache-result}

\subsection{Conclusion}
As the benchmark shows versions 40 and 41 of chrome have almost no difference between the implementations with and without the cache. To determine whether introducing the cache was worth it or not we also used Chrome Canary version 42. The latter shows a big increase in performance (near 20\%), which is why we decided to leave the \tfunction{} cache as part of our implementation.

\pagebreak

### 2015-03-08 dist/p-j-s-standalone.js
fixes #10 refactors JobPackager

@@ -105,13 +105,13 @@ var initialized = false;
 var workers = [];
 
 var WrappedTypedArray = function(source, parts){
+	this.packager = new JobPackager(this.parts, source);
 	this.source = source;
 	this.parts = parts;
 };
 
 WrappedTypedArray.prototype.__operationSkeleton = function (f, operation, collectorMapper, done) {
-	var packager = new JobPackager(this.parts, f, this.source, operation);
-	var packs = packager.generatePackages();
+	var packs = packager.generatePackages(f, operation);
 
 	var collector = new ResultCollector(this.parts, function(results){
 		var partial_results = results.map(collectorMapper);
@@ -204,34 +204,31 @@ var operations = Object.keys(operation_names).map(function (k) {
   return operation_names[k];
 });
 
-var JobPackager = module.exports = function (parts, code, elements, operation) {
+var JobPackager = module.exports = function (parts, elements) {
   if (!parts) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_PARTS);
-  }
-  if (!code) {
-    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
   }
   if (!elements) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_ELEMENTS);
   }
+  this.parts = parts;
+  this.elements = elements;
+};
+
+JobPackager.prototype.generatePackages = function (code, operation) {
+  if (!code) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
+  }
   if (!operation || -1 === operations.indexOf(operation)) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_OPERATION);
   }
-  this.parts = parts;
-  this.code = code;
-  this.elements = elements;
-  this.operation = operation;
-};
-
-JobPackager.prototype.generatePackages = function () {
-  var functionString = this.code.toString();
+  var functionString = code.toString();
   var match = functionString.match(FUNCTION_REGEX);
   var packageCodeArg = match[1].split(',')[0].trim();
   var packageCode = match[2];
   var elementsType = utils.getTypedArrayType(this.elements);
   var partitioner = new Partitioner(this.parts);
   var partitionedElements = partitioner.partition(this.elements);
-  var operation = this.operation;
 
   return partitionedElements.map(function (partitionedElement, index) {
     return {

### 2015-03-08 dist/p-j-s.js

@@ -103,34 +103,31 @@ var operations = Object.keys(operation_names).map(function (k) {
   return operation_names[k];
 });
 
-var JobPackager = module.exports = function (parts, code, elements, operation) {
+var JobPackager = module.exports = function (parts, elements) {
   if (!parts) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_PARTS);
-  }
-  if (!code) {
-    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
   }
   if (!elements) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_ELEMENTS);
   }
+  this.parts = parts;
+  this.elements = elements;
+};
+
+JobPackager.prototype.generatePackages = function (code, operation) {
+  if (!code) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
+  }
   if (!operation || -1 === operations.indexOf(operation)) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_OPERATION);
   }
-  this.parts = parts;
-  this.code = code;
-  this.elements = elements;
-  this.operation = operation;
-};
-
-JobPackager.prototype.generatePackages = function () {
-  var functionString = this.code.toString();
+  var functionString = code.toString();
   var match = functionString.match(FUNCTION_REGEX);
   var packageCodeArg = match[1].split(',')[0].trim();
   var packageCode = match[2];
   var elementsType = utils.getTypedArrayType(this.elements);
   var partitioner = new Partitioner(this.parts);
   var partitionedElements = partitioner.partition(this.elements);
-  var operation = this.operation;
 
   return partitionedElements.map(function (partitionedElement, index) {
     return {
@@ -427,13 +424,13 @@ var initialized = false;
 var workers = [];
 
 var WrappedTypedArray = function(source, parts){
+	this.packager = new JobPackager(this.parts, source);
 	this.source = source;
 	this.parts = parts;
 };
 
 WrappedTypedArray.prototype.__operationSkeleton = function (f, operation, collectorMapper, done) {
-	var packager = new JobPackager(this.parts, f, this.source, operation);
-	var packs = packager.generatePackages();
+	var packs = packager.generatePackages(f, operation);
 
 	var collector = new ResultCollector(this.parts, function(results){
 		var partial_results = results.map(collectorMapper);

### 2015-03-08 src/index.js

@@ -11,14 +11,13 @@ var initialized = false;
 var workers = [];
 
 var WrappedTypedArray = function(source, parts){
+	this.packager = new JobPackager(parts, source);
 	this.source = source;
 	this.parts = parts;
 };
 
 WrappedTypedArray.prototype.__operationSkeleton = function (f, operation, collectorMapper, done) {
-	var packager = new JobPackager(this.parts, f, this.source, operation);
-	var packs = packager.generatePackages();
-
+	var packs = this.packager.generatePackages(f, operation);
 	var collector = new ResultCollector(this.parts, function(results){
 		var partial_results = results.map(collectorMapper);
 		var m = merge_typed_arrays(partial_results);

### 2015-03-08 src/job_packager.js

@@ -9,34 +9,31 @@ var operations = Object.keys(operation_names).map(function (k) {
   return operation_names[k];
 });
 
-var JobPackager = module.exports = function (parts, code, elements, operation) {
+var JobPackager = module.exports = function (parts, elements) {
   if (!parts) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_PARTS);
-  }
-  if (!code) {
-    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
   }
   if (!elements) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_ELEMENTS);
   }
+  this.parts = parts;
+  this.elements = elements;
+};
+
+JobPackager.prototype.generatePackages = function (code, operation) {
+  if (!code) {
+    throw new errors.InvalidArgumentsError(errors.messages.INVALID_CODE);
+  }
   if (!operation || -1 === operations.indexOf(operation)) {
     throw new errors.InvalidArgumentsError(errors.messages.INVALID_OPERATION);
   }
-  this.parts = parts;
-  this.code = code;
-  this.elements = elements;
-  this.operation = operation;
-};
-
-JobPackager.prototype.generatePackages = function () {
-  var functionString = this.code.toString();
+  var functionString = code.toString();
   var match = functionString.match(FUNCTION_REGEX);
   var packageCodeArg = match[1].split(',')[0].trim();
   var packageCode = match[2];
   var elementsType = utils.getTypedArrayType(this.elements);
   var partitioner = new Partitioner(this.parts);
   var partitionedElements = partitioner.partition(this.elements);
-  var operation = this.operation;
 
   return partitionedElements.map(function (partitionedElement, index) {
     return {

### 2015-03-08 test/index.tests.js

@@ -91,10 +91,11 @@ describe('initialization', function(){
   });
 });
 
-describe ('wrap tests', function(){
+describe.only('wrap tests', function(){
 
   var pjs;
   var errors = require('../src/errors.js');
+  var utils = require('../src/utils.js');
 
   beforeEach(function () {
     pjs = require('../src/index.js');
@@ -118,4 +119,19 @@ describe ('wrap tests', function(){
       pjs([1,6,8,9,14]);
     }).to.throw(errors.InvalidArgumentsError);
   });
+
+  [Uint8Array, Int8Array, Uint8ClampedArray,
+    Uint16Array, Int16Array,
+    Uint32Array, Int32Array,
+    Float32Array, Float64Array].forEach(function (TypedArray) {
+      describe(utils.format('tests for {0}', utils.getTypedArrayConstructorType(TypedArray)), function(){
+        var sourceArray = new TypedArray([1,2,3,4,5]);
+
+        it('should not throw if object to wrap is typed array', function () {
+          expect(function () {
+            pjs(sourceArray);
+          }).to.not.throw(errors.InvalidArgumentsError);
+        });
+      });
+    });
 });

### 2015-03-08 test/job_packager.test.js

@@ -6,13 +6,14 @@ describe('job packager', function(){
   var utils = require('../src/utils');
 
   var parts = 4;
+
   var code = function (a) { return a + 1; };
   var elements = new Uint32Array([1,2,3,4,5,6,7,8]);
   var operationFilter = 'filter';
   var operationMap = 'map';
-  var packager = new JobPackager(parts, code, elements, operationFilter);
+  var packager = new JobPackager(parts, elements);
   var invalidOperation = 'add';
-  var packages = packager.generatePackages();
+  var packages = packager.generatePackages(code, operationFilter);
 
   it('should not support empty initialization', function () {
     expect(function () {
@@ -20,39 +21,37 @@ describe('job packager', function(){
     }).to.throw(errors.InvalidArgumentsError);
   });
 
-  it('should not support empty code initialization', function () {
+  it('should not support empty elements initialization', function () {
     expect(function () {
       var p = new JobPackager(parts);
     }).to.throw(errors.InvalidArgumentsError);
   });
 
-  it('should not support empty elements initialization', function () {
+  it('should not support empty operation for package generation', function () {
     expect(function () {
-      var p = new JobPackager(parts, code);
+      var p = new JobPackager(parts, elements);
+      var ps = p.generatePackages(); 
     }).to.throw(errors.InvalidArgumentsError);
   });
 
-  it('should not support empty operation initialization', function () {
+  it('should not support invalid operation for package generation', function () {
     expect(function () {
-      var p = new JobPackager(parts, code, elements);
+      var p = new JobPackager(parts, elements);
+      var ps = p.generatePackages(code, invalidOperation);
     }).to.throw(errors.InvalidArgumentsError);
   });
 
-  it('should not support invalid operation initialization', function () {
+  it('should support valid map operation for package generation', function () {
     expect(function () {
-      var p = new JobPackager(parts, code, elements, invalidOperation);
-    }).to.throw(errors.InvalidArgumentsError);
-  });
-
-  it('should support valid map operation initialization', function () {
-    expect(function () {
-      var p = new JobPackager(parts, code, elements, operationMap);
+      var p = new JobPackager(parts, elements);
+      var ps = p.generatePackages(code, operationMap);
     }).to.not.throw(errors.InvalidArgumentsError);
   });
 
-  it('should support valid filter operation initialization', function () {
+  it('should support valid filter operation for package generation', function () {
[…]

### 2015-03-08 test/worker_core.test.js

@@ -14,8 +14,8 @@ describe('worker core', function(){
       var code = function (a) { return a + 1; };
       var packageIndex = 3, result;
       var elements = new TypedArray([1,2,3,4,5,6,7,8]);
-      var packager = new JobPackager(parts, code, elements, 'map');
-      var packages = packager.generatePackages();
+      var packager = new JobPackager(parts, elements);
+      var packages = packager.generatePackages(code, 'map');
 
       var expectedElements = new TypedArray([8,9]);
 

### 2015-03-08 docs/report/inlining.tex

@@
 
-\note{The serial works with a \code{Uint8ClampedArray} but accesses three elements at the same time. The parallel implementation uses a \code{Uint32Array} and works on one pixel (A,R,G,B) at a time.}
+\note{The serial code works with a \code{Uint8ClampedArray} but accesses three elements at the same time.
+The parallel implementation uses a \code{Uint32Array} and works on one pixel (A,R,G,B) at a time.}
 
@@
 \subsection{The problem}
-Taking the experience from previous experiments into account we considered that the performance of the two implementation with a \code{Uint32Array} of about \(8E7\) (eight million) elements should be similar, even favoring the parallel one.
+Taking the experience from previous experiments into account we considered that the performance of the two implementations with a source \code{Uint32Array} of about \(8E7\) (eight million) elements should be similar, even favoring the parallel one.
 The initial measurements showed that the linear implementation fared approximately four times better than the parallel one.
@@
 
-\subsubsection{Serial implementation generated code}
+\subsubsection{Parallel implementation generated code}
 Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the serial implementation. As it can be seen from them, only \code{Math.random} is being inlined.

### 2015-03-08 test/index.tests.js
completes WrappedTypedArray's tests

@@ -91,15 +91,16 @@ describe('initialization', function(){
   });
 });
 
-describe.only('wrap tests', function(){
+describe('wrap tests', function(){
 
   var pjs;
   var errors = require('../src/errors.js');
   var utils = require('../src/utils.js');
+  var parts = 4;
 
   beforeEach(function () {
     pjs = require('../src/index.js');
-    pjs.init({maxWorkers:4});
+    pjs.init({maxWorkers: parts});
   });
 
   afterEach(function(){
@@ -131,6 +132,23 @@ describe.only('wrap tests', function(){
           expect(function () {
             pjs(sourceArray);
           }).to.not.throw(errors.InvalidArgumentsError);
+        });
+
+        it('should initialize packager after wrap a typed array', function () {
+          var wrapped = pjs(sourceArray);
+          expect(wrapped.packager).to.not.be.undefined;
+        });
+
+        it('should have source after wrap a typed array', function () {
+          var wrapped = pjs(sourceArray);
+          expect(wrapped.source).to.not.be.undefined;
+          expect(wrapped.source).to.equal(sourceArray);
+        });
+
+        it('should have parts after wrap a typed array', function () {
+          var wrapped = pjs(sourceArray);
+          expect(wrapped.parts).to.not.be.undefined;
+          expect(wrapped.parts).to.equal(parts);
         });
       });
     });

### 2015-03-08 docs/report/map_serial_vs_parallel.tex
Added comparison of serial vs parallel map

\section{\code{map}: Sequential vs parallel}

Since as explained in \ref{sec:serialization-and-transference} the cost of transferring objects is not zero, it is important to compare the performance of our parallel implementation against a serial one.

In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform. That is because in the parallel case, the source \ttarray{} must be copied to sub-arrays in order to be transferred. On the contratry, the serial implementation can take that time to actually operate on some elements.

\subsection{Algorithm selection}
For this benchmark we decided to use an algorithm that applies a sepia tone effect to pictures. This meets the criteria of both being useful and having some calls to \code{Math.random} which make it more computationally complex.

\subsection{Benchmark code}
The benchmark code is introduced in the following code listing:
\bmcode{http://jsperf.com/pjs-map-vs-serial/3}{../../spikes/3.1.3/comparison/jspBenchmark.js}

\note{Unlike other benchmarks in which we were comparing different ways to achieve the same, this benchmark compares the same thing with different element amounts. That's why the original charts are not displayed.}

The results for the benchmark are displayed in Figure~\ref{fig:map-par-vs-seq}.
\img{Relative performance of parallel implementation with regards to sequential one}{../../spikes/3.1.3/comparison/chart}{fig:map-par-vs-seq}

\subsection{Conclusion}
The parallel improvement might be more noticeable in lower end machines since the linear implementation would be able to operate on less elements ``while the parallel is doing the split'', since the first is a computationally expensive operationa and the other one isn't. This is consistent with our findings based on the machine specifications in \ref{sub-sec:serialization-and-transference}.

\pagebreak

### 2015-03-08 spikes/3.2.3/comparison/jspBenchmark.js
[3.2.3] adds jsperf benchmark for pjs' filter operation

@@ -1,0 +1,110 @@
+// 
+
+// HTML preparation
+<script src="https://rawgit.com/pjsteam/pjs/v0.2.0/dist/p-j-s.min.js"></script>
+<script>
+  var pjs = require('p-j-s');
+  pjs.init();
+
+  var generateElements = function (total) {
+    var typed = new Uint32Array(total);
+    for (var i = total; i > 0; i--){
+      typed[i - 1] = 0xdddddddd;
+    }
+    return typed;
+  };
+  var xs100 = generateElements(100);
+  var wrappedXs100 = pjs(xs100);
+  var xs1000 = generateElements(1000);
+  var wrappedXs1000 = pjs(xs1000);
+  var xs10000 = generateElements(10000);
+  var wrappedXs10000 = pjs(xs10000);
+  var xs100000 = generateElements(100000);
+  var wrappedXs100000 = pjs(xs100000);
+  var xs1000000 = generateElements(1000000);
+  var wrappedXs1000000 = pjs(xs1000000);
+
+  function predicate(pixel) {
+    var r = pixel & 0xFF;
+    var g = (pixel & 0xFF00) >> 8;
+    var b = (pixel & 0xFF0000) >> 16;
+
+    var component_r = r * r * 0.241;
+    var component_g = g * g * 0.691;
+    var component_b = b * b * 0.068;
+
+    var brightness = Math.sqrt(component_r + component_g + component_b);
+
+    return brightness > 130;
+  };
+
+  function serialFilter(pixels, l) {
+    var result = new Uint32Array(l);
+    var newLength = 0;
+    for (var i = 0; i < l; i++) {
+      var pixel = pixels[i];
+      var r = pixel & 0xFF;
+      var g = (pixel & 0xFF00) >> 8;
+      var b = (pixel & 0xFF0000) >> 16;
+
+      var component_r = r * r * 0.241;
+      var component_g = g * g * 0.691;
+      var component_b = b * b * 0.068;
+
+      var brightness = Math.sqrt(component_r + component_g + component_b);
+
+      if (brightness > 130) {
+        result[newLength++] = pixel;
+      }
+    }
+    return new Uint32Array(result).subarray(0, newLength);
+  };
+
+  function runSerial(xs) {
+    var r = serialFilter(xs, xs.length);
+    if (0 === r.length) {
+      console.log('ups serial');
+    }
+  };
+
[…]

### 2015-03-08 dist/p-j-s-standalone.js
Updated to latest dist

@@ -105,14 +105,13 @@ var initialized = false;
 var workers = [];
 
 var WrappedTypedArray = function(source, parts){
-	this.packager = new JobPackager(this.parts, source);
+	this.packager = new JobPackager(parts, source);
 	this.source = source;
 	this.parts = parts;
 };
 
 WrappedTypedArray.prototype.__operationSkeleton = function (f, operation, collectorMapper, done) {
-	var packs = packager.generatePackages(f, operation);
-
+	var packs = this.packager.generatePackages(f, operation);
 	var collector = new ResultCollector(this.parts, function(results){
 		var partial_results = results.map(collectorMapper);
 		var m = merge_typed_arrays(partial_results);

### 2015-03-08 dist/p-j-s.js

@@ -424,14 +424,13 @@ var initialized = false;
 var workers = [];
 
 var WrappedTypedArray = function(source, parts){
-	this.packager = new JobPackager(this.parts, source);
+	this.packager = new JobPackager(parts, source);
 	this.source = source;
 	this.parts = parts;
 };
 
 WrappedTypedArray.prototype.__operationSkeleton = function (f, operation, collectorMapper, done) {
-	var packs = packager.generatePackages(f, operation);
-
+	var packs = this.packager.generatePackages(f, operation);
 	var collector = new ResultCollector(this.parts, function(results){
 		var partial_results = results.map(collectorMapper);
 		var m = merge_typed_arrays(partial_results);

### 2015-03-08 spikes/3.2.3/comparison/jspBenchmark.js
[3.2.3] fixes jsperf benchmark for pjs' filter operation

@@ -1,7 +1,7 @@
-// 
+// http://jsperf.com/pjs-filter-vs-serial
 
 // HTML preparation
-<script src="https://rawgit.com/pjsteam/pjs/v0.2.0/dist/p-j-s.min.js"></script>
+<script src="https://rawgit.com/pjsteam/pjs/v0.2.1/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
   pjs.init();
@@ -79,32 +79,32 @@ __finish = function () {
   deferred.resolve();
 };
 
-// Test Case - serial map 100
+// Test Case - serial filter 100
 runSerial(xs100);
 
-// Test Case - pjs map 100
+// Test Case - pjs filter 100
 runPjs(wrappedXs100);
 
-// Test Case - serial map 1,000
+// Test Case - serial filter 1,000
 runSerial(xs1000);
 
-// Test Case - pjs map 1,000
+// Test Case - pjs filter 1,000
 runPjs(wrappedXs1000);
 
-// Test Case - serial map 10,000
+// Test Case - serial filter 10,000
 runSerial(xs10000);
 
-// Test Case - pjs map 10,000
+// Test Case - pjs filter 10,000
 runPjs(wrappedXs10000);
 
-// Test Case - serial map 100,000
+// Test Case - serial filter 100,000
 runSerial(xs100000);
 
-// Test Case - pjs map 100,000
+// Test Case - pjs filter 100,000
 runPjs(wrappedXs100000);
 
-// Test Case - serial map 1,000,000
+// Test Case - serial filter 1,000,000
 runSerial(xs1000000);
 
-// Test Case - pjs map 1,000,000
+// Test Case - pjs filter 1,000,000
 runPjs(wrappedXs1000000);

### 2015-03-14 docs/report/modules.tex
Added modules doc

\section{Modules system}

\note{At the time of writing, Google Chrome does not support the ES6 module specification so that is not a viable choice.}

The two most known module formats/API in JavaScript are known as:
\begin{description}
\item[CommonJS] Recognized for being used in node.js~\cite{nodejs}.
\item[AMD~\cite{amd}] The most known implementation being require.js~\cite{requirejs}.
\end{description}

CommonJS has many similar aspects with the ES6 module specification, although it does not provide a way to load modules asynchronously. AMD is very different from the ES6 module specification but does support asynchronous module loading which is useful for browser applications.

Instead of making the choice for users, our library will support both systems in a way that is easy for users.

\subsection{CommonJS support}
All that needs to be done to use the library is shown in the following code snippet:
\begin{lstlisting}[language=HTML, caption=CommonJS usage]
<script src="https://rawgit.com/pjsteam/pjs/v0.2.1/dist/p-j-s.min.js"></script>

<script>
  var pjs = require('p-j-s');
  pjs.init();

  /* use below */
</script>
\end{lstlisting}

\subsection{AMD support}
This example shows how to use the library with require.js. It assumes a folder structure like the one shown below:
\begin{lstlisting}[language=bash, caption=AMD folder structure]
.
|__ index.html
|__ scripts
    |__ main.js
    |__ require.js
\end{lstlisting}

The file \textbf{main.js} must have the following contents:
\begin{lstlisting}[caption=main.js AMD support]
require(["https://rawgit.com/pjsteam/pjs/v0.2.1/dist/p-j-s-standalone.min.js"], function(pjs) {
  pjs.init();

  /* use below */
});
\end{lstlisting}

\subsection{Conclusion}
As the previous examples show, using our library with either module system is really simple and is done in a way that should feel familiar for users of those module formats.

\pagebreak

### 2015-03-14 src/Skeleton.js
[3.2.4] adds filter-reduce and map-reduce chaining.

@@ -4,6 +4,7 @@ var merge_typed_arrays = require('./typed_array_merger');
 var utils = require('./utils');
 var operation_names = require('./operation_names');
 var operation_packager = require('./operation_packager');
+var errors = require('./errors');
 
 var finisher = {
   map: function (self, result, done) {
@@ -30,16 +31,21 @@ var Skeleton = function (source, parts, workers, operation, previousOperations) 
 };
 
 Skeleton.prototype.map = function (mapper) {
+  this.__verifyPreviousOperation();
   var operation = operation_packager(operation_names.MAP, mapper);
   return new Skeleton(this.source, this.parts, this.workers, operation, this.operations);
 };
 
 Skeleton.prototype.filter = function (predicate) {
+  this.__verifyPreviousOperation();
   var operation = operation_packager(operation_names.FILTER, predicate);
   return new Skeleton(this.source, this.parts, this.workers, operation, this.operations);
 };
 
-Skeleton.prototype.reduce = function () {
+Skeleton.prototype.reduce = function (predicate, seed, identity) {
+  this.__verifyPreviousOperation();
+  var operation = operation_packager(operation_names.REDUCE, predicate, seed, identity);
+  return new Skeleton(this.source, this.parts, this.workers, operation, this.operations);
 };
 
 Skeleton.prototype.seq = function (done) {
@@ -62,6 +68,12 @@ Skeleton.prototype.seq = function (done) {
 
     workers[index].postMessage(pack, [ pack.buffer ]);
   });
+};
+
+Skeleton.prototype.__verifyPreviousOperation = function () {
+  if (this.operation.name === 'reduce') {
+    throw new errors.InvalidOperationError(errors.messages.INVALID_CHAINING_OPERATION);
+  }
 };
 
 module.exports = Skeleton;

### 2015-03-14 src/errors.js

@@ -19,6 +19,7 @@ InvalidArgumentsError.prototype = new Error();
 InvalidArgumentsError.prototype.constructor = InvalidArgumentsError;
 errors.InvalidArgumentsError = InvalidArgumentsError;
 
+
 errors.messages = {
   CONSECUTIVE_INITS: 'You should not recall init if the library is already initialized.',
   TERMINATE_WITHOUT_INIT: 'You should not terminate pjs if it was not initialized before.',
@@ -29,9 +30,10 @@ errors.messages = {
   INVALID_CODE: 'Invalid code argument to package.',
   INVALID_ELEMENTS: 'Invalid number of elements to package.',
   INVALID_PACKAGE_INDEX: 'Package index should be not negative and less than {0}.',
-  INVALID_TYPED_ARRAY: 'Invalid argument. It should be of TypedArray',
-  INVALID_OPERATION: 'Invalid pjs operation. Possible values are \'filter\', \'map\' or \'reduce\'',
+  INVALID_TYPED_ARRAY: 'Invalid argument. It should be of TypedArray.',
+  INVALID_OPERATION: 'Invalid pjs operation. Possible values are \'filter\', \'map\' or \'reduce\'.',
   INVALID_OPERATIONS: 'Invalid operation chain sent to JobPackager. Either undefined or empty.',
-  MISSING_SEED: 'Missing Seed argument for reduce operation packaging',
-  MISSING_IDENTITY: 'Missing Identity argument for reduce operation packaging'
+  MISSING_SEED: 'Missing Seed argument for reduce operation packaging.',
+  MISSING_IDENTITY: 'Missing Identity argument for reduce operation packaging.',
+  INVALID_CHAINING_OPERATION: 'Can not perform more chaining after reduce operation.'
 };

### 2015-03-14 test/chaining.test.js

@@ -5,6 +5,7 @@ describe('chaining tests', function(){
   var pjs;
   var utils = require('../src/utils.js');
   var Skeleton = require('../src/skeleton');
+  var errors = require('../src/errors');
 
   beforeEach(function () {
     pjs = require('../src/index.js');
@@ -49,9 +50,10 @@ describe('chaining tests', function(){
             var mapper1 = function (e) { return e * 2; };
             var mapper2 = function (e) { return e * 3; };
             wrapped.map(mapper1).map(mapper2).seq(function (result) {
+              var normalChaining = normalSourceArray.map(mapper1).map(mapper2);
+              expect(0 < result.length).to.be.true;
               expect(result).to.have.length(sourceArray.length);
               expect(utils.getTypedArrayType(result)).to.equal(utils.getTypedArrayType(sourceArray));
-              var normalChaining = normalSourceArray.map(mapper1).map(mapper2);
               for (var i = sourceArray.length - 1; i >= 0; i--) {
                 expect(result[i]).to.equal(normalChaining[i]);
               };
@@ -65,11 +67,25 @@ describe('chaining tests', function(){
             var predicate = function (e) { return 1 == (e % 2); };
             wrapped.map(mapper).filter(predicate).seq(function (result) {
               var normalChaining = normalSourceArray.map(mapper).filter(predicate);
+              expect(0 < result.length).to.be.true;
               expect(result).to.have.length(normalChaining.length);
               expect(utils.getTypedArrayType(result)).to.equal(utils.getTypedArrayType(sourceArray));
               for (var i = sourceArray.length - 1; i >= 0; i--) {
                 expect(result[i]).to.equal(normalChaining[i]);
               };
+              done();
+            });
+          });
+
+          it ('sequenced map-reduce should return mapped-and-reduced element in callback', function (done) {
+            var wrapped = pjs(sourceArray);
+            var reducer = function (p, e) { return p + e; };
+            var mapper = function (e) { return e * 2; };
+            var seed = 1;
+            var identity = 0;
+            wrapped.map(mapper).reduce(reducer, seed, identity).seq(function (result) {
+              var normalChaining = normalSourceArray.map(mapper).reduce(reducer, seed);
+              expect(result).to.equal(normalChaining);
               done();
             });
           });
@@ -104,6 +120,7 @@ describe('chaining tests', function(){
             var predicate2 = function (e) { return 0 == e % 4; };
             wrapped.filter(predicate1).filter(predicate2).seq(function (result) {
[…]

### 2015-05-02 docs/paper/paper.tex
Added more content to paper

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{dschenkelman, mservetto}@fi.uba.ar|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Using algorithmic skeletons in EcmaScript to parallelize computation in browsers through Web Workers}

% a short form should be given in case it is too long for the running head
\titlerunning{Algorithmic skeletons to parallelizing computation in browsers}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Damian Schenkelman\and Matias Servetto}
%
\authorrunning{Algorithmic skeletons to parallelizing computation in browsers}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Buenos Aires University, School of Engineering,\\
Paseo Colon 850, Buenos Aires, Argentina\\
\mailsa\\
\url{http://www.fi.uba.ar/}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
With the growing tendency of browsers being used for CPU intensive applications (such as 3D games) it is important to take advantage of multiple cores to reduce user perceived latency. This paper presents a JavaScript library based on algorithmic skeletons that allows users to take advantage of parallel processing in browsers while maintaining a familiar coding style.
\keywords{browsers, ecmascript, parallelization, algorithmic skeletons}
\end{abstract}

\section{Introduction}

In 1965 Gordon Moore proposed the popularly known ``Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
As a consequence, programs that use a single processor/core will automatically become faster without the need to modify them with that purpose in mind.

In the year 2010 it was predicted that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation. This is the cause why newer computer have a greater number of processors instead of processors with more computing power.

To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors. In the case of application executed on a single device (e.g. browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing. For example, the .NET platform has the Task Parallel Library (TPL) and Parallel LINQ (PLINQ) \cite{ms-par} libraries that use algorithmic skeletons modelled with high order functions so developers can perform complex operations through a simple API.

On a different note, one of the tendencies that is beggining to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications that were previously only considered viable in a native environment, such as video games. This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications. Initiatives such as ParallelJS \cite{par-js} and River trail \cite{rivertrail} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.

The goal of this paper is to provide an alternative for parallel code execution in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce. This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

\section{Serialization and Transference}
Let's consider an array of \verb+N+ elements on which a particular transformation is to be performed through the \verb+map+ function. If the transformation is executed using a single thread (no parallelism) and the average time to process each element is \verb+t+ then the total time (\(T_{ser}\)) for the operation can be approximated as:

\begin{equation}
  T_{ser} = \sum_{i = 0}^{N}t = Nt \;  .
\end{equation}

When trying to parallelize this operation using \(K\) threads the ideal goal is to reach a total time (\(T_{par}\)) that is:

\begin{equation}
  T_{par} = \frac{T_{ser}}{K} \;  .
\end{equation}

Nevertheless, there are additional time consuming tasks other than the main computation that need to be considered when performing the operation in parallel in programs where not all memory is shared. These are:
\begin{itemize}
  \item Serializing/deserializing the elements to transfer.
  \item Transferring the elements back and forth between the UI thread and the workers.
  \item Transferring to each worker the functions for the transformation.
\end{itemize}

If we consider \(T_{ft}\) as the function transfer time, \(T_{et}\) as the elements transfer time, and \(T_{s}\) as the serialization/deserialization time it is clear that:

\begin{equation}
  T_{sync} = T_{ft} + T_{et} + T_{s} \;  .
\end{equation}

\begin{equation}
  T_{par} \approx \frac{T_{ser}}{K} + T_{sync} \;  .
\end{equation}

Based on the previous equations it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.

In our case we are trying to transfer objects between a browser's JavaScript UI thread and Web Workers so we are constrained by the means of that environment. The Worker interface is the following \cite{w3c-ww}:
\medskip

\noindent
{\it Web Worker interface}
\begin{verbatim}
[Constructor(DOMString scriptURL)]
interface Worker : EventTarget {
  void terminate();

  void postMessage(any message, optional sequence<Transferable> transfer);
  [TreatNonCallableAsNull] attribute Function? onmessage;
};
Worker implements AbstractWorker;
\end{verbatim}
%
\noindent
{\small (Example extracted from the W3C Web Workers specification)}

As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects. From section \textbf{2.7.5} of the HTML Standard \cite{html-whatwg}:
\textit{``Some objects support being copied and closed in one operation. This is called transferring the object, and is used in particular to transfer ownership of unsharable or expensive resources across worker boundaries."}

\subsection{Elements}
Considering the definition of a \verb+Transferable+ the latter seems like a good alternative to minimize both \(T_{et}\) and \(T_{s}\). Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.6} of that same standard).

To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+. Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1000000 (a million). The results are in Tab.~\ref{tab:transf-vs-clone}. As it can be seen from the  results, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales. On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.

For that reason our library only works with \verb+SharedTypedArray+s and \verb+TypedArrays+.

\begin{table}
  \centering
  \begin{tabular}{|l|l|}
    \hline
    Computer & Mac Book Air \\
    Processor & Intel Core i5 \\
    Clock Frequency & 1.8 GHz \\
    System Memory & 4 GB \\
    Operating System & OS X Yosemite 10.10.3 \\
    Browser version & Firefox Nightly 40.0a1 (2015-04-26) \\
    \hline
  \end{tabular}
  \caption{Benchmarks environment.}
  \label{tab:env}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    Elements & Cloning [ops/sec] & Transferring [ops/sec] \\
    \hline
    \(1E5\) & 1000 & 9000 \\
    \(1E6\) & 100 & 16000 \\
    \hline
  \end{tabular}
  \caption{Difference between transferring and cloning shared buffers.}
  \label{tab:transf-vs-clone}
\end{table}

\section{Paper Preparation}

Springer provides you with a complete integrated \LaTeX{} document class
(\texttt{llncs.cls}) for multi-author books such as those in the LNCS
series. Papers not complying with the LNCS style will be reformatted.
This can lead to an increase in the overall number of pages. We would
therefore urge you not to squash your paper.

Please always cancel any superfluous definitions that are
not actually used in your text. If you do not, these may conflict with
the definitions of the macro package, causing changes in the structure
of the text and leading to numerous mistakes in the proofs.

If you wonder what \LaTeX{} is and where it can be obtained, see the
``\textit{LaTeX project site}'' (\url{http://www.latex-project.org})
and especially the webpage ``\textit{How to get it}''
(\url{http://www.latex-project.org/ftp.html}) respectively.

When you use \LaTeX\ together with our document class file,
\texttt{llncs.cls},
your text is typeset automatically in Computer Modern Roman (CM) fonts.
Please do
\emph{not} change the preset fonts. If you have to use fonts other
than the preset fonts, kindly submit these with your files.

Please use the commands \verb+\label+ and \verb+\ref+ for
cross-references and the commands \verb+\bibitem+ and \verb+\cite+ for
references to the bibliography, to enable us to create hyperlinks at
these places.

For preparing your figures electronically and integrating them into
your source file we recommend using the standard \LaTeX{} \verb+graphics+ or
\verb+graphicx+ package. These provide the \verb+\includegraphics+ command.
In general, please refrain from using the \verb+\special+ command.

Remember to submit any further style files and
fonts you have used together with your source files.

\subsubsection{Headings.}

Headings should be capitalized
(i.e., nouns, verbs, and all other words
except articles, prepositions, and conjunctions should be set with an
initial capital) and should,
with the exception of the title, be aligned to the left.
Words joined by a hyphen are subject to a special rule. If the first
word can stand alone, the second word should be capitalized.

Here are some examples of headings: ``Criteria to Disprove
Context-Freeness of Collage Language", ``On Correcting the Intrusion of
Tracing Non-deterministic Programs by Software", ``A User-Friendly and
Extendable Data Distribution System", ``Multi-flip Networks:
Parallelizing GenSAT", ``Self-determinations of Man".

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems, etc. should
appear in consecutive order, starting with Lemma 1, and not, for
example, with Lemma 11.

\subsection{Figures}

For \LaTeX\ users, we recommend using the \emph{graphics} or \emph{graphicx}
package and the \verb+\includegraphics+ command.

Please check that the lines in line drawings are not
interrupted and are of a constant width. Grids and details within the
figures must be clearly legible and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi). The lettering in figures should have a height of
2~mm (10-point type). Figures should be numbered and should have a
caption which should always be positioned \emph{under} the figures, in
contrast to the caption belonging to a table, which should always appear
\emph{above} the table; this is simply achieved as matter of sequence in
your source.

Please center the figures or your tabular material by using the \verb+\centering+
declaration. Short captions are centered by default between the margins
and typeset in 9-point type (Fig.~\ref{fig:example} shows an example).
The distance between text and figure is preset to be about 8~mm, the
distance between figure and caption about 6~mm.

To ensure that the reproduction of your illustrations is of a reasonable
quality, we advise against the use of shading. The contrast should be as
pronounced as possible.

If screenshots are necessary, please make sure that you are happy with
the print quality before you send the files.
\begin{figure}
\centering
\includegraphics[height=6.2cm]{eijkel2}
\caption{One kernel at $x_s$ (\emph{dotted kernel}) or two kernels at
$x_i$ and $x_j$ (\textit{left and right}) lead to the same summed estimate
at $x_s$. This shows a figure consisting of different types of
lines. Elements of the figure described in the caption should be set in
italics, in parentheses, as shown in this sample caption.}
\label{fig:example}
\end{figure}

Please define figures (and tables) as floating objects. Please avoid
using optional location parameters like ``\verb+[h]+" for ``here".

\paragraph{Remark 1.}

In the printed volumes, illustrations are generally black and white
(halftones), and only in exceptional cases, and if the author is
prepared to cover the extra cost for color reproduction, are colored
pictures accepted. Colored pictures are welcome in the electronic
version free of charge. If you send colored figures that are to be
printed in black and white, please make sure that they really are
legible in black and white. Some colors as well as the contrast of
converted colors show up very poorly when printed in black and white.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin --
which is the default if you use the \emph{equation} environment, e.g.,
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Equations should be punctuated in the same way as ordinary
text but with a small space before the end punctuation mark.

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or period). Footnotes should appear at the bottom of
the
normal text area, with a line of about 2~cm set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing.}

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
report. Springer, New York)}

\subsection{Citations}

For citations in the text please use
square brackets and consecutive numbers: \cite{jour}, \cite{lncschap},
\cite{proceeding1} -- provided automatically
by \LaTeX 's \verb|\cite| \dots\verb|\bibitem| mechanism.

\subsection{Page Numbering and Running Heads}

There is no need to include page numbers. If your paper title is too
long to serve as a running head, it will be shortened. Your suggestion
as to how to shorten it would be most welcome.

\section{LNCS Online}

The online version of the volume will be available in LNCS Online.
Members of institutes subscribing to the Lecture Notes in Computer
Science series have access to all the pdfs of all the online
publications. Non-subscribers can only read as far as the abstracts. If
they try to go beyond this point, they are automatically asked, whether
they would like to order the pdf, and are given instructions as to how
to do so.

Please note that, if your email address is given in your paper,
it will also be included in the meta data of the online version.

\section{BibTeX Entries}

The correct BibTeX entries for the Lecture Notes in Computer Science
volumes can be found at the following Website shortly after the
publication of the book:
\url{http://www.informatik.uni-trier.de/~ley/db/journals/lncs.html}

\subsubsection*{Acknowledgments.} The heading should be treated as a
subsubsection heading and should not be assigned a number.

\section{The References Section}\label{references}

In order to permit cross referencing within LNCS-Online, and eventually
between different publishers and their online databases, LNCS will,
from now on, be standardizing the format of the references. This new
feature will increase the visibility of publications and facilitate
academic research considerably. Please base your references on the
examples below. References that don't adhere to this style will be
reformatted by Springer. You should therefore check your references
thoroughly when you receive the final pdf of your paper.
The reference section must be complete. You may not omit references.
Instructions as to where to find a fuller version of the references are
not permissible.

We only accept references written using the latin alphabet. If the title
of the book you are referring to is in Russian or Chinese, then please write
(in Russian) or (in Chinese) at the end of the transcript or translation
of the title.

The following section shows a sample reference list with entries for
journal articles \cite{jour}, an LNCS chapter \cite{lncschap}, a book
\cite{book}, proceedings without editors \cite{proceeding1} and
\cite{proceeding2}, as well as a URL \cite{url}.
Please note that proceedings published in LNCS are not cited with their
full titles, but with their acronyms!

\begin{thebibliography}{4}

\bibitem{moore65} Moore, G.E.: Cramming More Components onto Integrated Circuits. (1965)

\bibitem{ms-par} Parallel Programming in the .NET Framework, Microsoft Corp., \url{http://msdn.microsoft.com/en-us/library/dd460693(v=vs.110).aspx}

\bibitem{es-web} EcmaScript, ECMA, \url{http://www.ecmascript.org/}

\bibitem{asm-web} asm.js, \url{http://asmjs.org/}

\bibitem{webgl-web} WebGL, Khronos Group, \url{http://www.khronos.org/webgl/}

\bibitem{rivertrail} Herhut, S., Hudson, R. L., Shpeisman, T., Sreeram, J.: River trail: A path to parallelism in javascript. SIGPLAN Not., 48(10):729--744, (2013).

\bibitem{par-js} Wang, J., Rubin, N., Yalamanchili, S.: River trail: ParallelJS: An Execution Framework for JavaScript on Heterogeneous Systems. SIGPLAN Not. (2014)

\bibitem{js-simd} McCutchan, J., Feng, H., Matsakis, N. D., Anderson, Z., Jensen, P.: A SIMD Programming Model for Dart, JavaScript, and other dynamically typed scripting languages (2014)

\bibitem{w3c-ww} Web Workers, W3C, \url{http://www.w3.org/TR/workers/}

\bibitem{html-whatwg} HTML Standard, WHATWG, \url{https://html.spec.whatwg.org}

\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
Subsequences. J. Mol. Biol. 147, 195--197 (1981)

\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
Composing a Complex Biological Workflow through Web Services. In: Nagel,
W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
pp. 1148--1158. Springer, Heidelberg (2006)

\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
Infrastructure. Morgan Kaufmann, San Francisco (1999)

\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
Information Services for Distributed Resource Sharing. In: 10th IEEE
International Symposium on High Performance Distributed Computing, pp.
181--184. IEEE Press, New York (2001)

\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
Grid: an Open Grid Services Architecture for Distributed Systems
Integration. Technical report, Global Grid Forum (2002)

\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\section*{Appendix: Springer-Author Discount}

LNCS authors are entitled to a 33.3\% discount off all Springer
publications. Before placing an order, the author should send an email, 
giving full details of his or her Springer publication,
to \url{orders-HD-individuals@springer.com} to obtain a so-called token. This token is a
number, which must be entered when placing an order via the Internet, in
order to obtain the discount.

\section{Checklist of Items to be Sent to Volume Editors}
Here is a checklist of everything the volume editor requires from you:

\begin{itemize}
\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
\itemsep8pt\relax
\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}

\item The final \LaTeX{} source files
\item A final PDF file
\item A copyright form, signed by one author on behalf of all of the
authors of the paper.
\item A readme giving the name and email address of the
corresponding author.
\end{itemize}
\end{document}

### 2015-05-02 examples/sapiaPicture/default-pjs.js
fixes shared sapia picture sample.

@@ -46,7 +46,7 @@ var run;
       var start = new Date();
       pjs(copyData).map('f').seq(function(err, result) {
         var diff = new Date() - start;
-        canvasData.data.set(new Uint8ClampedArray(result.buffer));
+        canvasData.data.set(new SharedUint8ClampedArray(result.buffer));
         tempContext.putImageData(canvasData, 0, 0);
         log.innerHTML = "Process done in " + diff + " ms";
         runButton.style.visibility = "visible";

### 2015-05-02 docs/paper/paper.tex
More content

@@
   \item Transferring the elements back and forth between the UI thread and the workers.
+  \item Serializing/deserializing the function to transfer.
   \item Transferring to each worker the functions for the transformation.
@@
 
-If we consider \(T_{ft}\) as the function transfer time, \(T_{et}\) as the elements transfer time, and \(T_{s}\) as the serialization/deserialization time it is clear that:
+If we drill down into the different parts:
+\begin{itemize}
+  \item \(T_{ft}\) as the function transfer time
+  \item \(T_{et}\) as the elements transfer time
+  \item \(T_{fs}\) as the function serialization/deserialization time
+  \item \(T_{es}\) as the elements serialization/deserialization time
+\end{itemize}
 
+It is clear that:
 \begin{equation}
-  T_{sync} = T_{ft} + T_{et} + T_{s} \;  .
+  T_{sync} = T_{ft} + T_{et} + T_{fs} + T_{es} \;  .
+  \label{eq-tsync}
 \end{equation}
@@
   T_{par} \approx \frac{T_{ser}}{K} + T_{sync} \;  .
+  \label{eq-tpar}
 \end{equation}
 
-Based on the previous equations it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.
+Based on Eq.~\ref{eq-tsync} and Eq.~\ref{eq-tpar} it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.
 
@@
 \subsection{Elements}
-Considering the definition of a \verb+Transferable+ the latter seems like a good alternative to minimize both \(T_{et}\) and \(T_{s}\).
+Considering the definition of a \verb+Transferable+, it seems like a good alternative to minimize both \(T_{et}\) and \(T_{es}\).
 Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.6} of that same standard).
@@
 Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1000000 (a million).
-The results are in Tab.~\ref{tab:transf-vs-clone}.
-As it can be seen from the  results, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales.
+As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales.
 On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.
@@
   \label{tab:transf-vs-clone}
+\end{table}
+
+\subsection{Functions}
+The library must the distribution of the code between the different workers when an operation is executed.
+This is a challenging problem with a lot of possible alternatives \cite{cloud-haskell} \cite{distributed-scheme}.
+
+A naive approach would be to try to transfer a function as part of the \verb+message+ parameter of \verb+postMessage+ but that is not an option (it throws an error).
+
+Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered (Sec.~\ref{sec:context-sharing} proposes a solution to sharing additional data and functions between the UI thread and workers) a possible alternative is to:
+\begin{enumerate}
+  \item Serialize the \verb+Function+ to a \verb+String+.
+  \item Transfer the \verb+String+ to the worker.
+  \item Create a new \verb+Function+ on the worker from that \verb+String+.
+\end{enumerate}
+
+\subsubsection{Serialization and deserialization}
+One possible way of passing serialized \verb+Function+ to workers would be to encode them as binary and transfer them as an \verb+ArrayBuffer+. To encode the strings there are two possible approaches:
+\begin{itemize}
+  \item Using the Encoding API\cite{encoding-api}
+  \item Implementing non native encoding/decoding functions
+\end{itemize}
+
+The benchmarks provided the following results:
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|}
+    \hline
+    Native encoding API [ops/sec] & Non-native function [ops/sec] \\
+    \hline
+    \textbf{5247} & 4685 \\
+    \hline
+  \end{tabular}
+  \caption{Difference encoding/decoding strings natively and in JavaScript.}
+  \label{tab:encode-decode}
+\end{table}
+
+\subsubsection{Transference}
+\verb+Function+'s code not only needs to be serialized but also transferred.
+For that reason it was also worth comparing the time it takes to encode/decode a \verb+String+ and tranfer the resulting \verb+ArrayBuffer+ to the worker against just passing the \verb+String+ to the worker.
+
+Additionally, in most scenarios code would have to be sent along with a \verb+TypedArray+ so it would also be interesting to see if the transfer time was affected by this fact.
+
+We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time (results in Tab.~\ref{tab:send-function}).
+We measured three approaches to transfer the string:
+\begin{enumerate}
+  \item Encoding the \verb+String+ into an \verb+TypedArrayBuffer+, sending it as a \verb+Tranferable+ and decoding it on the Web Worker.
+  \item Sending the \verb+String+ directly.
+  \item Using a \verb+Blob+ to store the stream and retrieving that in the Worker.
+\end{enumerate}
+
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    Function length & Blob [ops/sec] & Copy [ops/sec] & Tranferable [ops/sec] \\
+    \hline
+    Short & 1251 & \textbf{7046} & 6219 \\
+    Long & 705 & 1176 & \textbf{1198} \\
+    \hline
+  \end{tabular}
+  \caption{Sending function strings from UI thread to Web Workers.}
+  \label{tab:send-function}
+\end{table}
+
+We also created a benchmark that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}.
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    Copying code [ops/sec] & Tranferring code [ops/sec] \\
+    \hline
+    4200 & \textbf{4340} \\
+    \hline
+  \end{tabular}
+  \caption{Sending code and a TypedArray to a Web Worker.}
+  \label{tab:send-function-and-array}
 \end{table}
@@
 
+\bibitem{cloud-haskell} Epstein, J., Black, A. P., Peyton-Jones, S.: Towards Haskell in the Cloud (2011)
+
+\bibitem{distributed-scheme} Schwendner, A.: Distributed Functional Programming in Scheme (2010)
+
+\bibitem{encoding-api} Encoding Standard, WHATWG, \url{https://encoding.spec.whatwg.org}
+
+% end of our comments, start of sample comments
+
 \bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular

### 2015-05-02 dist/p-j-s-standalone.js
adds benchmark results for shared paper.

@@ -991,9 +991,16 @@ var globalContext = {};
 
 var operations = {
   map: function (source, target, start, end, f, ctx) {
+    /*
+    var i = start | 0;
+    for ( ; (i | 0) < (end | 0); i = (i | 0) + (1 | 0)){
+      target[i | 0] = f(source[i | 0], ctx);
+    }
+    return end;
+    */
     var i = start;
     for ( ; i < end; i += 1){
-      target[i] = f(source[i], ctx);
+      target[i] = f(source[i]);
     }
     return end;
   },
@@ -1074,7 +1081,6 @@ module.exports = function(event){
     } else {
       localCtx = globalContext;
     }
-
     newEnd = operations[operation.name](sourceArray, targetArray, start, newEnd, f, localCtx, seed);
     sourceArray = targetArray;
   }

### 2015-05-02 dist/p-j-s.js

@@ -914,9 +914,16 @@ var globalContext = {};
 
 var operations = {
   map: function (source, target, start, end, f, ctx) {
+    /*
+    var i = start | 0;
+    for ( ; (i | 0) < (end | 0); i = (i | 0) + (1 | 0)){
+      target[i | 0] = f(source[i | 0], ctx);
+    }
+    return end;
+    */
     var i = start;
     for ( ; i < end; i += 1){
-      target[i] = f(source[i], ctx);
+      target[i] = f(source[i]);
     }
     return end;
   },
@@ -997,7 +1004,6 @@ module.exports = function(event){
     } else {
       localCtx = globalContext;
     }
-
     newEnd = operations[operation.name](sourceArray, targetArray, start, newEnd, f, localCtx, seed);
     sourceArray = targetArray;
   }

### 2015-05-02 examples/sapiaPicture/default-pjs.js

@@ -6,7 +6,7 @@ var run;
   var source = document.getElementById("source");
   var runButton = document.getElementById("runButton");
   var pjs = require('p-j-s');
-  pjs.init({ maxWorkers: 2 });
+  pjs.init({ maxWorkers: 4 });
   pjs.updateContext({
     f:function(pixel){
       // return pixel;

### 2015-05-02 spikes/shared/paper-bechmarks/section-4.1/jspBenchmark-long.js

@@ -1,0 +1,1 @@
+// http://jsperf.com/longer-transferrable-vs-cloning/2

### 2015-05-02 spikes/shared/paper-bechmarks/section-4.1/jspBenchmark-short.js

@@ -1,0 +1,1 @@
+// http://jsperf.com/transferrable-vs-cloning/2

### 2015-05-02 spikes/shared/paper-bechmarks/section-5.1/jspBenchmark.js

@@ -1,0 +1,1 @@
+// http://jsperf.com/pjs-serialization-comparison/2

### 2015-05-02 spikes/shared/paper-bechmarks/section-5.2.1/jspBenchmark-long.js

@@ -1,0 +1,1 @@
+// http://jsperf.com/pjs-serialization-long/2

### 2015-05-02 src/worker_core.js

@@ -26,9 +26,16 @@ var globalContext = {};
 
 var operations = {
   map: function (source, target, start, end, f, ctx) {
+    /*
+    var i = start | 0;
+    for ( ; (i | 0) < (end | 0); i = (i | 0) + (1 | 0)){
+      target[i | 0] = f(source[i | 0], ctx);
+    }
+    return end;
+    */
     var i = start;
     for ( ; i < end; i += 1){
-      target[i] = f(source[i], ctx);
+      target[i] = f(source[i]);
     }
     return end;
   },
@@ -109,7 +116,6 @@ module.exports = function(event){
     } else {
       localCtx = globalContext;
     }
-
     newEnd = operations[operation.name](sourceArray, targetArray, start, newEnd, f, localCtx, seed);
     sourceArray = targetArray;
   }

### 2015-05-09 docs/paper/paper.tex
Updated benchmark
Paper background section done

@@
 % unless you accept that it will be published
-\institute{Buenos Aires University, School of Engineering,\\
+\institute{School of Engineering. University of Buenos Aires.\\
 Paseo Colon 850, Buenos Aires, Argentina\\
@@
 This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.
+
+\section{Background}
+All modern versions of major browsers (Internet Explorer, Mozilla Firefox, Google Chrome, Safari and Opera) allow the execution of JavaScript code in a single threaded event loop runtime \cite{event-loop}.
+User actions add elements to the event loop queue and these are processesed sequentially.
+All process input/output is asynchronous to prevent the UI thread from blocking, thus keeping it responsive.
+While such a runtime is a good fit for most web applications, JavaScript and the browser are now also being used for some CPU intensive tasks, such as games, which commonly require physics calculations, image manipulation (in 2D) and graphics generation (in 3D).
+Despite the large improvement in JavaScript performance obtained through optimizing compilers such as Google Chrome's V8 \cite{v8} and Mozilla Firefox's SpiderMonkey \cite{spider-monkey}, it is important to be able to take advantage of multiple cores in modern devices to achieve even greater performance.\\
+In 2010, Web Workers were made part of the web standard.
+Web Workers allow the creation of ``thread like'' constructs in a browser environment, but they don't allow on shared memory and instead communicate via message passing.
+The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
+However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing the notion of shared memory through \verb+SharedArrayBuffer+s \cite{sab}.
+In essence, the proposal allows the same memory to be shared across multiple Web Workers, and also aims to provide the necessary atomic/lock constructs to deal with shared memory.
+In the particular case of embarrassingly parallel computations (such as map, filter and reduce operations on an array), it is simple to take advantage of the shared memory speed benefits without incurring in any overhead due to locks.
 
@@
     Operating System & OS X Yosemite 10.10.3 \\
-    Browser version & Firefox Nightly 40.0a1 (2015-04-26) \\
+    Browser version & Firefox Nightly 40.0a1 \\
     \hline
@@
     \hline
-    \(1E5\) & 1000 & 9000 \\
-    \(1E6\) & 100 & 16000 \\
+    \(1E5\) & 972 & \textbf{1705} \\
+    \(1E6\) & 114 & \textbf{754} \\
     \hline
@@
 \subsection{Functions}
-The library must the distribution of the code between the different workers when an operation is executed.
+The library must handle the distribution of the code between the different workers when an operation is executed.
 This is a challenging problem with a lot of possible alternatives \cite{cloud-haskell} \cite{distributed-scheme}.
@@
 
+\bibitem{event-loop} Concurrency model and Event Loop, Mozilla Developer Network, \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/EventLoop}
+
+\bibitem{v8} v8, Google, \url{https://developers.google.com/v8/intro}
+
+\bibitem{spider-monkey} SpiderMonkey, Mozilla, \url{https://developer.mozilla.org/en-US/docs/Mozilla/Projects/SpiderMonkey}
+
+\bibitem{thread-pool} Schmidt, D. C., Vinoski, S.,: Comparing Alternative Programming Techniques for Multi-threaded CORBA Servers: Thread Pool (1996)
+
+\bibitem{sab} JavaScript Shared Memory, Atomics, and Locks, \url{https://docs.google.com/document/d/1NDGA_gZJ7M7w1Bh8S0AoDyEqwDdRh4uSoTPSNn77PFk/edit}
+
 \bibitem{html-whatwg} HTML Standard, WHATWG, \url{https://html.spec.whatwg.org}

### 2015-05-09 test/chrome_version_helper.js
fixes promoses tests.

@@ -1,6 +1,6 @@
 'use strict';
 
-module.exports = function (minVersion, callback) {
+module.exports = function (minVersion, callback, shouldIgnoreNotMatchingCallback) {
   var match = window.navigator.appVersion.match(/Chrome\/(\d+)\./);
   if (match) {
     var chromeVersion = parseInt(match[1], 10);
@@ -8,6 +8,8 @@ module.exports = function (minVersion, callback) {
       callback();
     }
   } else {
-    callback();
+    if (!shouldIgnoreNotMatchingCallback) {
+      callback();
+    }
   }
 };

### 2015-05-09 test/promises.test.js

@@ -134,6 +134,6 @@ describe('promises tests', function () {
           done();
         });
       });
-    });
+    }, true);
   });
 });

### 2015-05-09 docs/paper/paper.tex
Added comparison

@@
 
-\section{Background}
+This paper is organized as follows: Section~\ref{sec-background} provides some context on the current state of thread-base parallelization in JavaScript.
+Section~\ref{sec-overview} provides an overview of how the library works.
+Section~\ref{sec-ser-transf} explains the challenges the library faces, the alternative solutions and the decisions taken.
+Section~\ref{sec-bench} compares the performance of using the library against a sequential approach for an image transformation.
+Finally, Section~\ref{sec-future} presents future research directions.
+
+\section{Background}\label{sec-background}
 All modern versions of major browsers (Internet Explorer, Mozilla Firefox, Google Chrome, Safari and Opera) allow the execution of JavaScript code in a single threaded event loop runtime \cite{event-loop}.
@@
 
-\section{Serialization and Transference}
+\section{Library Overview}\label{sec-overview}
+JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs.
+In JavaScript, \verb+Array+s have \verb+map+, \verb+filter+, and \verb+reduce+ functions and version 6 of the EcmaScript adds those same methods to \verb+TypedArray+s \cite{es6-ta}.
+For example, to multiply all the elements of an \verb+Array+ by two:
+\medskip
+
+\noindent
+{\it plain JS map sample}
+\begin{verbatim}
+var newValues = [1,2,3,4].map(function(e){
+  return e * 2;
+});
+
+// newValues is a new array [2,4,6,8]
+\end{verbatim}
+%
+\noindent
+{\small (Example creating a new array with the values of the original one multiplied by two)}\\
+
+Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec-ser-transf}.
+It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation.
+Given a \verb+SharedTypedArray+, the following code would do the same as the previous example but parallelizing the work:
+\medskip
+
+\noindent
+{\it p-j-s map sample}
+\begin{verbatim}
+var pjs = require('p-j-s');
+pjs.init({ maxWorkers: 4 });
+var xs = new SharedUint8Array(4);
+xs.set([1,2,3,4]);
+pjs(xs).map(function(e){
+  return e * 2;
+}).seq().then(function(newValues){
+  // newValues is a new array [2,4,6,8]
+});
+\end{verbatim}
+%
+\noindent
+{\small (Example creating a new array with the values of the original one multiplied by two using p-j-s)}\\
+
+The call to \verb+pjs.init+ initializes the library and the amount of workers to be used for processing (a worker pool is created with them).
+In browsers that have an API to detect the amount of available cores in a machine, the \verb+maxWorkers+ value is limited by the amount of cores.
+
+The API provided by p-j-s \cite{pjs-api} is very similar to the synchronous one that is provided by the language.
+The asynchronous nature of the operation is reflected by the fact that the call to \verb+seq+ returns a \verb+Promise+ (alternatively a callback could be passed as a parameter to it).
+It is possible to chain multiple operations before calling \verb+seq+ to avoid passing data back and forth between the Web Workers and the UI thread.
+
+\section{Serialization and Transference}\label{sec-ser-transf}
 Let's consider an array of \verb+N+ elements on which a particular transformation is to be performed through the \verb+map+ function.
@@
   \centering
-  \begin{tabular}{|c|c|c|c|}
+  \begin{tabular}{|c|c|}
     \hline
@@
 
-\section{Paper Preparation}
+\section{Experiment}
+It is important to compare the performance of our parallel implementation against a serial one, as some processing is required before the computation begins and after it ends to distribute the data and work and gather the results respectively, as explained in Section~\ref{sec-ser-transf}.
 
-Springer provides you with a complete integrated \LaTeX{} document class
-(\texttt{llncs.cls}) for multi-author books such as those in the LNCS
-series. Papers not complying with the LNCS style will be reformatted.
-This can lead to an increase in the overall number of pages. We would
-therefore urge you not to squash your paper.
+In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform.
+That is because in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.
 
-Please always cancel any superfluous definitions that are
-not actually used in your text. If you do not, these may conflict with
-the definitions of the macro package, causing changes in the structure
-of the text and leading to numerous mistakes in the proofs.
+\subsection{Algorithm selection}
+For the experiment we decided to use an algorithm that applies a sepia tone effect to pictures.
+This meets the criteria of both being useful and having some calls to \verb+Math.random+ which make it more complex from a computational standpoint.
+The benchmark is publicly available \cite{benchmark}.
 
-If you wonder what \LaTeX{} is and where it can be obtained, see the
-``\textit{LaTeX project site}'' (\url{http://www.latex-project.org})
-and especially the webpage ``\textit{How to get it}''
-(\url{http://www.latex-project.org/ftp.html}) respectively.
+\subsection{Results}
+The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark}.
+\begin{table}
+  \centering
+  \begin{tabular}{|r|c|c|}
+    \hline
+    Elements & Serial [ops/sec] & Parallel [ops/sec] \\
+    \hline
+    100 & \textbf{154442} & 377.2\\
+    1000 & \textbf{17460.2} &  368.4 \\
+    10000 & \textbf{1768.6} & 311.2 \\
+    100000 & \textbf{179.2} & 172.2 \\
+    1000000 & 16.96 & \textbf{33.2} \\
+    \hline
+  \end{tabular}
+  \caption{Comparing p-j-s (parallel) and plain JS (serial) performance for an image transformation.}
+  \label{tab:benchmark}
+\end{table}
 
-When you use \LaTeX\ together with our document class file,
-\texttt{llncs.cls},
-your text is typeset automatically in Computer Modern Roman (CM) fonts.
-Please do
-\emph{not} change the preset fonts. If you have to use fonts other
-than the preset fonts, kindly submit these with your files.
+Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
 
-Please use the commands \verb+\label+ and \verb+\ref+ for
-cross-references and the commands \verb+\bibitem+ and \verb+\cite+ for
-references to the bibliography, to enable us to create hyperlinks at
-these places.
-
-For preparing your figures electronically and integrating them into
-your source file we recommend using the standard \LaTeX{} \verb+graphics+ or
-\verb+graphicx+ package. These provide the \verb+\includegraphics+ command.
-In general, please refrain from using the \verb+\special+ command.
-
-Remember to submit any further style files and
-fonts you have used together with your source files.
-
-\subsubsection{Headings.}
-
-Headings should be capitalized
-(i.e., nouns, verbs, and all other words
-except articles, prepositions, and conjunctions should be set with an
-initial capital) and should,
-with the exception of the title, be aligned to the left.
-Words joined by a hyphen are subject to a special rule. If the first
-word can stand alone, the second word should be capitalized.
-
-Here are some examples of headings: ``Criteria to Disprove
-Context-Freeness of Collage Language", ``On Correcting the Intrusion of
-Tracing Non-deterministic Programs by Software", ``A User-Friendly and
-Extendable Data Distribution System", ``Multi-flip Networks:
-Parallelizing GenSAT", ``Self-determinations of Man".
-
-\subsubsection{Lemmas, Propositions, and Theorems.}
-
-The numbers accorded to lemmas, propositions, and theorems, etc. should
-appear in consecutive order, starting with Lemma 1, and not, for
-example, with Lemma 11.
-
-\subsection{Figures}
-
-For \LaTeX\ users, we recommend using the \emph{graphics} or \emph{graphicx}
-package and the \verb+\includegraphics+ command.
-
-Please check that the lines in line drawings are not
-interrupted and are of a constant width. Grids and details within the
-figures must be clearly legible and may not be written one on top of
-the other. Line drawings should have a resolution of at least 800 dpi
-(preferably 1200 dpi). The lettering in figures should have a height of
-2~mm (10-point type). Figures should be numbered and should have a
-caption which should always be positioned \emph{under} the figures, in
-contrast to the caption belonging to a table, which should always appear
-\emph{above} the table; this is simply achieved as matter of sequence in
-your source.
-
-Please center the figures or your tabular material by using the \verb+\centering+
-declaration. Short captions are centered by default between the margins
-and typeset in 9-point type (Fig.~\ref{fig:example} shows an example).
-The distance between text and figure is preset to be about 8~mm, the
-distance between figure and caption about 6~mm.
-
-To ensure that the reproduction of your illustrations is of a reasonable
-quality, we advise against the use of shading. The contrast should be as
-pronounced as possible.
-
-If screenshots are necessary, please make sure that you are happy with
-the print quality before you send the files.
 \begin{figure}
 \centering
-\includegraphics[height=6.2cm]{eijkel2}
-\caption{One kernel at $x_s$ (\emph{dotted kernel}) or two kernels at
-$x_i$ and $x_j$ (\textit{left and right}) lead to the same summed estimate
-at $x_s$. This shows a figure consisting of different types of
-lines. Elements of the figure described in the caption should be set in
-italics, in parentheses, as shown in this sample caption.}
-\label{fig:example}
+\includegraphics[height=6.2cm]{comparison}
+\caption{Relative performance of parallel transformation compared to sequential one.}
+\label{fig:relative}
 \end{figure}
 
-Please define figures (and tables) as floating objects. Please avoid
-using optional location parameters like ``\verb+[h]+" for ``here".
+\subsection{Conclusion}
+A couple of interesting findings result from the experiments.
+An expected one is that the amount of operations per second that can be performed with a serial implementation is almost inversely proportional to the amount of items to transform with a ratio of 1.
+On the other hand, this is not the case for the parallel approach.
 
-\paragraph{Remark 1.}
+An acceptable speedup of approximately \emph{2x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
+Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \verb+SharedArrayBuffer+s as they are likely not optimized, being that they are only an experimental feature in Firefox Nightly.
 
-In the printed volumes, illustrations are generally black and white
-(halftones), and only in exceptional cases, and if the author is
-prepared to cover the extra cost for color reproduction, are colored
-pictures accepted. Colored pictures are welcome in the electronic
-version free of charge. If you send colored figures that are to be
-printed in black and white, please make sure that they really are
-legible in black and white. Some colors as well as the contrast of
-converted colors show up very poorly when printed in black and white.
+\section{Future Work}
+There are still some parts of the library that would benefit from optimizations.
+Of particular interest to us is to perform further profiling the code executing in each WebWorker.
+As of today, using browser profiling tools for Web Workers is not as simple as doing so for the UI thread.\\
 
-\subsection{Formulas}
+We intend to perform some tests to evaluate the possibility of serializing functions via \verb+SharedArrayBuffer+s.
+We have not done this up to now due to time constraints, but we believe that a performance improvements could be achieved by serializing function strings once, creating an array for the bytes and then sharing the backing buffer with all workers.\\
 
-Displayed equations or formulas are centered and set on a separate
-line (with an extra line or halfline space above and below). Displayed
-expressions should be numbered for reference. The numbers should be
-consecutive within each section or within the contribution,
-with numbers enclosed in parentheses and set on the right margin --
-which is the default if you use the \emph{equation} environment, e.g.,
-\begin{equation}
-  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
-  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
-\end{equation}
+During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times.
+Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high performance degrades.
+We are plan to spend some time on this issue.\\
 
-Equations should be punctuated in the same way as ordinary
-text but with a small space before the end punctuation mark.
+There is a proposal to support typed objects in a future version of EcmaScript, which in turn would allow \verb+TypedArray+s of said typed objects.
+Once a browser has an experimental version available we intend to make sure that typed arrays of typed objects work with the library.
 
-\subsection{Footnotes}
-
-The superscript numeral used to refer to a footnote appears in the text
-either directly after the word to be discussed or -- in relation to a
-phrase or a sentence -- following the punctuation sign (comma,
-semicolon, or period). Footnotes should appear at the bottom of
-the
-normal text area, with a line of about 2~cm set
-immediately above them.\footnote{The footnote numeral is set flush left
-and the text follows with the usual word spacing.}
-
-\subsection{Program Code}
-
-Program listings or program commands in the text are normally set in
-typewriter font, e.g., CMTT10 or Courier.
-
-\medskip
-
-\noindent
-{\it Example of a Computer Program}
-\begin{verbatim}
-program Inflation (Output)
-  {Assuming annual inflation rates of 7%, 8%, and 10%,...
-   years};
-   const
-     MaxYears = 10;
-   var
-     Year: 0..MaxYears;
-     Factor1, Factor2, Factor3: Real;
-   begin
-     Year := 0;
-     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
-     WriteLn('Year  7% 8% 10%'); WriteLn;
-     repeat
-       Year := Year + 1;
-       Factor1 := Factor1 * 1.07;
-       Factor2 := Factor2 * 1.08;
-       Factor3 := Factor3 * 1.10;
-       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
-     until Year = MaxYears
-end.
-\end{verbatim}
-%
-\noindent
-{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
-report. Springer, New York)}
-
-\subsection{Citations}
-
-For citations in the text please use
-square brackets and consecutive numbers: \cite{jour}, \cite{lncschap},
-\cite{proceeding1} -- provided automatically
-by \LaTeX 's \verb|\cite| \dots\verb|\bibitem| mechanism.
-
-\subsection{Page Numbering and Running Heads}
-
-There is no need to include page numbers. If your paper title is too
-long to serve as a running head, it will be shortened. Your suggestion
-as to how to shorten it would be most welcome.
-
-\section{LNCS Online}
-
-The online version of the volume will be available in LNCS Online.
-Members of institutes subscribing to the Lecture Notes in Computer
-Science series have access to all the pdfs of all the online
-publications. Non-subscribers can only read as far as the abstracts. If
-they try to go beyond this point, they are automatically asked, whether
-they would like to order the pdf, and are given instructions as to how
-to do so.
-
-Please note that, if your email address is given in your paper,
-it will also be included in the meta data of the online version.
-
-\section{BibTeX Entries}
-
-The correct BibTeX entries for the Lecture Notes in Computer Science
-volumes can be found at the following Website shortly after the
-publication of the book:
-\url{http://www.informatik.uni-trier.de/~ley/db/journals/lncs.html}
-
-\subsubsection*{Acknowledgments.} The heading should be treated as a
-subsubsection heading and should not be assigned a number.
-
-\section{The References Section}\label{references}
-
-In order to permit cross referencing within LNCS-Online, and eventually
-between different publishers and their online databases, LNCS will,
-from now on, be standardizing the format of the references. This new
-feature will increase the visibility of publications and facilitate
-academic research considerably. Please base your references on the
-examples below. References that don't adhere to this style will be
-reformatted by Springer. You should therefore check your references
-thoroughly when you receive the final pdf of your paper.
-The reference section must be complete. You may not omit references.
-Instructions as to where to find a fuller version of the references are
-not permissible.
-
-We only accept references written using the latin alphabet. If the title
-of the book you are referring to is in Russian or Chinese, then please write
-(in Russian) or (in Chinese) at the end of the transcript or translation
-of the title.
-
-The following section shows a sample reference list with entries for
-journal articles \cite{jour}, an LNCS chapter \cite{lncschap}, a book
-\cite{book}, proceedings without editors \cite{proceeding1} and
-\cite{proceeding2}, as well as a URL \cite{url}.
-Please note that proceedings published in LNCS are not cited with their
-full titles, but with their acronyms!
+Once the \verb+SharedArrayBuffer+ type is available in other browsers (there is already a proposal for Google Chrome) we plan to perform benchmarks and tests to compare with Firefox Nightly.
 
@@
 
+\bibitem{es6-ta} TypedArray objects, ECMA, \url{https://people.mozilla.org/~jorendorff/es6-draft.html#sec-typedarray-objects}
+
+\bibitem{pjs-api} p-j-s API, \url{https://github.com/pjsteam/pjs/blob/dev/README.md}
+
 \bibitem{html-whatwg} HTML Standard, WHATWG, \url{https://html.spec.whatwg.org}
@@
 
-% end of our comments, start of sample comments
-
-\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
-Subsequences. J. Mol. Biol. 147, 195--197 (1981)
-
-\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
-Composing a Complex Biological Workflow through Web Services. In: Nagel,
-W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
-pp. 1148--1158. Springer, Heidelberg (2006)
-
-\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
-Infrastructure. Morgan Kaufmann, San Francisco (1999)
-
-\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
-Information Services for Distributed Resource Sharing. In: 10th IEEE
-International Symposium on High Performance Distributed Computing, pp.
-181--184. IEEE Press, New York (2001)
-
-\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
-Grid: an Open Grid Services Architecture for Distributed Systems
-Integration. Technical report, Global Grid Forum (2002)
-
-\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}
+\bibitem{benchmark} Sepia tone transformation benchmark comparison between p-j-s and plain JS, \url{http://jsperf.com/pjs-shared-vs-normal-vs-seq-100/4}
 
@@
 
-\section*{Appendix: Springer-Author Discount}
-
-LNCS authors are entitled to a 33.3\% discount off all Springer
-publications. Before placing an order, the author should send an email, 
-giving full details of his or her Springer publication,
-to \url{orders-HD-individuals@springer.com} to obtain a so-called token. This token is a
-number, which must be entered when placing an order via the Internet, in
-order to obtain the discount.
-
-\section{Checklist of Items to be Sent to Volume Editors}
-Here is a checklist of everything the volume editor requires from you:
-
-\begin{itemize}
-\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
-\itemsep8pt\relax
-\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}
-
-\item The final \LaTeX{} source files
-\item A final PDF file
-\item A copyright form, signed by one author on behalf of all of the
-authors of the paper.
-\item A readme giving the name and email address of the
-corresponding author.
-\end{itemize}
 \end{document}

### 2015-05-09 docs/paper/paper.tex
Finished first paper draft

@@
 
-This paper is organized as follows: Section~\ref{sec-background} provides some context on the current state of thread-base parallelization in JavaScript.
-Section~\ref{sec-overview} provides an overview of how the library works.
-Section~\ref{sec-ser-transf} explains the challenges the library faces, the alternative solutions and the decisions taken.
-Section~\ref{sec-bench} compares the performance of using the library against a sequential approach for an image transformation.
-Finally, Section~\ref{sec-future} presents future research directions.
+This paper is organized as follows: Section~\ref{sec:background} provides some context on the current state of thread-base parallelization in JavaScript.
+Section~\ref{sec:overview} provides an overview of how the library works.
+Section~\ref{sec:implementation} explains the challenges the library faces, the alternative solutions and the decisions taken.
+Section~\ref{sec:experiment} compares the performance of using the library against a sequential approach for an image transformation.
+Finally, Section~\ref{sec:future} presents future research directions.
 
-\section{Background}\label{sec-background}
+\section{Background}\label{sec:background}
 All modern versions of major browsers (Internet Explorer, Mozilla Firefox, Google Chrome, Safari and Opera) allow the execution of JavaScript code in a single threaded event loop runtime \cite{event-loop}.
@@
 
-\section{Library Overview}\label{sec-overview}
+\section{Library Overview}\label{sec:overview}
 JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs.
@@
 
-Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec-ser-transf}.
+Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:overview}.
 It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation.
@@
 }).seq().then(function(newValues){
-  // newValues is a new array [2,4,6,8]
+  // newValues is a new SharedUint8Array [2,4,6,8]
 });
@@
 The asynchronous nature of the operation is reflected by the fact that the call to \verb+seq+ returns a \verb+Promise+ (alternatively a callback could be passed as a parameter to it).
+
 It is possible to chain multiple operations before calling \verb+seq+ to avoid passing data back and forth between the Web Workers and the UI thread.
+Before \verb+seq+ is invoked, an operation is a chain of one or more steps.
+From the same main chain, multiple chains could be created:
+\medskip
 
-\section{Serialization and Transference}\label{sec-ser-transf}
+\noindent
+{\it p-j-s chaining sample}
+\begin{verbatim}
+var xs = new SharedUint8Array(4);
+xs.set([1,2,3,4]);
+
+var chain = pjs(xs).map(function(e){
+  return e * 3;
+});
+
+var evenChain = chain.filter(function(e){
+  return e % 2 === 0;
+});
+
+var sumChain = chain.reduce(function(c, v){
+  return c + v;
+}, 0, 0);
+
+evenChain.seq().then(function(evens){
+  // evens is [6, 12]
+  sumChain.seq().then(function(sum){
+    // sum is [3,6,9,12]
+  });
+});
+\end{verbatim}
+%
+\noindent
+{\small (Example creating multiple chains from the same base chain)}\\
+
+\section{Implementation}\label{sec:implementation}
 Let's consider an array of \verb+N+ elements on which a particular transformation is to be performed through the \verb+map+ function.
@@
   T_{sync} = T_{ft} + T_{et} + T_{fs} + T_{es} \;  .
-  \label{eq-tsync}
+  \label{eq:tsync}
 \end{equation}
@@
   T_{par} \approx \frac{T_{ser}}{K} + T_{sync} \;  .
-  \label{eq-tpar}
+  \label{eq:tpar}
 \end{equation}
 
-Based on Eq.~\ref{eq-tsync} and Eq.~\ref{eq-tpar} it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.
+Based on Eq.~\ref{eq:tsync} and Eq.~\ref{eq:tpar} it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.
 
@@
 
-\subsection{Functions}
+\subsection{Functions}\label{sub-sec:functions}
 The library must handle the distribution of the code between the different workers when an operation is executed.
@@
 
-Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered (Sec.~\ref{sec:context-sharing} proposes a solution to sharing additional data and functions between the UI thread and workers) a possible alternative is to:
+Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered (Sec.~\ref{sub-sec:context} proposes a solution to sharing additional data and functions between the UI thread and workers) a possible alternative is to:
 \begin{enumerate}
@@
 
-\section{Experiment}
-It is important to compare the performance of our parallel implementation against a serial one, as some processing is required before the computation begins and after it ends to distribute the data and work and gather the results respectively, as explained in Section~\ref{sec-ser-transf}.
+\subsection{Contexts}\label{sub-sec:context}
+Functions in JavaScript can capture variables as long as they are within the variable's lexical scope.
+Functions that make use  of this feature are closures.
+As explained in subsection~\ref{sub-sec:functions}, \verb+Function+ objects cannot be passed as messages to workers.
+This means that there is no out of the box mechanism to make variables that are available in the environment when the parallel computation is started accessible when executing the callback functions in each Web Worker.
+
+To solve this problem, p-j-s introduces the \emph{context} concept. There are two kinds of contexts:
+\begin{description}
+\item[Global context] Shared across all chains.
+\item[Local context] Specific to a particular step in a chain.
+\end{description}
+
+The following code snippet shows how developers can work with each type of context:
+\medskip
+
+\noindent
+{\it p-j-s context sample}
+\begin{verbatim}
+var xs = new SharedUint8Array(4);
+xs.set([1,2,3,4]);
+
+pjs.updateContext({
+  max: 3
+}).then(function(){
+  pjs(xs).filter(function(e, ctx){
+    return e <= ctx.max && e >= ctx.min;
+  }, {
+    min: 2
+  }).seq().then(function(range){
+    // range is [2,3]
+  });
+});
+
+\end{verbatim}
+%
+\noindent
+{\small (Example using local and global context for chains)}\\
+
+When executing each step in a chain, the local context is merged with the global context (neither is modified) into a separate object which is the \verb+ctx+ parameter that the function receives.
+If the same key is present in both the global and local context, the one in the local context is the one that will be available.
+Given the following code snippet, the available \verb+ctx+ for each function is displayed in Fig.~\ref{fig:context}.
+\medskip
+
+\noindent
+{\it p-j-s context merge sample}
+\begin{verbatim}
+var xs = new SharedUint8Array(4);
+xs.set([1,2,3,4]);
+
+pjs.updateContext({
+  A: 0,
+  B: 25,
+  C: 44
+}).then(function(){
+  var chain = pjs(xs)
+    .map(function(e, ctx){
+      // code here
+    }, { A: 5, B: 13 })
+    .filter(function(e, ctx){
+      // code here
+    }, { A: 2, E: 9 })
+});
+
+\end{verbatim}
+%
+\noindent
+{\small (Example for resulting ctx)}\\
+
+\begin{figure}
+\centering
+\includegraphics[width=1\textwidth]{context}
+\caption{Relative performance of parallel transformation compared to sequential one.}
+\label{fig:context}
+\end{figure}
+
+\section{Experiment}\label{sec:experiment}
+It is important to compare the performance of our parallel implementation against a serial one, as some processing is required before the computation begins and after it ends to distribute the data and work and gather the results respectively, as explained in Section~\ref{sec:implementation}.
 
@@
 \subsection{Results}
-The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark}.
+The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark} and Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
 \begin{table}
@@
 
-Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
-
 \begin{figure}
 \centering
-\includegraphics[height=6.2cm]{comparison}
+\includegraphics[height=8.2cm]{comparison}
 \caption{Relative performance of parallel transformation compared to sequential one.}
@@
 
-An acceptable speedup of approximately \emph{2x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
+An acceptable speedup of approximately \textbf{2x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
 Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \verb+SharedArrayBuffer+s as they are likely not optimized, being that they are only an experimental feature in Firefox Nightly.
 
-\section{Future Work}
+\section{Future Work}\label{sec:future}
 There are still some parts of the library that would benefit from optimizations.
-Of particular interest to us is to perform further profiling the code executing in each WebWorker.
+Of particular interest to us is to perform further profiling of the code executing in each Web Worker.
 As of today, using browser profiling tools for Web Workers is not as simple as doing so for the UI thread.\\
@@
 We intend to perform some tests to evaluate the possibility of serializing functions via \verb+SharedArrayBuffer+s.
-We have not done this up to now due to time constraints, but we believe that a performance improvements could be achieved by serializing function strings once, creating an array for the bytes and then sharing the backing buffer with all workers.\\
+We have not done this up to now due to time constraints, but we believe that performance improvements could be achieved by serializing function string once, creating an \verb+SharedTypedArray+ for the bytes and then sharing the backing buffer with all workers.\\
 
@@
 Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high performance degrades.
-We are plan to spend some time on this issue.\\
+We plan to spend some time on this issue.\\
 
-There is a proposal to support typed objects in a future version of EcmaScript, which in turn would allow \verb+TypedArray+s of said typed objects.
-Once a browser has an experimental version available we intend to make sure that typed arrays of typed objects work with the library.
+There is a strawman proposal to support typed objects \cite{typed-objects} in a future version of EcmaScript, which in turn would allow \verb+TypedArray+s of said typed objects.
+Once a browser has an experimental version available we intend to make sure that typed arrays of typed objects work with the library.\\
 
-Once the \verb+SharedArrayBuffer+ type is available in other browsers (there is already a proposal for Google Chrome) we plan to perform benchmarks and tests to compare with Firefox Nightly.
+Once the \verb+SharedArrayBuffer+ type is available in other browsers (there is already a proposal for Google Chrome \cite{sab-chrome}) we plan to perform benchmarks and tests to compare with Firefox Nightly.
 
@@
 
+\bibitem{typed-objects} Typed Objects, ECMA, \url{http://wiki.ecmascript.org/doku.php?id=harmony:typed_objects}
+
+\bibitem{sab-chrome} Shared Array Buffers, Chromium Dashboard, \url{https://www.chromestatus.com/feature/4570991992766464}
+
 \end{thebibliography}

### 2015-05-09 docs/paper/paper.tex
Some feedback from Rosita

@@
 % a short form should be given in case it is too long for the running head
-\titlerunning{Algorithmic skeletons to parallelizing computation in browsers}
+\titlerunning{Algorithmic skeletons to parallelize computation in browsers}
 
@@
 
-Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:overview}.
+Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:implementation}.
 It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation.
@@
 To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+.
-Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1000000 (a million).
+Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1,000,000 (a million).
 As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales.
@@
 \includegraphics[width=1\textwidth]{context}
-\caption{Relative performance of parallel transformation compared to sequential one.}
+\caption{Example of global and local context being merged for chain steps.}
 \label{fig:context}
@@
 In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform.
-That is because in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.
+The reason for this is that, in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.
 
@@
     100 & \textbf{154442} & 377.2\\
-    1000 & \textbf{17460.2} &  368.4 \\
-    10000 & \textbf{1768.6} & 311.2 \\
-    100000 & \textbf{179.2} & 172.2 \\
-    1000000 & 16.96 & \textbf{33.2} \\
+    1,000 & \textbf{17460.2} &  368.4 \\
+    10,000 & \textbf{1768.6} & 311.2 \\
+    100,000 & \textbf{179.2} & 172.2 \\
+    1,000,000 & 16.96 & \textbf{33.2} \\
     \hline
@@
 During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times.
-Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high performance degrades.
+Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high, performance degrades.
 We plan to spend some time on this issue.\\
@@
 
+\bibitem{promise} Promise, MDN, \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Global_Objects/Promise}
+
 \bibitem{html-whatwg} HTML Standard, WHATWG, \url{https://html.spec.whatwg.org}

### 2015-05-09 docs/paper/paper.tex
Added some things

@@
 \subsection{Results}
-The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark} and Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
+The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark}.
+Based on those results, we created a chart (Figure~\ref{fig:relative}) that shows the relative performance of the parallel approach compared to the sequential one.
 \begin{table}
@@
 Once the \verb+SharedArrayBuffer+ type is available in other browsers (there is already a proposal for Google Chrome \cite{sab-chrome}) we plan to perform benchmarks and tests to compare with Firefox Nightly.
+
+\subsubsection*{Acknowledgments.}
+Thanks to Rosa Wachenhauzer, our tutor, for reviewing this paper and also for all her support and guidance throughout this entire project.
 

### 2015-05-16 docs/paper/paper.tex
Improvements for ways of passing functions

@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
-%
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
-%
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass[runningheads,a4paper]{llncs}
@@
 The library must handle the distribution of the code between the different workers when an operation is executed.
-This is a challenging problem with a lot of possible alternatives \cite{cloud-haskell} \cite{distributed-scheme}.
+This is a challenging problem with many possible alternatives: Cloud Haskell \cite{cloud-haskell} for example, requires a customized version of the Glasgow Haskell Compiler (GHC) to serialize closures and their captured variables as a code pointer and an environment, operating under the assumption that the same code is executed in all nodes.
+A proposal for using Scheme for distributed programming \cite{distributed-scheme} uses a custom serializable data structure to represent procedures, and macros in order to capture a closure's captured variables and serialize their values.
+It also assumes that all nodes are executing the same program since serialized functions store pointers to parts of source code files.
+
+In our case, modifying the runtime is not a viable alternative since we want our library to be usable by anyone creating applications for major browser versions.
 
@@
 
-Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered (Sec.~\ref{sub-sec:context} proposes a solution to sharing additional data and functions between the UI thread and workers) a possible alternative is to:
+Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered a possible alternative is to:
 \begin{enumerate}
@@
 \end{enumerate}
+
+Sec.~\ref{sub-sec:context} proposes a solution to sharing additional data and functions between the UI thread and Web Workers as a replacement for variables captured by closures.
 
@@
 
+One alternative to hide the \emph{context} concept from application developers could be to introduce a \textit{build} step that detects the variables captured by p-j-s callbacks and transforms the code.
+By doing this, theyg wouldn't need to do anything out of the ordinary with regards to closures and their environment when working with p-j-s.
+However, this approach would require transforming the original code, and would be complex and time consuming to implement.
+
 \section{Experiment}\label{sec:experiment}

### 2015-05-16 docs/paper/paper.tex
Updated major comparison to firefox 41 nightly

@@
     Operating System & OS X Yosemite 10.10.3 \\
-    Browser version & Firefox Nightly 40.0a1 \\
+    Browser version & Firefox Nightly 41.0a1 \\
     \hline
@@
     \hline
-    100 & \textbf{154442} & 377.2\\
-    1,000 & \textbf{17460.2} &  368.4 \\
-    10,000 & \textbf{1768.6} & 311.2 \\
-    100,000 & \textbf{179.2} & 172.2 \\
-    1,000,000 & 16.96 & \textbf{33.2} \\
+    100 & \textbf{142223} & 328 \\
+    1,000 & \textbf{16503.8} &  359 \\
+    10,000 & \textbf{1683.8} & 321.8 \\
+    100,000 & \textbf{169.6} & 184.4 \\
+    1,000,000 & 16.8 & \textbf{34.1} \\
     \hline
@@
 \centering
-\includegraphics[height=8.2cm]{comparison}
+\includegraphics[height=7.7cm]{comparison}
 \caption{Relative performance of parallel transformation compared to sequential one.}

### 2015-05-16 docs/paper/paper.tex
Minor fix
Updated all current benchmarks to use FF 41

@@
 To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+.
-Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1,000,000 (a million).
+Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one\footnote{http://jsperf.com/transferrable-vs-cloning/2} has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one\footnote{http://jsperf.com/longer-transferrable-vs-cloning/2} 1,000,000 (a million).
 As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales.
@@
     \hline
-    \(1E5\) & 972 & \textbf{1705} \\
-    \(1E6\) & 114 & \textbf{754} \\
+    \(1E5\) & 734 & \textbf{1596} \\
+    \(1E6\) & 91 & \textbf{1134} \\
     \hline
@@
 
-The benchmarks provided the following results:
+The results for the benchmark\footnote{http://jsperf.com/pjs-serialization-comparison/2} are in Tab.~\ref{tab:encode-decode}:
 \begin{table}
@@
     \hline
-    \textbf{5247} & 4685 \\
+    \textbf{5017} & 4398 \\
     \hline
@@
 
-We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time (results in Tab.~\ref{tab:send-function}).
+We created benchmarks\footnote{http://jsperf.com/pjs-serialization-long/3 and http://jsperf.com/pjs-serialization/3} using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time (results in Tab.~\ref{tab:send-function}).
 We measured three approaches to transfer the string:
@@
     \hline
-    Short & 1251 & \textbf{7046} & 6219 \\
-    Long & 705 & 1176 & \textbf{1198} \\
+    Short & 1070 & \textbf{6629} & 5927 \\
+    Long & 1075 & \textbf{3273} & 2967 \\
     \hline
@@
 
-We also created a benchmark that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}.
+We also created a benchmark\footnote{http://jsperf.com/pjs-encoding/2} that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}.
 \begin{table}
@@
     \hline
-    4200 & \textbf{4340} \\
+    \textbf{4509} & 4270 \\
     \hline
@@
     10,000 & \textbf{1683.8} & 321.8 \\
-    100,000 & \textbf{169.6} & 184.4 \\
+    100,000 & 169.6 & \textbf{184.4} \\
     1,000,000 & 16.8 & \textbf{34.1} \\

### 2015-05-16 test/multiple_chaining.test.js
adds multiple chaining tests.

@@ -19,7 +19,7 @@ describe('chaining tests', function(){
     }
   });
 
-  it('should properly calculate the result without sibling chains\' interferience', function (done) {
+  it('should properly calculate the result without sibling chain operation interferience', function (done) {
     var xs = new Uint8Array(4);
     xs.set([1,2,3,4]);
 
@@ -39,6 +39,88 @@ describe('chaining tests', function(){
       expect(evens[0]).to.equal(6);
       expect(evens[1]).to.equal(12);
       done();
+    });
+  });
+
+  it('should properly calculate the result without sibling chain calculation interferience', function (done) {
+    var xs = new Uint8Array(4);
+    xs.set([1,2,3,4]);
+
+    var chain = pjs(xs).map(function(e){
+    return e * 3;
+    });
+
+    var evenChain = chain.filter(function(e){
+    return e % 2 === 0;
+    });
+
+    var sumChain = chain.reduce(function(c, v){
+    return c + v;
+    }, 0, 0);
+
+    evenChain.seq(function (err, evens) {
+      expect(evens[0]).to.equal(6);
+      expect(evens[1]).to.equal(12);
+      sumChain.seq(function (err, sum) {
+        expect(sum).to.equal(30);
+        done();
+      });
+    });
+  });
+
+  it('should properly calculate the result without previous chain calculation interferience', function (done) {
+    var xs = new Uint8Array(4);
+    xs.set([1,2,3,4]);
+
+    var chain = pjs(xs).map(function(e){
+    return e * 3;
+    });
+
+    var evenChain = chain.filter(function(e){
+    return e % 2 === 0;
+    });
+
+    var sumChain = chain.reduce(function(c, v){
+    return c + v;
+    }, 0, 0);
+
+    chain.seq(function (err, results) {
+      expect(results[0]).to.equal(3);
+      expect(results[1]).to.equal(6);
+      expect(results[2]).to.equal(9);
+      expect(results[3]).to.equal(12);
+      sumChain.seq(function (err, sum) {
+        expect(sum).to.equal(30);
+        done();
+      });
+    });
+  });
+
+  it('should properly calculate the result without next chain calculation interferience', function (done) {
+    var xs = new Uint8Array(4);
+    xs.set([1,2,3,4]);
+
+    var chain = pjs(xs).map(function(e){
+    return e * 3;
+    });
+
+    var evenChain = chain.filter(function(e){
+    return e % 2 === 0;
+    });
+
[…]

### 2015-05-16 docs/paper/paper.tex

@@
 The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
-However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing the notion of shared memory through \verb+SharedArrayBuffer+s \cite{sab}.
+However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing shared memory through \verb+SharedArrayBuffer+s \cite{sab}.
 In essence, the proposal allows the same memory to be shared across multiple Web Workers, and also aims to provide the necessary atomic/lock constructs to deal with shared memory.
@@
 On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.
-
 For that reason our library only works with \verb+SharedTypedArray+s and \verb+TypedArrays+.
@@
 \begin{enumerate}
-  \item Encoding the \verb+String+ into an \verb+TypedArrayBuffer+, sending it as a \verb+Tranferable+ and decoding it on the Web Worker.
+  \item Encoding the \verb+String+ into an \verb+TypedArray+, sending it as a \verb+Tranferable+ and decoding it on the Web Worker.
   \item Sending the \verb+String+ directly.
@@
 
-We also created a benchmark\footnote{http://jsperf.com/pjs-encoding/2} that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}.
+We created a separate benchmark\footnote{http://jsperf.com/shared-function-transfer} to compare the best option (copying) against encoding the function into a \verb+SharedTypedArray+ and transferring it to four workers.
+We expected the latter to be slower since it involves encoding the code once and decoding it in each worker.
+The results are shown in Tab.~\ref{tab:copy-vs-shared}.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|}
+    \hline
+    Copying [ops/sec] & SharedArrayBuffer [ops/sec] \\
+    \hline
+    \textbf{1315} & 1226 \\
+    \hline
+  \end{tabular}
+  \caption{Sending code to four workers by copying compared to encoding it and using a SharedArrayBuffer.}
+  \label{tab:copy-vs-shared}
+\end{table}
+
+We also created a benchmark\footnote{http://jsperf.com/pjs-encoding/2} that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}).
 \begin{table}
@@
 
-We intend to perform some tests to evaluate the possibility of serializing functions via \verb+SharedArrayBuffer+s.
-We have not done this up to now due to time constraints, but we believe that performance improvements could be achieved by serializing function string once, creating an \verb+SharedTypedArray+ for the bytes and then sharing the backing buffer with all workers.\\
-
 During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times.

### 2015-05-19 docs/paper/paper.tex
Minor fix
Paper fixes

@@
 In the year 2010 it was predicted that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation.
-This is the cause why newer computer have a greater number of processors instead of processors with more computing power.
+This is the reason why newer computers have a greater number of processors instead of processors with more computing power.
 
 To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors.
-In the case of application executed on a single device (e.g.
+In the case of applications executed on a single device (e.g.
 browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing.
@@
 
-On a different note, one of the tendencies that is beggining to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications that were previously only considered viable in a native environment, such as video games.
+On a different note, one of the tendencies that is beginning to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications, such as video games, that were previously only considered viable in a native environment.
 This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
@@
 While such a runtime is a good fit for most web applications, JavaScript and the browser are now also being used for some CPU intensive tasks, such as games, which commonly require physics calculations, image manipulation (in 2D) and graphics generation (in 3D).
-Despite the large improvement in JavaScript performance obtained through optimizing compilers such as Google Chrome's V8 \cite{v8} and Mozilla Firefox's SpiderMonkey \cite{spider-monkey}, it is important to be able to take advantage of multiple cores in modern devices to achieve even greater performance.\\
+Despite the large improvement in JavaScript's performance, obtained through optimizing compilers such as Google Chrome's V8 \cite{v8} and Mozilla Firefox's SpiderMonkey \cite{spider-monkey}, it is important to be able to take advantage of multiple cores in modern devices to achieve even greater performance.\\
 In 2010, Web Workers were made part of the web standard.
-Web Workers allow the creation of ``thread like'' constructs in a browser environment, but they don't allow on shared memory and instead communicate via message passing.
+Web Workers allow the creation of ``thread like'' constructs in a browser environment, but they don't allow shared memory and instead communicate via message passing.
 The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
@@
 In essence, the proposal allows the same memory to be shared across multiple Web Workers, and also aims to provide the necessary atomic/lock constructs to deal with shared memory.
-In the particular case of embarrassingly parallel computations (such as map, filter and reduce operations on an array), it is simple to take advantage of the shared memory speed benefits without incurring in any overhead due to locks.
+In the particular case of embarrassingly parallel computations (such as map, filter and reduce operations on an array), it is simple to take advantage of the speed benefits of shared memory without incurring in any overhead due to synchronization.
 
@@
 Considering the definition of a \verb+Transferable+, it seems like a good alternative to minimize both \(T_{et}\) and \(T_{es}\).
-Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.6} of that same standard).
+Even more so when one considers that otherwise objects are copied using structured cloning (explained in section \textbf{2.7.6} of that same standard).
 
@@
 One alternative to hide the \emph{context} concept from application developers could be to introduce a \textit{build} step that detects the variables captured by p-j-s callbacks and transforms the code.
-By doing this, theyg wouldn't need to do anything out of the ordinary with regards to closures and their environment when working with p-j-s.
+By doing this, they wouldn't need to do anything out of the ordinary with regards to closures and their environment when working with p-j-s.
 However, this approach would require transforming the original code, and would be complex and time consuming to implement.

### 2015-05-23 test/queueing_workers_jobs.test.js
Refactors queueing tests and adds error on queued operation test.

@@ -22,33 +22,51 @@ describe('queueing workers jobs', function(){
         var sourceArrayA = new TypedArray([1,2,3,4,5,6,7,8,9,10,11,12,13,14]);
         var sourceArrayB = new TypedArray([20,21,22,23,24,25,26,27,28,29,30,31]);
 
-        it('should queue consecutive map operations', function(done){
+        it('should respect chains order', function () {
           var mapperA = function(e) { return e * 2; };
-          var mapperB = function(e) { return e * 4; }
+          var mapperB = function(e) { return e * 4; };
           var chainA = pjs(sourceArrayA).map(mapperA);
           var chainB = pjs(sourceArrayB).map(mapperB);
 
-          chainA.seq(function(err, result) {
-            if (err) { return done(err); }
-            finisher(err, result, sourceArrayA, mapperA);
-          });
-          chainB.seq(function(err, result) {
-            if (err) { return done(err); }
-            finisher(err, result, sourceArrayB, mapperB);
-          });
+          var promiseA = chainA.seq();
+          var promiseB = chainB.seq();
 
-          var finishCount = 0;
-          var finisher = function (err, result, sourceArray, mapper) {
-            finishCount++;
-            expect(result).to.have.length(sourceArray.length);
-            expect(utils.getTypedArrayType(result)).to.equal(utils.getTypedArrayType(sourceArray));
-            for (var i = sourceArray.length - 1; i >= 0; i--) {
-              expect(result[i]).to.equal(mapper(sourceArray[i]));
+          return Promise.race([promiseA, promiseB]).then(function (result) {
+            expect(result).to.have.length(sourceArrayA.length);
+            expect(utils.getTypedArrayType(result)).to.equal(utils.getTypedArrayType(sourceArrayA));
+            for (var i = sourceArrayA.length - 1; i >= 0; i--) {
+              expect(result[i]).to.equal(mapperA(sourceArrayA[i]));
             }
-            if (finishCount === 2) {
-              done();
-            }
-          }
+            return promiseB.then(function (result) {
+              expect(result).to.have.length(sourceArrayB.length);
+              expect(utils.getTypedArrayType(result)).to.equal(utils.getTypedArrayType(sourceArrayB));
+              for (var i = sourceArrayB.length - 1; i >= 0; i--) {
+                expect(result[i]).to.equal(mapperB(sourceArrayB[i]));
+              }
+            });
+          });
+        });
+
[…]

### 2015-05-23 docs/report/intro.tex
Corrections up to and including initialization.tex

--- 2015-02-08 docs/report/intro.tex
@@
 
-In the year 2010 it was predicted \cite{itrs} that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation.
-This is the cause why newer computer have a greater number of processors instead of processors with more computing power.
+In the year 2010 it was predicted that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation.
+This is the reason why newer computers have a greater number of processors instead of processors with more computing power.
 
 To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors.
-In the case of application executed on a single device (e.g.
+In the case of applications executed on a single device (e.g.
 browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing.
@@
 
-On a different note, one of the tendencies that is beggining to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications that were previously only considered viable in a native environment, such as video games.
+On a different note, one of the tendencies that is beginning to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications, such as video games, that were previously only considered viable in a native environment.
 This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
 In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications.
-Initiatives such as ParallelJS \cite{par-js} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.
+Initiatives such as ParallelJS \cite{par-js} and River trail \cite{rivertrail} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.
 
-The goal of this professional assignment is to provide an alternative for parallel code eection in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce.
+The goal of this paper is to provide an alternative for parallel code execution in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce.
 This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.

### 2015-05-23 docs/report/micro_benchmarks.tex
Corrections up to and including initialization.tex

@@
 A common practice to measure the performance of a fragment of code, or to compare it to the performance of some other code fragment is to create a micro benchmark.
-Throughout this document, there are multiple references to benchmarks of this kind since their results helped us make decisions about how to best implement some specific library features.
+Throughout this document, there are multiple references to benchmarks of this kind, since their results helped us make decisions about how to best implement some specific library features.
 
@@
 The first point is applicable to all programming languages but the second one is only important in engines/VMs that perform such optimizations.
-Since this document discussed the performance of JavaScript and in particular v8, the second point does affect our benchmarks.
+This document discusses the performance of JavaScript and in particular v8, so the second point does affect our benchmarks.
 
 \subsection{A small example\protect\footnote{The example is based on \cite{mraleph-bc}.}}
-Let's assume that we want to profile to the following function:
+Let's assume that we want to profile the following function:
 \begin{lstlisting}[caption=Function to benchmark]
@@
 
-The problem with the above is that it does not consider optimizations. Let's assume that the executing engine is smart enough to:
+The problem with the above benchmark is that it does not consider optimizations. Let's assume that the executing engine is smart enough to:
 \begin{enumerate}
@@
 
-If it applies \#2 it realizes it does not need to perform the comparison in every iteration so it could extract it outside the loop, resulting in
+When it applies \#2, it realizes it does not need to perform the comparison in every iteration so it could extract it outside the loop, resulting in
 \begin{lstlisting}[caption=Naive benchmark after evaluating loop invariants]
@@
 To verify that is not the case the tool IRHydra\cite{ir-hydra} can be used , which \textit{can display intermediate representations used by V8}.
-As shown in the following figure, v8 inline the function (the blue chevron marks that) but the loop invariant was not detected and optimized:
+As shown in the following figure, v8 inlined the function (the blue chevron marks that) but the loop invariant was not detected and optimized:
 \begin{figure}[h!]
@@
 
-\note{All benchmarks are performed using two different versions of Google Chrome (40.0.2214 and 41.0.2272) to be able to get an idea of the evolution of each feature.}
+\subsection{Details}
+\label{sub-sec:serialization-and-transference}
+The benchmarks are executed using two different machines:
+
+\textbf{Machine 1}
+\begin{itemize}
+  \item Mac Book Pro Retina
+  \item Intel Core i7 2.8GHz
+  \item 16 GB RAM DDR3 1600 MHz
+  \item SSD 250 GB
+  \item Chrome 41.0.2272
+\end{itemize}
+
+\textbf{Machine 2}
+\begin{itemize}
+  \item Mac Book Air
+  \item Intel Core i5 1.8GHz
+  \item 4 GB RAM DDR3 1600 MHz
+  \item SSD 128 GB
+  \item Chrome 43.0.2357
+\end{itemize}
 

### 2015-05-23 src/workers.js
Fixes queueing.

@@ -2,7 +2,7 @@ var ResultCollector = require('./result_collector');
 var work = require('webworkify');
 
 var workers = [];
-var packQueue = [];
+var packQueue;
 var sentPack;
 
 Object.defineProperty(module.exports, 'length', {
@@ -16,6 +16,8 @@ module.exports.init = function(workersCount){
     var worker = work(require('./worker.js'));
     workers.push(worker);
   }
+  packQueue = [];
+  sentPack = undefined;
 };
 
 module.exports.terminate = function(){
@@ -24,6 +26,8 @@ module.exports.terminate = function(){
   });
 
   workers = [];
+  packQueue = undefined;
+  sentPack = undefined;
 };
 
 module.exports.sendPacks = function(packs, callback){

### 2015-05-23 docs/report/array_sharing.tex
Corrections up to code sharing

@@
 
-As explained in section~\ref{sec:serialization-and-transference} of this document transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance.
-Since our library will receive \ttarray{} objects to be partiotioned and handled between multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.
+As explained in section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance.
+Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned}
 
 If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s.
-For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but since there's no shared memory and element ownership is tranferred to the Web Workers the elements must be copied to the sub-arrays.
+For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but \ttarray{}s do not support shared memory and element ownership is tranferred to the Web Workers, the elements must be copied to the sub-arrays.
 
@@
 \subsubsection{Conclusion}
-The benchmark shows that there are three very similar alternatives to split a \ttarray{}.
+The benchmark shows that there are three very similar alternatives to partition a \ttarray{}.
 We are going to use \textit{Buffer subarray} as it yields better performance in a newer version of Chrome.
 The implementation of the slice function could look something like this:
-\begin{lstlisting}[caption=Function to benchmark]
+\begin{lstlisting}[caption=Possible approach to copy a slice of a \ttarray{}]
 function typedArraySlice(array, from, to) {

### 2015-05-23 docs/report/serialization_transference.tex
Corrections up to code sharing

@@
 Let's assume that we have an array of \code{N} elements on which we want to perform a particular transformation through the \code{map} function.
-If we run in a single thread (no parallelism) and the average time to process is element is \code{t} then we could approximate the total time \(T_{ser}\) for the operation as:
+If code is executed in a single thread (no parallelism) and the average time to process an element is \(t\) then we could approximate the total time \(T_{ser}\) for the operation as:
 \[T_{ser} = \sum_{i = 0}^{N}t = Nt\]
@@
 
-Based on the previous equations we deduce that we can reduce \(T_{sync}\) the closer to the best possible scenario we will be.
+Based on the previous equations we deduce that: the more we can reduce \(T_{sync}\), the closer to the best possible scenario we will be.
 
 In our case we are trying to transfer objects between a browser's JavaScript UI thread and Web Workers so we are constrained by the means of that environment.
-The Worker interface is the following \cite{w3c-ww}:
+The Worker interface is the following one \cite{w3c-ww}:
 \begin{lstlisting}[language=Java, caption=The Worker interface]
@@
 
-As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects.
+As the aforementioned interface states, one can either send just a message or send a message with a sequence of transferable objects.
 From section \textbf{2.7.5} of the HTML Standard \cite{html-whatwg}:
 \textit{``Some objects support being copied and closed in one operation.
-This is called transferring the object, and is used in particular to transfer ownership of unsharable or expensive resources across worker boundaries."}
+This is called transferring the object, and is used in particular to transfer ownership of unsharable\footnote{Note that the definition does not consider \tsabuffer{} instances as those are sharable and transferable.} or expensive resources across worker boundaries."}
 
 Considering the definition of a \code{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\).
-Even more so when one considers that otherwise objects are copied using structured cloning(explained in section \textbf{2.7.6} of that same standard).
+Even more so when one considers that otherwise objects are copied using structured cloning (explained in section \textbf{2.7.6} of that same standard).
 
@@
 To verify the above we put together two benchmarks that aim to verify the difference between invoking \code{postMessage} with and without structured cloning for a \ttarray{}.
-Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \ttarray{} and the other one 1000000 (a million).
+Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the \ttarray{} and the other one 1,000,000 (a million).
 
-\subsubsection{With 100000 elements}
+\subsubsection{With 100,000 elements}
 The following is the code for the benchmark:
@@
 The results for it are displayed in Figure~\ref{fig:transferrable-vs-cloning}.
-\img{Results with 100000 elements}{../../spikes/1.2/transferrableVsCloning}{fig:transferrable-vs-cloning}
+\img{Results with 100,000 elements}{../../spikes/1.2/transferrableVsCloning}{fig:transferrable-vs-cloning}
 
-\subsubsection{With 1000000 elements}
+\subsubsection{With 1,000,000 elements}
 The following is the code for the benchmark:
@@
 The results for it are displayed in Figure~\ref{fig:long-transferrable-vs-cloning}.
-\img{Results with 100000 elements}{../../spikes/1.2/longerTransferrableVsCloning}{fig:long-transferrable-vs-cloning}
+\img{Results with 100,000 elements}{../../spikes/1.2/longerTransferrableVsCloning}{fig:long-transferrable-vs-cloning}
 
 \subsubsection{Conclusion}
-As we can see from the previous results increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \code{Transferable} objects, it scales.
+As we can see from the previous results, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \code{Transferable} objects; it scales.
 On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.

### 2015-05-23 dist/p-j-s-standalone.js
1.0.0-beta

@@ -1105,7 +1105,7 @@ var ResultCollector = require('./result_collector');
 var work = require('webworkify');
 
 var workers = [];
-var packQueue = [];
+var packQueue;
 var sentPack;
 
 Object.defineProperty(module.exports, 'length', {
@@ -1119,6 +1119,8 @@ module.exports.init = function(workersCount){
     var worker = work(require('./worker.js'));
     workers.push(worker);
   }
+  packQueue = [];
+  sentPack = undefined;
 };
 
 module.exports.terminate = function(){
@@ -1127,6 +1129,8 @@ module.exports.terminate = function(){
   });
 
   workers = [];
+  packQueue = undefined;
+  sentPack = undefined;
 };
 
 module.exports.sendPacks = function(packs, callback){

### 2015-05-23 dist/p-j-s.js

@@ -1028,7 +1028,7 @@ var ResultCollector = require('./result_collector');
 var work = require('webworkify');
 
 var workers = [];
-var packQueue = [];
+var packQueue;
 var sentPack;
 
 Object.defineProperty(module.exports, 'length', {
@@ -1042,6 +1042,8 @@ module.exports.init = function(workersCount){
     var worker = work(require('./worker.js'));
     workers.push(worker);
   }
+  packQueue = [];
+  sentPack = undefined;
 };
 
 module.exports.terminate = function(){
@@ -1050,6 +1052,8 @@ module.exports.terminate = function(){
   });
 
   workers = [];
+  packQueue = undefined;
+  sentPack = undefined;
 };
 
 module.exports.sendPacks = function(packs, callback){

### 2015-05-23 docs/report/array_merging.tex
Finished initial corrections pass

@@
 \subsection{\ttarray{} merge benchmark}
-Unlike the case for \ttarray{} splitting (\ref{subsec:copying-typed-arrays}) there aren't that many native alternatives to create a \ttarray{} from smaller ones.
+Unlike the case for \ttarray{} splitting (Section~\ref{subsec:copying-typed-arrays}) there aren't that many native alternatives to create a \ttarray{} from smaller ones.
 
@@
 
-For illustrative purposes, a naive implementation (without optimizations or error checking) would be similar to the following one:
+For illustration purposes, a naive implementation (without optimizations or error checking) would be similar to the following one:
 \begin{lstlisting}[caption=Simple \ttarray{} merge function]

### 2015-05-23 docs/report/map_serial_vs_parallel.tex

@@
 
-Since as explained in \ref{sec:serialization-and-transference} the cost of transferring objects is not zero, it is important to compare the performance of our parallel implementation against a serial one.
+Since as explained in Section~\ref{sec:serialization-and-transference} the cost of transferring objects is not zero, it is important to compare the performance of our parallel implementation against a serial one.
 
@@
 That is because in the parallel case, the source \ttarray{} must be copied to sub-arrays in order to be transferred.
-On the contratry, the serial implementation can take that time to actually operate on some elements.
+On the contrary, the serial implementation operates on elements during that time.
 
@@
 
-\note{Unlike other benchmarks in which we were comparing different ways to achieve the same, this benchmark compares the same thing with different element amounts.
-That's why the original charts are not displayed.}
+\note{Unlike other benchmarks in which we were comparing different ways to achieve the same result, this benchmark compares a single way with a changing amount of elements.
+For that reason, the original charts are not displayed.}
 

### 2015-05-23 docs/report/modules.tex

@@
 
-\note{At the time of writing, Google Chrome does not support the ES6 module specification so that is not a viable choice.}
+\note{At the time of writing, neither Google Chrome nor Mozilla Firefox support the ES6 module specification so that is not a viable choice.}
 
@@
 
-Instead of making the choice for users, our library will support both systems in a way that is easy for users.
+Instead of making the choice for users, our library will support both systems in a way that makes it easier for users to consume either of them.
 
@@
 \begin{lstlisting}[language=HTML, caption=CommonJS usage]
-<script src="https://rawgit.com/pjsteam/pjs/v0.2.1/dist/p-j-s.min.js"></script>
+<script src="https://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 
@@
 \begin{lstlisting}[caption=main.js AMD support]
-require(["https://rawgit.com/pjsteam/pjs/v0.2.1/dist/p-j-s-standalone.min.js"], function(pjs) {
+require(["https://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s-standalone.min.js"], function(pjs) {
   pjs.init();

### 2015-05-23 spikes/2.2.1/test-howToSendCode/longSerialization.js
updates tests and adds cloning comparison result images.

@@ -1,4 +1,4 @@
-// http://jsperf.com/pjs-serialization-long/2
+// http://jsperf.com/pjs-serialization-long/3
 
 // HTML setup
 <script>

### 2015-05-23 spikes/6.1/cloningComparison/jspBenchmark.js

@@ -1,4 +1,4 @@
-// http://jsperf.com/additional-clonningdata-comparison/2
+// http://jsperf.com/additional-clonningdata-comparison/3
 
 // HTML setup
 <script>

### 2015-05-23 docs/paper/paper.tex

@@
 Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one\footnote{http://jsperf.com/transferrable-vs-cloning/2} has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one\footnote{http://jsperf.com/longer-transferrable-vs-cloning/2} 1,000,000 (a million).
-As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales.
+As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects; it scales.
 On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.

### 2015-05-23 docs/report/background.tex

--- 2015-05-23 docs/paper/paper.tex
@@
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-
-\usepackage{url}
-\urldef{\mailsa}\path|{dschenkelman, mservetto}@fi.uba.ar|
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
-
-\begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
-\title{Using algorithmic skeletons in EcmaScript to parallelize computation in browsers through Web Workers}
-
-% a short form should be given in case it is too long for the running head
-\titlerunning{Algorithmic skeletons to parallelize computation in browsers}
-
-% the name(s) of the author(s) follow(s) next
-%
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
-%
-\author{Damian Schenkelman\and Matias Servetto}
-%
-\authorrunning{Algorithmic skeletons to parallelizing computation in browsers}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-\institute{School of Engineering. University of Buenos Aires.\\
-Paseo Colon 850, Buenos Aires, Argentina\\
-\mailsa\\
-\url{http://www.fi.uba.ar/}}
-
-%
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
-%
-
-\toctitle{Lecture Notes in Computer Science}
-\tocauthor{Authors' Instructions}
-\maketitle
-
-\begin{abstract}
-With the growing tendency of browsers being used for CPU intensive applications (such as 3D games) it is important to take advantage of multiple cores to reduce user perceived latency.
-This paper presents a JavaScript library based on algorithmic skeletons that allows users to take advantage of parallel processing in browsers while maintaining a familiar coding style.
-\keywords{browsers, ecmascript, parallelization, algorithmic skeletons}
-\end{abstract}
-
-\section{Introduction}
-
-In 1965 Gordon Moore proposed the popularly known ``Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
-As a consequence, programs that use a single processor/core will automatically become faster without the need to modify them with that purpose in mind.
-
-In the year 2010 it was predicted that this tendency would be slowly reaching its end, mostly due to reasons related to heat dissipation.
-This is the reason why newer computers have a greater number of processors instead of processors with more computing power.
-
-To be able to maximize the performance of these computers, it is paramount that systems professionals can create programs that process data in parallel, through different mechanisms, such as threads or processes executing simultaneously in different processors.
-In the case of applications executed on a single device (e.g.
-browsers and web pages, applications for mobile devices, desktop apps, server applications) a lot of languages and platforms provide a simple way for programmers to abstract the complexity and coordination required for this type of processing.
-For example, the .NET platform has the Task Parallel Library (TPL) and Parallel LINQ (PLINQ) \cite{ms-par} libraries that use algorithmic skeletons modelled with high order functions so developers can perform complex operations through a simple API.
-
-On a different note, one of the tendencies that is beginning to take off is the use of EcmaScript \cite{es-web} (also known as Javascript) to create applications, such as video games, that were previously only considered viable in a native environment.
-This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
-In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications.
-Initiatives such as ParallelJS \cite{par-js} and River trail \cite{rivertrail} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.
-
-The goal of this paper is to provide an alternative for parallel code execution in EcmaScript, through a library that exposes high order functions to model algorithmic skeletons such as map, filter and reduce.
-This library will take advantage of Web Workers \cite{w3c-ww} (the mechanism proposed by browsers for parallelism at the time of writing) to allow the simultaneous execution of code in multiple processors.
-
-This paper is organized as follows: Section~\ref{sec:background} provides some context on the current state of thread-base parallelization in JavaScript.
-Section~\ref{sec:overview} provides an overview of how the library works.
-Section~\ref{sec:implementation} explains the challenges the library faces, the alternative solutions and the decisions taken.
-Section~\ref{sec:experiment} compares the performance of using the library against a sequential approach for an image transformation.
-Finally, Section~\ref{sec:future} presents future research directions.
-
 \section{Background}\label{sec:background}
@@
 The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
-However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing shared memory through \verb+SharedArrayBuffer+s \cite{sab}.
+However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing shared memory through \tsabuffer{}s \cite{sab}.
 In essence, the proposal allows the same memory to be shared across multiple Web Workers, and also aims to provide the necessary atomic/lock constructs to deal with shared memory.
@@
 
-\section{Library Overview}\label{sec:overview}
-JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs.
-In JavaScript, \verb+Array+s have \verb+map+, \verb+filter+, and \verb+reduce+ functions and version 6 of the EcmaScript adds those same methods to \verb+TypedArray+s \cite{es6-ta}.
-For example, to multiply all the elements of an \verb+Array+ by two:
-\medskip
+We developed our library to work with instances of the already standarized \tabuffer{} type using the Google Chrome broswer.
+Additionally, we support the experimental \tsabuffer{}, which at the time of writing is only available in Firefox Nightly\footnote{Current version is 41.0a1}.
 
-\noindent
-{\it plain JS map sample}
-\begin{verbatim}
-var newValues = [1,2,3,4].map(function(e){
-  return e * 2;
-});
-
-// newValues is a new array [2,4,6,8]
-\end{verbatim}
-%
-\noindent
-{\small (Example creating a new array with the values of the original one multiplied by two)}\\
-
-Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:implementation}.
-It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation.
-Given a \verb+SharedTypedArray+, the following code would do the same as the previous example but parallelizing the work:
-\medskip
-
-\noindent
-{\it p-j-s map sample}
-\begin{verbatim}
-var pjs = require('p-j-s');
-pjs.init({ maxWorkers: 4 });
-var xs = new SharedUint8Array(4);
-xs.set([1,2,3,4]);
-pjs(xs).map(function(e){
-  return e * 2;
-}).seq().then(function(newValues){
-  // newValues is a new SharedUint8Array [2,4,6,8]
-});
-\end{verbatim}
-%
-\noindent
-{\small (Example creating a new array with the values of the original one multiplied by two using p-j-s)}\\
-
-The call to \verb+pjs.init+ initializes the library and the amount of workers to be used for processing (a worker pool is created with them).
-In browsers that have an API to detect the amount of available cores in a machine, the \verb+maxWorkers+ value is limited by the amount of cores.
-
-The API provided by p-j-s \cite{pjs-api} is very similar to the synchronous one that is provided by the language.
-The asynchronous nature of the operation is reflected by the fact that the call to \verb+seq+ returns a \verb+Promise+ (alternatively a callback could be passed as a parameter to it).
-
-It is possible to chain multiple operations before calling \verb+seq+ to avoid passing data back and forth between the Web Workers and the UI thread.
-Before \verb+seq+ is invoked, an operation is a chain of one or more steps.
-From the same main chain, multiple chains could be created:
-\medskip
-
-\noindent
-{\it p-j-s chaining sample}
-\begin{verbatim}
-var xs = new SharedUint8Array(4);
-xs.set([1,2,3,4]);
-
-var chain = pjs(xs).map(function(e){
-  return e * 3;
-});
-
-var evenChain = chain.filter(function(e){
-  return e % 2 === 0;
-});
-
-var sumChain = chain.reduce(function(c, v){
-  return c + v;
-}, 0, 0);
-
-evenChain.seq().then(function(evens){
-  // evens is [6, 12]
-  sumChain.seq().then(function(sum){
-    // sum is [3,6,9,12]
-  });
-});
-\end{verbatim}
-%
-\noindent
-{\small (Example creating multiple chains from the same base chain)}\\
-
-\section{Implementation}\label{sec:implementation}
-Let's consider an array of \verb+N+ elements on which a particular transformation is to be performed through the \verb+map+ function.
-If the transformation is executed using a single thread (no parallelism) and the average time to process each element is \verb+t+ then the total time (\(T_{ser}\)) for the operation can be approximated as:
-
-\begin{equation}
-  T_{ser} = \sum_{i = 0}^{N}t = Nt \;  .
-\end{equation}
-
-When trying to parallelize this operation using \(K\) threads the ideal goal is to reach a total time (\(T_{par}\)) that is:
-
-\begin{equation}
-  T_{par} = \frac{T_{ser}}{K} \;  .
-\end{equation}
-
-Nevertheless, there are additional time consuming tasks other than the main computation that need to be considered when performing the operation in parallel in programs where not all memory is shared.
-These are:
-\begin{itemize}
-  \item Serializing/deserializing the elements to transfer.
-  \item Transferring the elements back and forth between the UI thread and the workers.
-  \item Serializing/deserializing the function to transfer.
-  \item Transferring to each worker the functions for the transformation.
-\end{itemize}
-
-If we drill down into the different parts:
-\begin{itemize}
-  \item \(T_{ft}\) as the function transfer time
-  \item \(T_{et}\) as the elements transfer time
-  \item \(T_{fs}\) as the function serialization/deserialization time
-  \item \(T_{es}\) as the elements serialization/deserialization time
-\end{itemize}
-
-It is clear that:
-\begin{equation}
-  T_{sync} = T_{ft} + T_{et} + T_{fs} + T_{es} \;  .
-  \label{eq:tsync}
-\end{equation}
-
-\begin{equation}
-  T_{par} \approx \frac{T_{ser}}{K} + T_{sync} \;  .
-  \label{eq:tpar}
-\end{equation}
-
-Based on Eq.~\ref{eq:tsync} and Eq.~\ref{eq:tpar} it can be deduced that the more \(T_{sync}\) can be reduced, the closer to the ideal scenario the computation will be.
-
-In our case we are trying to transfer objects between a browser's JavaScript UI thread and Web Workers so we are constrained by the means of that environment.
-The Worker interface is the following \cite{w3c-ww}:
-\medskip
-
-\noindent
-{\it Web Worker interface}
-\begin{verbatim}
-[Constructor(DOMString scriptURL)]
-interface Worker : EventTarget {
-  void terminate();
-
-  void postMessage(any message, optional sequence<Transferable> transfer);
-  [TreatNonCallableAsNull] attribute Function? onmessage;
-};
-Worker implements AbstractWorker;
-\end{verbatim}
-%
-\noindent
-{\small (Example extracted from the W3C Web Workers specification)}
-
-As the aforementioned interface states one can either send just a message or send a message with a sequence of transferable objects.
-From section \textbf{2.7.5} of the HTML Standard \cite{html-whatwg}:
-\textit{``Some objects support being copied and closed in one operation.
-This is called transferring the object, and is used in particular to transfer ownership of unsharable or expensive resources across worker boundaries."}
-
-\subsection{Elements}
-Considering the definition of a \verb+Transferable+, it seems like a good alternative to minimize both \(T_{et}\) and \(T_{es}\).
-Even more so when one considers that otherwise objects are copied using structured cloning (explained in section \textbf{2.7.6} of that same standard).
-
-To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+.
-Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one\footnote{http://jsperf.com/transferrable-vs-cloning/2} has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one\footnote{http://jsperf.com/longer-transferrable-vs-cloning/2} 1,000,000 (a million).
-As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects; it scales.
-On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.
-For that reason our library only works with \verb+SharedTypedArray+s and \verb+TypedArrays+.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|l|l|}
-    \hline
-    Computer & Mac Book Air \\
-    Processor & Intel Core i5 \\
-    Clock Frequency & 1.8 GHz \\
-    System Memory & 4 GB \\
-    Operating System & OS X Yosemite 10.10.3 \\
-    Browser version & Firefox Nightly 41.0a1 \\
-    \hline
-  \end{tabular}
-  \caption{Benchmarks environment.}
-  \label{tab:env}
-\end{table}
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|c|}
-    \hline
-    Elements & Cloning [ops/sec] & Transferring [ops/sec] \\
-    \hline
-    \(1E5\) & 734 & \textbf{1596} \\
-    \(1E6\) & 91 & \textbf{1134} \\
-    \hline
-  \end{tabular}
-  \caption{Difference between transferring and cloning shared buffers.}
-  \label{tab:transf-vs-clone}
-\end{table}
-
-\subsection{Functions}\label{sub-sec:functions}
-The library must handle the distribution of the code between the different workers when an operation is executed.
-This is a challenging problem with many possible alternatives: Cloud Haskell \cite{cloud-haskell} for example, requires a customized version of the Glasgow Haskell Compiler (GHC) to serialize closures and their captured variables as a code pointer and an environment, operating under the assumption that the same code is executed in all nodes.
-A proposal for using Scheme for distributed programming \cite{distributed-scheme} uses a custom serializable data structure to represent procedures, and macros in order to capture a closure's captured variables and serialize their values.
-It also assumes that all nodes are executing the same program since serialized functions store pointers to parts of source code files.
-
-In our case, modifying the runtime is not a viable alternative since we want our library to be usable by anyone creating applications for major browser versions.
-
-A naive approach would be to try to transfer a function as part of the \verb+message+ parameter of \verb+postMessage+ but that is not an option (it throws an error).
-
-Considering the fact that \verb+Function+ objects cannot be directly transferred a different serialization approach is to be considered a possible alternative is to:
-\begin{enumerate}
-  \item Serialize the \verb+Function+ to a \verb+String+.
-  \item Transfer the \verb+String+ to the worker.
-  \item Create a new \verb+Function+ on the worker from that \verb+String+.
-\end{enumerate}
-
-Sec.~\ref{sub-sec:context} proposes a solution to sharing additional data and functions between the UI thread and Web Workers as a replacement for variables captured by closures.
-
-\subsubsection{Serialization and deserialization}
-One possible way of passing serialized \verb+Function+ to workers would be to encode them as binary and transfer them as an \verb+ArrayBuffer+. To encode the strings there are two possible approaches:
-\begin{itemize}
-  \item Using the Encoding API\cite{encoding-api}
-  \item Implementing non native encoding/decoding functions
-\end{itemize}
-
-The results for the benchmark\footnote{http://jsperf.com/pjs-serialization-comparison/2} are in Tab.~\ref{tab:encode-decode}:
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|}
-    \hline
-    Native encoding API [ops/sec] & Non-native function [ops/sec] \\
-    \hline
-    \textbf{5017} & 4398 \\
-    \hline
-  \end{tabular}
-  \caption{Difference encoding/decoding strings natively and in JavaScript.}
-  \label{tab:encode-decode}
-\end{table}
-
-\subsubsection{Transference}
-\verb+Function+'s code not only needs to be serialized but also transferred.
-For that reason it was also worth comparing the time it takes to encode/decode a \verb+String+ and tranfer the resulting \verb+ArrayBuffer+ to the worker against just passing the \verb+String+ to the worker.
-
-Additionally, in most scenarios code would have to be sent along with a \verb+TypedArray+ so it would also be interesting to see if the transfer time was affected by this fact.
-
-We created benchmarks\footnote{http://jsperf.com/pjs-serialization-long/3 and http://jsperf.com/pjs-serialization/3} using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time (results in Tab.~\ref{tab:send-function}).
-We measured three approaches to transfer the string:
-\begin{enumerate}
-  \item Encoding the \verb+String+ into an \verb+TypedArray+, sending it as a \verb+Tranferable+ and decoding it on the Web Worker.
-  \item Sending the \verb+String+ directly.
-  \item Using a \verb+Blob+ to store the stream and retrieving that in the Worker.
-\end{enumerate}
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    Function length & Blob [ops/sec] & Copy [ops/sec] & Tranferable [ops/sec] \\
-    \hline
-    Short & 1070 & \textbf{6629} & 5927 \\
-    Long & 1075 & \textbf{3273} & 2967 \\
-    \hline
-  \end{tabular}
-  \caption{Sending function strings from UI thread to Web Workers.}
-  \label{tab:send-function}
-\end{table}
-
-We created a separate benchmark\footnote{http://jsperf.com/shared-function-transfer} to compare the best option (copying) against encoding the function into a \verb+SharedTypedArray+ and transferring it to four workers.
-We expected the latter to be slower since it involves encoding the code once and decoding it in each worker.
-The results are shown in Tab.~\ref{tab:copy-vs-shared}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|}
-    \hline
-    Copying [ops/sec] & SharedArrayBuffer [ops/sec] \\
-    \hline
-    \textbf{1315} & 1226 \\
-    \hline
-  \end{tabular}
-  \caption{Sending code to four workers by copying compared to encoding it and using a SharedArrayBuffer.}
-  \label{tab:copy-vs-shared}
-\end{table}
-
-We also created a benchmark\footnote{http://jsperf.com/pjs-encoding/2} that tranfers a \verb+SharedTypedArray+ and a \verb+String+ that are properties of the same \verb+message+ object back and forth between the UI thread an a Web Worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \verb+String+ (results in Tab.~\ref{tab:send-function-and-array}).
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|}
-    \hline
-    Copying code [ops/sec] & Tranferring code [ops/sec] \\
-    \hline
-    \textbf{4509} & 4270 \\
-    \hline
-  \end{tabular}
-  \caption{Sending code and a TypedArray to a Web Worker.}
-  \label{tab:send-function-and-array}
-\end{table}
-
-\subsection{Contexts}\label{sub-sec:context}
-Functions in JavaScript can capture variables as long as they are within the variable's lexical scope.
-Functions that make use  of this feature are closures.
-As explained in subsection~\ref{sub-sec:functions}, \verb+Function+ objects cannot be passed as messages to workers.
-This means that there is no out of the box mechanism to make variables that are available in the environment when the parallel computation is started accessible when executing the callback functions in each Web Worker.
-
-To solve this problem, p-j-s introduces the \emph{context} concept. There are two kinds of contexts:
-\begin{description}
-\item[Global context] Shared across all chains.
-\item[Local context] Specific to a particular step in a chain.
-\end{description}
-
-The following code snippet shows how developers can work with each type of context:
-\medskip
-
-\noindent
-{\it p-j-s context sample}
-\begin{verbatim}
-var xs = new SharedUint8Array(4);
-xs.set([1,2,3,4]);
-
-pjs.updateContext({
-  max: 3
-}).then(function(){
-  pjs(xs).filter(function(e, ctx){
-    return e <= ctx.max && e >= ctx.min;
-  }, {
-    min: 2
-  }).seq().then(function(range){
-    // range is [2,3]
-  });
-});
-
-\end{verbatim}
-%
-\noindent
-{\small (Example using local and global context for chains)}\\
-
-When executing each step in a chain, the local context is merged with the global context (neither is modified) into a separate object which is the \verb+ctx+ parameter that the function receives.
-If the same key is present in both the global and local context, the one in the local context is the one that will be available.
-Given the following code snippet, the available \verb+ctx+ for each function is displayed in Fig.~\ref{fig:context}.
-\medskip
-
-\noindent
-{\it p-j-s context merge sample}
-\begin{verbatim}
-var xs = new SharedUint8Array(4);
-xs.set([1,2,3,4]);
-
-pjs.updateContext({
-  A: 0,
-  B: 25,
-  C: 44
-}).then(function(){
-  var chain = pjs(xs)
-    .map(function(e, ctx){
-      // code here
-    }, { A: 5, B: 13 })
-    .filter(function(e, ctx){
-      // code here
-    }, { A: 2, E: 9 })
-});
-
-\end{verbatim}
-%
-\noindent
-{\small (Example for resulting ctx)}\\
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth]{context}
-\caption{Example of global and local context being merged for chain steps.}
-\label{fig:context}
-\end{figure}
-
-One alternative to hide the \emph{context} concept from application developers could be to introduce a \textit{build} step that detects the variables captured by p-j-s callbacks and transforms the code.
-By doing this, they wouldn't need to do anything out of the ordinary with regards to closures and their environment when working with p-j-s.
-However, this approach would require transforming the original code, and would be complex and time consuming to implement.
-
-\section{Experiment}\label{sec:experiment}
-It is important to compare the performance of our parallel implementation against a serial one, as some processing is required before the computation begins and after it ends to distribute the data and work and gather the results respectively, as explained in Section~\ref{sec:implementation}.
-
-In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform.
-The reason for this is that, in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.
-
-\subsection{Algorithm selection}
-For the experiment we decided to use an algorithm that applies a sepia tone effect to pictures.
-This meets the criteria of both being useful and having some calls to \verb+Math.random+ which make it more complex from a computational standpoint.
-The benchmark is publicly available \cite{benchmark}.
-
-\subsection{Results}
-The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark}.
-Based on those results, we created a chart (Figure~\ref{fig:relative}) that shows the relative performance of the parallel approach compared to the sequential one.
-\begin{table}
-  \centering
-  \begin{tabular}{|r|c|c|}
-    \hline
-    Elements & Serial [ops/sec] & Parallel [ops/sec] \\
-    \hline
-    100 & \textbf{142223} & 328 \\
-    1,000 & \textbf{16503.8} &  359 \\
-    10,000 & \textbf{1683.8} & 321.8 \\
-    100,000 & 169.6 & \textbf{184.4} \\
-    1,000,000 & 16.8 & \textbf{34.1} \\
-    \hline
-  \end{tabular}
-  \caption{Comparing p-j-s (parallel) and plain JS (serial) performance for an image transformation.}
-  \label{tab:benchmark}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[height=7.7cm]{comparison}
-\caption{Relative performance of parallel transformation compared to sequential one.}
-\label{fig:relative}
-\end{figure}
-
-\subsection{Conclusion}
-A couple of interesting findings result from the experiments.
-An expected one is that the amount of operations per second that can be performed with a serial implementation is almost inversely proportional to the amount of items to transform with a ratio of 1.
-On the other hand, this is not the case for the parallel approach.
-
-An acceptable speedup of approximately \textbf{2x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
-Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \verb+SharedArrayBuffer+s as they are likely not optimized, being that they are only an experimental feature in Firefox Nightly.
-
-\section{Future Work}\label{sec:future}
-There are still some parts of the library that would benefit from optimizations.
-Of particular interest to us is to perform further profiling of the code executing in each Web Worker.
-As of today, using browser profiling tools for Web Workers is not as simple as doing so for the UI thread.\\
-
-During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times.
-Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high, performance degrades.
-We plan to spend some time on this issue.\\
-
-There is a strawman proposal to support typed objects \cite{typed-objects} in a future version of EcmaScript, which in turn would allow \verb+TypedArray+s of said typed objects.
-Once a browser has an experimental version available we intend to make sure that typed arrays of typed objects work with the library.\\
-
-Once the \verb+SharedArrayBuffer+ type is available in other browsers (there is already a proposal for Google Chrome \cite{sab-chrome}) we plan to perform benchmarks and tests to compare with Firefox Nightly.
-
-\subsubsection*{Acknowledgments.}
-Thanks to Rosa Wachenhauzer, our tutor, for reviewing this paper and also for all her support and guidance throughout this entire project.
-
-\begin{thebibliography}{4}
-
-\bibitem{moore65} Moore, G.E.: Cramming More Components onto Integrated Circuits. (1965)
-
-\bibitem{ms-par} Parallel Programming in the .NET Framework, Microsoft Corp., \url{http://msdn.microsoft.com/en-us/library/dd460693(v=vs.110).aspx}
-
-\bibitem{es-web} EcmaScript, ECMA, \url{http://www.ecmascript.org/}
-
-\bibitem{asm-web} asm.js, \url{http://asmjs.org/}
-
-\bibitem{webgl-web} WebGL, Khronos Group, \url{http://www.khronos.org/webgl/}
-
-\bibitem{rivertrail} Herhut, S., Hudson, R. L., Shpeisman, T., Sreeram, J.: River trail: A path to parallelism in javascript. SIGPLAN Not., 48(10):729--744, (2013).
-
-\bibitem{par-js} Wang, J., Rubin, N., Yalamanchili, S.: River trail: ParallelJS: An Execution Framework for JavaScript on Heterogeneous Systems. SIGPLAN Not. (2014)
-
-\bibitem{js-simd} McCutchan, J., Feng, H., Matsakis, N. D., Anderson, Z., Jensen, P.: A SIMD Programming Model for Dart, JavaScript, and other dynamically typed scripting languages (2014)
-
-\bibitem{w3c-ww} Web Workers, W3C, \url{http://www.w3.org/TR/workers/}
-
-\bibitem{event-loop} Concurrency model and Event Loop, Mozilla Developer Network, \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/EventLoop}
-
-\bibitem{v8} v8, Google, \url{https://developers.google.com/v8/intro}
-
-\bibitem{spider-monkey} SpiderMonkey, Mozilla, \url{https://developer.mozilla.org/en-US/docs/Mozilla/Projects/SpiderMonkey}
-
-\bibitem{thread-pool} Schmidt, D. C., Vinoski, S.,: Comparing Alternative Programming Techniques for Multi-threaded CORBA Servers: Thread Pool (1996)
-
-\bibitem{sab} JavaScript Shared Memory, Atomics, and Locks, \url{https://docs.google.com/document/d/1NDGA_gZJ7M7w1Bh8S0AoDyEqwDdRh4uSoTPSNn77PFk/edit}
-
-\bibitem{es6-ta} TypedArray objects, ECMA, \url{https://people.mozilla.org/~jorendorff/es6-draft.html#sec-typedarray-objects}
-
-\bibitem{pjs-api} p-j-s API, \url{https://github.com/pjsteam/pjs/blob/dev/README.md}
-
-\bibitem{promise} Promise, MDN, \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Global_Objects/Promise}
-
-\bibitem{html-whatwg} HTML Standard, WHATWG, \url{https://html.spec.whatwg.org}
-
-\bibitem{cloud-haskell} Epstein, J., Black, A. P., Peyton-Jones, S.: Towards Haskell in the Cloud (2011)
-
-\bibitem{distributed-scheme} Schwendner, A.: Distributed Functional Programming in Scheme (2010)
-
-\bibitem{encoding-api} Encoding Standard, WHATWG, \url{https://encoding.spec.whatwg.org}
-
-\bibitem{benchmark} Sepia tone transformation benchmark comparison between p-j-s and plain JS, \url{http://jsperf.com/pjs-shared-vs-normal-vs-seq-100/4}
-
-\bibitem{typed-objects} Typed Objects, ECMA, \url{http://wiki.ecmascript.org/doku.php?id=harmony:typed_objects}
-
-\bibitem{sab-chrome} Shared Array Buffers, Chromium Dashboard, \url{https://www.chromestatus.com/feature/4570991992766464}
-
-\end{thebibliography}
-
-\end{document}
+\pagebreak

### 2015-05-23 docs/report/code_sharing.tex

@@
 \section{Code sharing}
+The library must handle the distribution of the code between the different workers when an operation is executed.
+This is a challenging problem with many possible alternatives: Cloud Haskell \cite{cloud-haskell} for example, requires a customized version of the Glasgow Haskell Compiler (GHC) to serialize closures and their captured variables as a code pointer and an environment, operating under the assumption that the same code is executed in all nodes.
+A proposal for using Scheme for distributed programming \cite{distributed-scheme} uses a custom serializable data structure to represent procedures, and macros in order to capture a closure's captured variables and serialize their values.
+It also assumes that all nodes are executing the same program since serialized functions store pointers to parts of source code files.
 
-One of the aspects that must be handled by the library is the distribution of the code between the different workers when a skeleton is created.
+In our case, modifying the runtime is not a viable alternative since we want our library to be usable by anyone creating applications for major browser versions.
 
@@
 
-That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a \code{Function} does not meet that requirement as explained in section 2.7.5 of that same standard.
-It could be assumed that this is because of the challenges of serializing function closures\cite{cloud-haskell}\cite{distributed-scheme}.
+That error is expected.
+As explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a \code{Function} does not meet that requirement as explained in section 2.7.5 of that same standard.
+It could be assumed that this is because of the challenges of serializing functions.
 
 \subsection{Serializing functions}
-Considering the fact that \tfunction{} objects cannot be directly transferred a different serialization approach is to be considered.
+Considering the fact that \tfunction{} objects cannot be directly transferred, a different serialization approach is to be considered.
 Since function closures will not be transferrable a possible alternative is to:
@@
 \subsubsection{\tstring{} serialization benchmarks}
-One possible way of passing serialized \tfunction{} to workers would be to encode them as binary and transfer them as an \tabuffer{}. To encode the strings there are two possible approaches:
+One possible way of passing serialized \tfunction{}s to workers would be to encode them as binary and transfer them as an \tabuffer{}. To encode the strings there are two possible approaches:
 \begin{itemize}
-  \item Using the Encoding API\cite{encoding-api}
-  \item Implementing non native encoding/decoding functions
+  \item Use the Encoding API\cite{encoding-api}
+  \item Implement non native encoding/decoding functions
 \end{itemize}
@@
   \item Sending the string directly
-  \item Using a \code{Blob} to store the stream and retrieving that in the Worker.
+  \item Using a \code{Blob} to store the stream and retrieving it in the Worker.
 \end{enumerate}

### 2015-05-23 docs/report/initialization.tex

@@
 This factor raises the issue of when workers should be instantiated.
-On relation to this, we performed a benchmark to get an idea of the difference between cold starting a worker against having an already available one process messages.
+On relation to this, we performed a benchmark to get an idea of the difference in time taken to process messages.
+One scenario considers cold starting a Web Worker, and the other one having a Web Worker already available.
 
@@
 \subsubsection{Conclusion}
-The previous benchmark shows a huge difference between the time required to process a message on a pre-existent Web Worker and that necessary to start Web Worker and process a message.
-For that reason, we decided to create a pool of Web Workers when the library is initialized.
+The previous benchmark shows a huge difference between the time required to process a message on a pre-existent Web Worker compared to the one necessary to start a Web Worker and process a message.
+For that reason, we decided to create a pool of Web Workers when the library is initialized and dispatch work to them when it arrives.
+Since operations are asynchronous, if an operation is executing when another one is added for processing, the latter will be added to a queue and processed when the former completes.
 
@@
 Since the work being done is CPU intensive (i.e.: the workers' goal is not to perform I/O operations), the maximum amount of workers to be created should not be greater than the amount of cores in the machine \(W_{cores}\), otherwise interleaving would start to affect performance.
-Additionally, considering the necessity of users to control how many workers are created, an upper bound can be specified \(W_{upper bound}\).
+Additionally, considering the necessity of users to control how many workers are created, an upper bound can be specified \(W_{upper bound}\)\footnote{If the executing browser does not have an API to detect the amount of cores in a device \(W_{cores}=W_{upper bound}\).
+If no \(W_{upper bound}\) is provided, \(W_{cores}=1\).}.
 If no upper bound is specified, the user upper bound for workers is assumed to be infinite (because the user is not providing a restriction).

### 2015-05-23 docs/report/inlining.tex

@@
 \section{The importance of function inlining}
-One of the things found while comparing the performance of the parallel and non-parallel implementations of \code{map} is that our parallel implementation performed a lot worse than what the one expected when considering the size of the \ttarray{} and the \code{mapper} function.
+One of the things found while comparing the performance of the parallel and non-parallel implementations of \code{map} is that our parallel implementation performed a lot worse than expected when considering the size of the \ttarray{} and the \code{mapper} function.
 
@@
 As the two code samples were similar our hypothesis was that for some reason some v8 optimizations were not taking place.
-It was a definite possibility considering that the parallel implementation depends on the \tfunction constructor to create \tfunction objects in the Web Workers.
+It was a definite possibility considering that the parallel implementation depends on the \tfunction{} constructor to create \tfunction{} objects in the Web Workers.
 Specifically we believed that the internal functions \code{noise}, \code{clamp} and \code{colorDistance} were not being inlined.

### 2015-05-23 spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark.js
updates test and result image.

@@ -1,4 +1,4 @@
-// http://jsperf.com/additional-sendingdata-comparison/3
+// http://jsperf.com/additional-sendingdata-comparison/2
 
 // HTML setup
 <script>

### 2015-05-23 docs/report/context.tex
Context initial version

\section{Context}\label{sec:context}

Functions in JavaScript can capture variables as long as they are within the variable's lexical scope. Functions that make use of this feature are closures. As explained in Section~\ref{sec:code-sharing}, \code{Function} objects cannot be passed as messages to workers. This means that there is no out of the box mechanism to make variables that are available in the environment when the parallel computation is started accessible when executing the callback functions in each Web Worker.

To solve this problem, p-j-s introduces the \emph{context} concept. There are two kinds of contexts:
\begin{description}
\item[Global context] Shared across all chains.
\item[Local context] Specific to a particular step in a chain.
\end{description}

The following code snippet shows how developers can work with each type of context:
\begin{lstlisting}[caption=Example using local and global context for chains]
var xs = new SharedUint8Array(4);
xs.set([1,2,3,4]);

pjs.updateContext({
  max: 3
}).then(function(){
  pjs(xs).filter(function(e, ctx){
    return e <= ctx.max && e >= ctx.min;
  }, {
    min: 2
  }).seq().then(function(range){
    // range is [2,3]
  });
});
\end{lstlisting}

When executing each step in a chain, the local context is merged with the global context (neither is modified) into a separate object which is the \code{ctx} parameter that the function receives. If the same key is present in both the global and local context, the one in the local context is the one that will be available. Given the following code snippet, the available \code{ctx} for each function is displayed in Fig.~\ref{fig:context}.
\begin{lstlisting}[caption=Example for resulting ctx]
var xs = new SharedUint8Array(4);
xs.set([1,2,3,4]);

pjs.updateContext({
  A: 0,
  B: 25,
  C: 44
}).then(function(){
  var chain = pjs(xs)
    .map(function(e, ctx){
      // code here
    }, { A: 5, B: 13 })
    .filter(function(e, ctx){
      // code here
    }, { A: 2, E: 9 })
});
\end{lstlisting}

\img{Comparing \ttarray{} merge approaches}{../../spikes/6.1/context}{fig:context}

\subsection{Implementation}
Internally the context is sent as a message to all workers to make its properties available. Two possible implementatios were possible:
\begin{itemize}
  \item Send the object directly and rely on structured cloning
  \item Serialize the object using \code{JSON.stringify} and deserialize it in the workers
\end{itemize}

\note{None of the aforementioned approaches support sending functions as context keys. This is covered in Subsection~\ref{sub-sec:contexts-with-function}.}

We created to separate benchmarks, one that sends small objects to a Web Worker and another one that sends large objects to a Web Worker. Both benchmarks compare the approaches against each other.

The code for the benchmark with the small object is in the following listing:
\bmcode{http://jsperf.com/additional-clonningdata-comparison/3}{../../spikes/6.1/cloningComparison/jspBenchmark.js}

The results for the benchmark are displayed in Figure~\ref{fig:context-clone-small-cmp}.
\img{Comparing small context message passing}{../../spikes/6.1/cloningComparison/result}{fig:context-clone-small-cmp}

The code for the benchmark with the large object is in the following listing:
\bmcode{http://jsperf.com/additional-clonningdata-comparison-long/3}{../../spikes/6.1/cloningComparison/jspBenchmarkLong.js}

The results for the benchmark are displayed in Figure~\ref{fig:context-clone-large-cmp}.
\img{Comparing large context message passing}{../../spikes/6.1/cloningComparison/resultLong}{fig:context-clone-large-cmp}

\subsection{Contexts with functions}\label{sub-sec:contexts-with-function}
Properties whose values are \code{Function} instances are serialized into an object. For example if the following object were passed as a context:
\begin{lstlisting}[caption=Example context with a function property]
{
  a: 2,
  b: function(c) {
    return c + 1;
  }
}
\end{lstlisting}

It would result in the following object before being serialized:
\begin{lstlisting}[caption=Example processed context]
{
  a: 2,
  b: {
    args: [ 'c' ],
    code: '\n\treturn c + 1;\n',
    __isFunction: true
  }
}
\end{lstlisting}

\subsection{Conclusion}
Based on the benchmark results we decided to use \code{JSON.stringify} to serialize \code{context} objects. This involves a tradeoff as, unlike structured cloning, it does not support cyclic references. For this scenario, we believe that the performance benefit is more valuable than supporting cyclic references.

\pagebreak

### 2015-05-23 docs/report/context.tex

@@
 
+The reason why there are two kinds of context is because there will usually be data that does not need to be passed to the Web Workers with each chain, so having a global context allows this.
+Additionally, the local context allows users to avoid polluting the global one with keys that are not necessary for all operations.
+
+We created a benchmark to explore the difference between sending the context and chains to execute as part of the same message, and using two different messages.
+The code for the benchmark is in the following listing:
+\bmcode{http://jsperf.com/additional-sendingdata-comparison/2}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark.js}
+
+The results for the benchmark are displayed in Figure~\ref{fig:context-send}.
+\img{Comparing small context message passing}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark-all-browsers}{fig:context-send}
+
 \subsection{Implementation}

### 2015-05-23 docs/report/code_sharing.tex
Added link to context section from code sharing

@@
-\section{Code sharing}
+\section{Code sharing}\label{sec:code-sharing}
 The library must handle the distribution of the code between the different workers when an operation is executed.
@@
 
-\note{An approach to provide contextual data to worker functions is explored in \todo{link to section}.}
+\note{An approach to provide contextual data to worker functions is explored in Section~\ref{sec:context}.}
 

### 2015-05-23 docs/report/overview.tex
Added library overview section

\section{Library Overview}\label{sec:overview}
JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs. In JavaScript, \code{Array}s have \code{map}, \code{filter}, and \code{reduce} functions and version 6 of the EcmaScript standard adds those same methods to \ttarray{}s \cite{es6-ta}. For example, to multiply all the elements of an \code{Array} by two:
\begin{lstlisting}[caption=Example creating a new array with the values of the original one multiplied by two]
var newValues = [1,2,3,4].map(function(e){
  return e * 2;
});

// newValues is a new array [2,4,6,8]
\end{lstlisting}

Our library (p-j-s) only supports \tstarray{}s and \ttarray{}s due to performance reasons discussed in Section~\ref{sec:serialization-and-transference}. It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation. Given a \tstarray{}, the following code would do the same as the previous example but parallelizing the work:
\begin{lstlisting}[caption=Example creating a new array with the values of the original one multiplied by two using p-j-s]
var pjs = require('p-j-s');
pjs.init({ maxWorkers: 4 });
var xs = new SharedUint8Array(4);
xs.set([1,2,3,4]);
pjs(xs).map(function(e){
  return e * 2;
}).seq().then(function(newValues){
  // newValues is a new SharedUint8Array [2,4,6,8]
});
\end{lstlisting}

The call to \code{pjs.init} initializes the library and the amount of workers to be used for processing (a worker pool is created with them).\footnote{More details in Section~\ref{sec:initialization}.}

The API provided by p-j-s \cite{pjs-api} is very similar to the synchronous one that is provided by the language. The asynchronous nature of the operation is reflected by the fact that the call to \code{seq} returns a \code{Promise} \cite{promise} (alternatively a callback function could be passed as a parameter to it).

It is possible to chain multiple operations before calling \code{seq} to avoid passing data back and forth between the Web Workers and the UI thread. Before \code{seq} is invoked, an operation is a chain of one or more steps. From the same main chain, multiple chains could be created:

\begin{lstlisting}[caption=Example creating multiple chains from the same base chain]
var xs = new SharedUint8Array(4);
xs.set([1,2,3,4]);

var chain = pjs(xs).map(function(e){
  return e * 3;
});

var evenChain = chain.filter(function(e){
  return e % 2 === 0;
});

var sumChain = chain.reduce(function(c, v){
  return c + v;
}, 0, 0);

evenChain.seq().then(function(evens){
  // evens is [6, 12]
  sumChain.seq().then(function(sum){
    // sum is [3,6,9,12]
  });
});
\end{lstlisting}

\pagebreak

### 2015-05-23 spikes/3.1.3/comparison/jspBenchmark.js
updates tests and result images.

@@ -1,10 +1,10 @@
-// http://jsperf.com/pjs-map-vs-serial/3
+// http://jsperf.com/pjs-map-vs-serial/8
 
 // HTML preparation
-<script src="https://rawgit.com/pjsteam/pjs/v0.1.1/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
-  pjs.init();
+  pjs.init({ maxWorkers: 4});
 
   var generateElements = function (total) {
     var typed = new Uint32Array(total);
@@ -69,7 +69,7 @@
   };
 
   function runPjs(wrappedXs) {
-    wrappedXs.map(mapper, function (r) {
+    wrappedXs.map(mapper).seq(function (err, r) {
       __finish();
     });
   };

### 2015-05-23 spikes/3.1.3/pjsMapInlining/pjsMapInlining.js

@@ -1,10 +1,10 @@
-// http://jsperf.com/pjs-map-inlining/2
+// http://jsperf.com/pjs-map-inlining/6
 
 // HTML setup
-<script src="https://rawgit.com/pjsteam/pjs/v0.1.1/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
-  pjs.init();
+  pjs.init({ maxWorkers: 4});
 
   var generateElements = function (total) {
     var typed = new Uint32Array(total);
@@ -56,20 +56,33 @@
 
     return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
   };
+
+  function notInlinedTest() {
+    wrappedXs.map(notInlinedMapper).seq(function (result) {
+      if (!result) {
+        console.log('error');
+      }
+      __finish();
+    });
+  };
+
+  function inlinedTest() {
+    wrappedXs.map(inlinedMapper).seq(function (result) {
+      if (!result) {
+        console.log('error');
+      }
+      __finish();
+    });
+  };
 </script>
 
-// Test case 1 - Not inlined map
-wrappedXs.map(notInlinedMapper, function (result) {
-  if (!result) {
-    console.log('error');
-  }
+// Javascript setup
+__finish = function () {
   deferred.resolve();
-});
+};
+
+// Test case 1 - Not inlined map
+notInlinedTest();
 
 // Test case 2 - Inlined map
-wrappedXs.map(inlinedMapper, function (result) {
-  if (!result) {
-    console.log('error');
-  }
-  deferred.resolve();
-});
+inlinedTest();

### 2015-05-23 spikes/6.2/inlinedVsGlobalInlined/jspBenchmark.js

@@ -1,7 +1,7 @@
-// http://jsperf.com/global-ctx-inlined-func
+// http://jsperf.com/global-ctx-inlined-func/3
 
 // HTML setup
-<script src="http://rawgit.com/pjsteam/pjs/v0.5.2/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
   pjs.init();
@@ -14,7 +14,7 @@
     return typed;
   };
 
-  var xs = generateElements(50000);
+  var xs = generateElements(100000); // 100.000
 
   var wrappedXs = pjs(xs);
 
@@ -34,7 +34,7 @@
   };
 
   pjs.updateContext({
-    inlinedMapper: inlinedMapper
+    x: inlinedMapper
   });
 </script>
 
@@ -48,10 +48,10 @@ var __finish = function (result) {
 
 // Test case 1 - map with inlining
 wrappedXs.map(inlinedMapper).seq(function (result) {
-  __finish(result);
+  __finish();
 });
 
 // Test case 2 - map with global inlining
-  wrappedXs.map('inlinedMapper').seq(function (result) {
-    __finish(result);
-  });
+wrappedXs.map('x').seq(function (result) {
+  __finish();
+});

### 2015-05-23 spikes/6.2/inlinedVsGlobalNotInlined/jspBenchmark.js

@@ -1,7 +1,7 @@
-// http://jsperf.com/global-ctx-no-inlined-vs-inlined
+// http://jsperf.com/global-ctx-no-inlined-vs-inlined/2
 
 // HTML setup
-<script src="http://rawgit.com/pjsteam/pjs/v0.5.2/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
   pjs.init();

### 2015-05-23 spikes/6.2/noInlinedVsGlobalNotInlinedWithContext/jspBenchmark.js

@@ -1,7 +1,7 @@
-// http://jsperf.com/global-ctx-no-inlined-func
+// http://jsperf.com/global-ctx-no-inlined-func/2
 
 // HTML setup
-<script src="http://rawgit.com/pjsteam/pjs/v0.5.2/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
   pjs.init();

### 2015-05-23 spikes/6.2/noInlinedVsGlobalWithGlobalContext/jspBenchmark.js

@@ -1,7 +1,7 @@
-// http://jsperf.com/all-global-ctx-no-inlined-func
+// http://jsperf.com/all-global-ctx-no-inlined-func/2
 
 // HTML setup
-<script src="http://rawgit.com/pjsteam/pjs/v0.5.2/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var pjs = require('p-j-s');
   pjs.init();

### 2015-05-23 docs/paper/paper.tex

@@
 JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs.
-In JavaScript, \verb+Array+s have \verb+map+, \verb+filter+, and \verb+reduce+ functions and version 6 of the EcmaScript adds those same methods to \verb+TypedArray+s \cite{es6-ta}.
+In JavaScript, \verb+Array+s have \verb+map+, \verb+filter+, and \verb+reduce+ functions and version 6 of the EcmaScript standard adds those same methods to \verb+TypedArray+s \cite{es6-ta}.
 For example, to multiply all the elements of an \verb+Array+ by two:
@@
 The API provided by p-j-s \cite{pjs-api} is very similar to the synchronous one that is provided by the language.
-The asynchronous nature of the operation is reflected by the fact that the call to \verb+seq+ returns a \verb+Promise+ (alternatively a callback could be passed as a parameter to it).
+The asynchronous nature of the operation is reflected by the fact that the call to \verb+seq+ returns a \verb+Promise+ \cite{promise} (alternatively a callback could be passed as a parameter to it).
 

### 2015-05-23 docs/report/function_transfer_time.tex
Initial function transfer time

\section{Decreasing function transfer time}\label{sec:function-transfer-time}

Each time \code{seq} is invoked on a chain, all callback functions related to \code{map}, \code{filter} and \code{reduce} steps are serialized and sent to the Web Workers.

Each Web Worker has a cache to avoid using different instances of the same function (as explained in Section~\ref{sec:func-caching}). Nevertheless, if the same function is going to be used multiple times in a program as the callback for steps, it might make sense to send them to the Web Workers once and then simply make use of them.

This can be achived by taking advantage of the global context object (detailed in Section~\ref{sec:context}).

The first thing that needs to be done is to call \code{pjs.updateContext} by providing a \tfunction{} as a property of the context parameter. An example is shown in the following code snippet:
\begin{lstlisting}[caption=Sending function in global context]
var promise = pjs.updateContext({
  add: function (x) { return x + 2; },
});
\end{lstlisting}

Once the global context has been updated, the \tstring{} representation of the function's key (\code{'add'}) can be used as a callback parameter instead of using a \tfunction{} instance. An example is shown in the following code snippet:
\begin{lstlisting}[caption=Passing a function name from the global context instead of a funciton as a callback]
var promise = pjs.updateContext({
  add: function (x) { return x + 2; },
}).then(function(){
  pjs(new Uint8Array([1,2,3,4])).map('add').seq()
    .then(function(result){
      // result is [ 3, 4, 5, 6 ]
    });
});
\end{lstlisting}

\subsection{Benchmark}
We created a benchmark to analyze the improvement obtained by just sending a \tstring{} instead of a \tfunction{} instance.

\note{The improvement will vary if the difference in character length between the key and the function's code does.}

### 2015-05-23 docs/report/function_transfer_time.tex

@@
 
-This can be achived by taking advantage of the global context object (detailed in Section~\ref{sec:context}).
+This can be achieved by taking advantage of the global context object (detailed in Section~\ref{sec:context}).
 
@@
 
-\note{The improvement will vary if the difference in character length between the key and the function's code does.}
+The following listing shows the code for the benchmark:
+\bmcode{http://jsperf.com/global-ctx-inlined-func/3}{../../spikes/6.2/inlinedVsGlobalInlined/jspBenchmark.js}
+
+The results for the benchmark are displayed in Figure~\ref{fig:func-key-vs-instance}.
+\img{Comparing sending function key in global context against function instance}{../../spikes/6.2/inlinedVsGlobalInlined/result}{fig:func-key-vs-instance}
+
+\note{The improvement shown in Figure~\ref{fig:func-key-vs-instance} will vary if the difference in character length between the key and the function's code does.}
+
+\pagebreak

### 2015-05-23 dist/p-j-s-standalone.js
removes ww's function cache for benchmarks.

@@ -968,24 +968,6 @@ var immutableExtend = require('xtend/immutable');
 var contextUtils = require('./context');
 var chainContext = require('./chain_context');
 
-function getMapFactory(){
-  if (typeof Map === 'function'){
-    return function(){
-      /* jshint ignore:start */
-      return new Map();
-      /* jshint ignore:end */
-    };
-  } else {
-    return function(){
-      return Object.create(null);
-    };
-  }
-}
-
-var mapFactory = getMapFactory();
-
-var functionCache = mapFactory();
-
 var globalContext = {};
 
 var operations = {
@@ -1025,14 +1007,7 @@ function getFunction(operation){
 
   var args = operation.args;
   var code = operation.code;
-  var cacheKey = args.join(',') + code;
-  var f = functionCache[cacheKey];
-  if (!f){
-    f = utils.createFunction(args, code);
-    functionCache[cacheKey] = f;
-  }
-
-  return f;
+  return utils.createFunction(args, code);
 }
 
 module.exports = function(event){

### 2015-05-23 dist/p-j-s.js

@@ -891,24 +891,6 @@ var immutableExtend = require('xtend/immutable');
 var contextUtils = require('./context');
 var chainContext = require('./chain_context');
 
-function getMapFactory(){
-  if (typeof Map === 'function'){
-    return function(){
-      /* jshint ignore:start */
-      return new Map();
-      /* jshint ignore:end */
-    };
-  } else {
-    return function(){
-      return Object.create(null);
-    };
-  }
-}
-
-var mapFactory = getMapFactory();
-
-var functionCache = mapFactory();
-
 var globalContext = {};
 
 var operations = {
@@ -948,14 +930,7 @@ function getFunction(operation){
 
   var args = operation.args;
   var code = operation.code;
-  var cacheKey = args.join(',') + code;
-  var f = functionCache[cacheKey];
-  if (!f){
-    f = utils.createFunction(args, code);
-    functionCache[cacheKey] = f;
-  }
-
-  return f;
+  return utils.createFunction(args, code);
 }
 
 module.exports = function(event){
@@ -1154,7 +1129,7 @@ WrappedTypedArray.prototype.__operation = function(name, code, localContext, see
 };
 
 module.exports = WrappedTypedArray;
-},{"./chain":4,"./chain_context":5,"./operation_names":10,"./operation_packager":11,"./utils":15}],"p-j-s":[function(require,module,exports){
+},{"./chain":4,"./chain_context":5,"./operation_names":10,"./operation_packager":11,"./utils":15}],"p-j-s-no-cache":[function(require,module,exports){
 var errors = require('./errors');
 var utils = require('./utils');
 var workers = require('./workers');

### 2015-05-23 spikes/3.1.3/functionCache/functionCache.js

@@ -2,7 +2,7 @@
 
 // HTML setup
 <script src="http://127.0.0.1:3000/p-j-s.min.js"></script>
-<script src="https://cdn.rawgit.com/pjsteam/pjs/v0.1.0/dist/p-j-s.min.js"></script>
+<script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>
 <script>
   var __finish;
   var pjs = require('p-j-s');

### 2015-05-23 src/worker_core.js

@@ -4,24 +4,6 @@ var immutableExtend = require('xtend/immutable');
 var contextUtils = require('./context');
 var chainContext = require('./chain_context');
 
-function getMapFactory(){
-  if (typeof Map === 'function'){
-    return function(){
-      /* jshint ignore:start */
-      return new Map();
-      /* jshint ignore:end */
-    };
-  } else {
-    return function(){
-      return Object.create(null);
-    };
-  }
-}
-
-var mapFactory = getMapFactory();
-
-var functionCache = mapFactory();
-
 var globalContext = {};
 
 var operations = {
@@ -61,14 +43,7 @@ function getFunction(operation){
 
   var args = operation.args;
   var code = operation.code;
-  var cacheKey = args.join(',') + code;
-  var f = functionCache[cacheKey];
-  if (!f){
-    f = utils.createFunction(args, code);
-    functionCache[cacheKey] = f;
-  }
-
-  return f;
+  return utils.createFunction(args, code);
 }
 
 module.exports = function(event){

### 2015-05-24 docs/report/serialization_transference.tex

@@
 
-Nevertheless, there are additional time consuming tasks other than the main computation that need to be considered when performing the operation in parallel in programs with no shared memory.
+Nevertheless, there are additional time consuming tasks other than the main computation that need to be considered when performing the operation in parallel in programs where memory is not shared by default.
 These are:
@@
 
-Considering the definition of a \code{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{s}\).
+Considering the definition of a \code{Transferable} the latter seems like a good alternative to minimize both \(T_{s}\) and \(T_{t}\).
 Even more so when one considers that otherwise objects are copied using structured cloning (explained in section \textbf{2.7.6} of that same standard).
+
+In the case of \tsabuffer{} instances, \(T_{p}\) is also minimized.
+Instead of actually making copies of the buffer's data, all that is required a simple calculation to determine the boundaries on which each Web Worker should operate.
+Additionally, \(T_{t}\) should also decrease as there is no need to transfer the \tsabuffer{} back to the UI thread.
 
@@
 To verify the above we put together two benchmarks that aim to verify the difference between invoking \code{postMessage} with and without structured cloning for a \ttarray{}.
-Both transfer a \ttarray{} back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the \ttarray{} and the other one 1,000,000 (a million).
+Since \tsabuffer{}s cannot be cloned, we created a separate benchmark to compare the performance of transferring \tsabuffer{}s and \tabuffer{}s.
 
-\subsubsection{With 100,000 elements}
+The benchmarks transfer the arrays back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the array and the other one 1,000,000 (a million).
+
+\subsubsection{TypedArray With 100,000 elements}
 The following is the code for the benchmark:
@@
 
-\subsubsection{With 1,000,000 elements}
+\subsubsection{TypedArray With 1,000,000 elements}
 The following is the code for the benchmark:
@@
 
+\subsubsection{Transferring SharedArrayBuffers and ArrayBuffers}
+The following is the code for the benchmark:
+\bmcode{http://jsperf.com/transfer-shared-vs-not-shared}{../../spikes/1.2/transferableSharedVsNotShared.js}
+
+The results for it are displayed in Figure~\ref{fig:sab-vs-ab-transfer}.
+\img{Transferring SharedArrayBuffers and ArrayBuffers}{../../spikes/1.2/transferableSharedVsNotShared}{fig:sab-vs-ab-transfer}
+
 \subsubsection{Conclusion}
@@
 
-For that reason we will only work with {\ttarray{}}s and all of them will be passed to/from Web Workers by transferring their ownership.
+For that reason we will only work with \ttarray{}s and \tstarray{}s and all of them will be passed to/from Web Workers by transferring their ownership.
+
+Although it does not affect our decision, it is also interesting to note the big difference between transferring \tsabuffer{}s and \tabuffer{} in Firefox Nightly.
+This difference is probably going to be smaller over the course of time.
 

### 2015-05-24 spikes/2.4.1/sharedMerge.js
adds shared merging benchmarks.

@@ -1,0 +1,116 @@
+// http://jsperf.com/sharedtypedarray-merge
+
+// JavaScript setup
+<script>
+  var arraySize = 250000;
+  var arraysCount = 4;
+
+  function generateElements() {
+    var arrays = [];
+
+    for (var i = 0; i < arraysCount; i++) {
+      arrays[i] = new SharedUint8Array(arraySize);
+    }
+    
+    for (var j = 0; j < arraySize; j++) {
+      for (var i = 0; i < arrays.length; i++) {
+        arrays[i][j] = j + i * arraySize;
+      }
+    }
+    return arrays;
+  };
+
+  var parts = generateElements();
+
+  function caseNonNativeFunction(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var result = new SharedUint8Array(totalSize);
+
+    var from = 0;
+    for (var j = 0; j < parts.length; j++) {
+      var part = parts[j];
+      var partLength = part.length;
+      for (var i = 0; i < partLength; i++) {
+        result[from + i] = part[i];
+      }
+      from += partLength;
+    }
+    return result;
+  };
+
+  function caseTypedArraySet(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var result = new SharedUint8Array(totalSize);
+
+    var from = 0;
+    for (var j = 0; j < parts.length; j++) {
+      var array = parts[j];
+      result.set(array, from);
+      from += array.length;
+    }
+    return result;
+  };
+
+  function caseArrayLikeFunction(parts) {
+    var totalSize = 0;
+    for (var j = 0; j < parts.length; j++) {
+      totalSize += parts[j].length;
+    }
+    var arrayLike = new Array(totalSize);
+    var from = 0;
+    for (var i = 0; i < parts.length; i++) {
+      var part = parts[i];
+      var partLength = part.length;
+      for (var j = 0; j < partLength; j++) {
+        arrayLike[from + j] = part[j];
+      }
+      from += partLength;
+    }
+    var result = new SharedUint8Array(arrayLike.length);
+    result.set(arrayLike);
+    return result;
+  }
+
+  function withNonNativeFunction(parts) {
+    var result = caseNonNativeFunction(parts);
+    if (result.length !== (arraySize * arraysCount)) {
+      console.log('error');
+    }
+  }
+
[…]

### 2015-05-24 docs/report/array_partitioning.tex
Renamed array sharing to array partitioning

--- 2015-05-23 docs/report/array_sharing.tex
@@
-\section{Array sharing}
+\section{Array partitioning}
 
 As explained in section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance.
-Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned}
+Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned.}
 

### 2015-05-24 docs/report/array_merging.tex
Added SAB description for array merging

@@
 
-Once all the workers have executed the function provided to them on each element of their respective \ttarray{}s they are to notify the main thread and provide the resulting \ttarray{}.
+Once all the workers have executed the function provided to them on each element of their portion of the original array, they are to notify the main thread and provide the resulting \tabuffer{} or \tsabuffer{}.
 
-At this point the main thread must consolidate the results into a new \ttarray{} that represents the result for the entire operation.
-This means copying all the elements of each partial result into a larger \ttarray{} while maintaining the original order.
+At this point the UI thread must consolidate the results from the Web Workers into a new \ttarray{} or \tstarray{} that represents the result for the entire operation.\footnote{The case for chains with \code{reduce} as the last step similar and simpler.
+Thus, it is not worth analyzing.}
+
+For the \tabuffer{} case, this means copying all the elements of each partial result into a larger \ttarray{} while maintaining the original order.
+
+For the \tsabuffer{} case there are two different options:
+\begin{enumerate}
+  \item If all steps in the chain perform the \code{map} transformation, then no changes are required.
+  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textif{holes}.
+\end{enumerate}
+
+The latter is similar to the \tabuffer{} case, but instead of copying from one array to another, the source and target are the same.
 

### 2015-05-24 docs/report/array_merging.tex
Adds shared merge benchmark doc.

--- 2015-05-23 docs/report/array_merging.tex
@@
 
+\subsection{\tstarray{} merge benchmark}
+
+The benchmark presented modifies \ttarray{} merge benchmarks to support \tstarray{}s. The implementation removes \dataview{} test case as it is not supported by \tstarray{}:
+\bmcode{http://jsperf.com/sharedtypedarray-merge}{../../spikes/2.4.1/sharedMerge.js}
+
+The results for the benchmark are displayed in Figure~\ref{fig:shared-typed-array-merge}.
+\img{Comparing \tstarray{} merge approaches}{../../spikes/2.4.1/sharedMerge}{fig:shared-typed-array-merge}
+
 \subsection{Conclusion}
-The benchmark shows that the best alternative to merge small \ttarray{}s into a larger one is using \code{TypedArray::Set} method.
+The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of eighter \ttarray or \tstarray.
 

### 2015-05-24 spikes/2.4.1/sharedMerge.js

@@ -91,13 +91,6 @@
     }
   }
 
-  function withDataViewSet(parts) {
-    var result = caseDataViewSet(parts);
-    if (result.length !== (arraySize * arraysCount)) {
-      console.log('error');
-    }
-  }
-
   function withArrayLikeFunction(parts) {
     var result = caseArrayLikeFunction(parts);
     if (result.length !== (arraySize * arraysCount)) {

### 2015-05-24 docs/report/code_sharing.tex
updates report benchmarks links.

@@
 We created a benchmark to understand the difference between each approach:
-\bmcode{http://jsperf.com/pjs-serialization-comparison/2}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI.js}
+\bmcode{http://jsperf.com/pjs-serialization-comparison/4}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI.js}
 
@@
 The benchmark with the short function was the following one:
-\bmcode{http://jsperf.com/pjs-serialization/2}{../../spikes/2.2.1/test-howToSendCode/shortSerialization.js}
+\bmcode{http://jsperf.com/pjs-serialization/5}{../../spikes/2.2.1/test-howToSendCode/shortSerialization.js}
 
@@
 The benchmark with the long function was the following one:
-\bmcode{http://jsperf.com/pjs-serialization-long}{../../spikes/2.2.1/test-howToSendCode/longSerialization.js}
+\bmcode{http://jsperf.com/pjs-serialization-long/3}{../../spikes/2.2.1/test-howToSendCode/longSerialization.js}
 
@@
 The benchmark's code is the following one:
-\bmcode{http://jsperf.com/pjs-encoding}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding.js}
+\bmcode{http://jsperf.com/pjs-encoding/3}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding.js}
 

### 2015-05-24 docs/report/array_partitioning.tex
Benchmark link fixes

@@
 A benchmark was created based on the aforementioned alternatives:
-\bmcode{http://jsperf.com/arraybuffer-copy}{../../spikes/2.3.1/copyComparison.js}
+\bmcode{http://jsperf.com/arraybuffer-copy/4}{../../spikes/2.3.1/copyComparison.js}
 
@@
 Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array into parts:
-\bmcode{http://jsperf.com/arraybuffer-copy}{../../spikes/2.3.1/splitComparison.js}
+\bmcode{http://jsperf.com/arraybuffer-split/3}{../../spikes/2.3.1/splitComparison.js}
 

### 2015-05-24 docs/report/function_caching.tex
Benchmark link fixes

@@
-\section{Function caching}
+\section{Function caching}\label{sec:func-caching}
 
@@
 
-To verify its performance we created a benchmark that runs with and without the function cache:
-\bmcode{http://jsperf.com/p-j-s-with-vs-without-function-cache}{../../spikes/3.1.3/functionCache/functionCache.js}
+To verify its performance we created a benchmark that runs both with and without the function cache:
+\bmcode{http://jsperf.com/p-j-s-with-vs-without-function-cache/3}{../../spikes/3.1.3/functionCache/functionCache.js}
 
@@
 \subsection{Conclusion}
-As the benchmark shows versions 40 and 41 of chrome have almost no difference between the implementations with and without the cache.
-To determine whether introducing the cache was worth it or not we also used Chrome Canary version 42.
-The latter shows a big increase in performance (near 20\%), which is why we decided to leave the \tfunction{} cache as part of our implementation.
+As the benchmark shows versions 40 and 41 of Google Chrome have almost no difference between the implementations with and without the cache.
+To determine whether introducing the cache was worth it or not we considered the result from Firefox Nightly, which shows a big increase in performance (near 25\%), which is why we decided to make the \tfunction{} cache a part of our implementation.
 

### 2015-05-24 docs/report/array_merging.tex
Benchmark link fixes
minor fix

@@
   \item If all steps in the chain perform the \code{map} transformation, then no changes are required.
-  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textif{holes}.
+  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textit{holes}.
 \end{enumerate}
@@
 The benchmark presented only has one alternative that uses a method with a native implementation:
-\bmcode{http://jsperf.com/typedarray-merge}{../../spikes/2.4.1/merge.js}
+\bmcode{http://jsperf.com/typedarray-merge/2}{../../spikes/2.4.1/merge.js}
 

### 2015-05-24 docs/report/map_serial_vs_parallel.tex
Benchmark link fixes

@@
 On the contrary, the serial implementation operates on elements during that time.
+To observe the effects of this, we also perform the same benchmark using \tstarray{}s.
 
@@
 
-\subsection{Benchmark code}
+\subsection{TypedArray Benchmark}
 The benchmark code is introduced in the following code listing:
-\bmcode{http://jsperf.com/pjs-map-vs-serial/3}{../../spikes/3.1.3/comparison/jspBenchmark.js}
+\bmcode{http://jsperf.com/pjs-map-vs-serial/8}{../../spikes/3.1.3/comparison/jspBenchmark.js}
 
@@
 The results for the benchmark are displayed in Figure~\ref{fig:map-par-vs-seq}.
-\img{Relative performance of parallel implementation with regards to sequential one}{../../spikes/3.1.3/comparison/chart}{fig:map-par-vs-seq}
+\img{Relative performance of parallel implementation with regards to sequential one using ArrayBuffers}{../../spikes/3.1.3/comparison/chart}{fig:map-par-vs-seq}
+
+\subsection{SharedTypedArray Benchmark}
+The benchmark code is introduced in the following code listing:
+\bmcode{http://jsperf.com/pjs-map-vs-serial/9}{../../spikes/3.1.3/comparison/jspBenchmarkShared.js}
+
+\note{Unlike other benchmarks in which we were comparing different ways to achieve the same result, this benchmark compares a single way with a changing amount of elements.
+For that reason, the original charts are not displayed.}
+
+The results for the benchmark are displayed in Figure~\ref{fig:map-par-vs-seq-shared}.
+\img{Relative performance of parallel implementation with regards to sequential one using SharedArrayBuffers}{../../spikes/3.1.3/comparison/chartShared}{fig:map-par-vs-seq-shared}
 
 \subsection{Conclusion}
-The parallel improvement might be more noticeable in lower end machines since the linear implementation would be able to operate on less elements ``while the parallel is doing the split'', since the first is a computationally expensive operationa and the other one isn't.
-This is consistent with our findings based on the machine specifications in \ref{sub-sec:serialization-and-transference}.
+A couple of interesting findings result from the experiments.
+An expected one is that the amount of operations per second that can be performed with a serial implementation is almost inversely proportional to the amount of items to transform with a ratio of 1.
+On the other hand, this is not the case for the parallel approach.
+
+An acceptable speedup of approximately \textbf{2.4x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
+Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \tsabuffer{}s as they are likely not fully optimized, being that they are only an experimental feature in Firefox Nightly.
+
+The speedup obtained without shared memory is relative small (approximately \textbf{1.2x}) when compared to that obtained with shared memory.
+That can be explained by the overhead of copying arrays elements, to work around the lack of shared memory, when communicating with the Web Workers.
 

### 2015-05-24 docs/report/code_sharing.tex
Adds report benchmark for shared function transfer vs serialization.

@@
 
+\subsubsection{\tstring{} serialization and shared tranfer benchmark}
+
+We created a benchmark that compares these two scenarios:
+
+\begin{enumerate}
+\item Sending the \tstring{} representation of a function to four Web Workers.
+\item Encoding a function into a \tsabuffer{} and sending it to four Web Workers.
+\end{enumerate}
+
+The code for the benchmark is presented in the following code listing:
+\bmcode{http://jsperf.com/shared-function-transfer}{../../spikes/shared/2.2.1/jspBenchmark.longSendCode.js}
+
+The results for it are displayed in Figure~\ref{fig:serialization-vs-shared-transfer}.
+\img{Comparing string serialization vs shared buffer transfer approaches}{../../spikes/shared/2.2.1/jspBenchmarklongSendCode.png}{fig:serialization-vs-shared-transfer}
+
 \subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}

### 2015-05-24 spikes/3.1.3/comparison/jspBenchmarkShared.js
fixes shared map benchmark

@@ -74,13 +74,6 @@
     });
   };
 </script>
-<script>
-  Benchmark.prototype.setup = function() {
-    __finish = function () {
-      deferred.resolve();
-    };
-  };
-</script>
 
 // Javascript setup
 __finish = function () {

### 2015-05-24 docs/report/intro.tex
fixes title and intro tex sections.

@@
 \section{Introduction}
-In 1965 Gordon Moore proposed the popularly known ``Moore's law" \cite{moore65}, which predicts that the number of transitors in integrated circuits will double every two years.
+In 1965 Gordon Moore proposed the popularly known ``Moore's law" \cite{moore65}, which predicts that the number of transistors in integrated circuits will double every two years.
 As a consequence, programs that use a single processor/core will automatically become faster without the need to modify them with that purpose in mind.
@@
 This has been possible thanks to components like asm.js \cite{asm-web} and WebGL \cite{webgl-web} together with the constant evolution of browsers and JavaScript runtime engines.
-In this context, one of the plans of the commitee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications.
+In this context, one of the plans of the committee that develops the language is to provide an API to simplify the processing of data in parallel for version 7 (ES7) of the language, with the goal of keeping up with native applications.
 Initiatives such as ParallelJS \cite{par-js} and River trail \cite{rivertrail} or the possibility of taking advantage of SIMD instructions \cite{js-simd} (single instruction multiple data) are some of the options to implement this.

### 2015-05-24 docs/paper/paper.tex

@@
     \hline
-    100 & \textbf{142223} & 328 \\
-    1,000 & \textbf{16503.8} &  359 \\
-    10,000 & \textbf{1683.8} & 321.8 \\
-    100,000 & 169.6 & \textbf{184.4} \\
-    1,000,000 & 16.8 & \textbf{34.1} \\
+    100 & \textbf{130572} & 385 \\
+    1,000 & \textbf{14697} &  421 \\
+    10,000 & \textbf{1469} & 357 \\
+    100,000 & 151 & \textbf{188} \\
+    1,000,000 & 14 & \textbf{32} \\
     \hline
@@
 
-An acceptable speedup of approximately \textbf{2x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
-Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \verb+SharedArrayBuffer+s as they are likely not optimized, being that they are only an experimental feature in Firefox Nightly.
+An acceptable speedup of approximately \textbf{2.3x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
+Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \verb+SharedArrayBuffer+s as they are likely not fully optimized, being that they are only an experimental feature in Firefox Nightly.
 
@@
 \subsubsection*{Acknowledgments.}
-Thanks to Rosa Wachenhauzer, our tutor, for reviewing this paper and also for all her support and guidance throughout this entire project.
+Thanks to Rosa Wachenchauzer, our tutor, for reviewing this paper and also for all her support and guidance throughout this entire project.
 
@@
 
-\bibitem{benchmark} Sepia tone transformation benchmark comparison between p-j-s and plain JS, \url{http://jsperf.com/pjs-shared-vs-normal-vs-seq-100/4}
+\bibitem{benchmark} Sepia tone transformation benchmark comparison between p-j-s and plain JS, \url{http://jsperf.com/pjs-map-vs-serial/9}
 

### 2015-05-24 docs/report/map_serial_vs_parallel.tex
Fixes for map serial vs parallel cmp

@@
 
-In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform.
-That is because in the parallel case, the source \ttarray{} must be copied to sub-arrays in order to be transferred.
+In this regard, the more complex the computation to be performed on each array element, the better the parallel implementation will perform.
+That is because, in the parallel case, the source \ttarray{} must be copied to sub-arrays in order to be transferred.
 On the contrary, the serial implementation operates on elements during that time.
-To observe the effects of this, we also perform the same benchmark using \tstarray{}s.
+To observe the effects of this difference, we also perform the same benchmark using \tstarray{}s.
 
@@
 
-\note{Unlike other benchmarks in which we were comparing different ways to achieve the same result, this benchmark compares a single way with a changing amount of elements.
+\note{Unlike other benchmarks in which we were comparing different ways to achieve the same result, this benchmark compares a single way with a varying amount of elements.
 For that reason, the original charts are not displayed.}
@@
 The results for the benchmark are displayed in Figure~\ref{fig:map-par-vs-seq}.
-\img{Relative performance of parallel implementation with regards to sequential one using ArrayBuffers}{../../spikes/3.1.3/comparison/chart}{fig:map-par-vs-seq}
+\img{Relative performance of parallel implementation with regards to sequential one using \tabuffer{}s}{../../spikes/3.1.3/comparison/chart}{fig:map-par-vs-seq}
 
@@
 
-\note{Unlike other benchmarks in which we were comparing different ways to achieve the same result, this benchmark compares a single way with a changing amount of elements.
-For that reason, the original charts are not displayed.}
-
 The results for the benchmark are displayed in Figure~\ref{fig:map-par-vs-seq-shared}.
-\img{Relative performance of parallel implementation with regards to sequential one using SharedArrayBuffers}{../../spikes/3.1.3/comparison/chartShared}{fig:map-par-vs-seq-shared}
+\img{Relative performance of parallel implementation with regards to sequential one using \tsabuffer{}s}{../../spikes/3.1.3/comparison/chartShared}{fig:map-par-vs-seq-shared}
 
@@
 A couple of interesting findings result from the experiments.
-An expected one is that the amount of operations per second that can be performed with a serial implementation is almost inversely proportional to the amount of items to transform with a ratio of 1.
+An expected one is that the amount of operations per second that can be performed with a serial implementation is almost inversely proportional to the amount of items to transform (with a ratio of 1).
 On the other hand, this is not the case for the parallel approach.
 
-An acceptable speedup of approximately \textbf{2.4x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive computation (generating pseudo-random numbers).
+An acceptable speedup of approximately \textbf{2.4x} can be achieved with a relatively large amount of items (\(10E6\)) and performing a computationally expensive operation (generating pseudo-random numbers).
 Larger speedups might be possible by applying optimizations to the library and there might also be some potential improvements for operations with \tsabuffer{}s as they are likely not fully optimized, being that they are only an experimental feature in Firefox Nightly.
@@
 The speedup obtained without shared memory is relative small (approximately \textbf{1.2x}) when compared to that obtained with shared memory.
-That can be explained by the overhead of copying arrays elements, to work around the lack of shared memory, when communicating with the Web Workers.
+That can be explained by the overhead of copying array elements, to work around the lack of shared memory, when communicating with the Web Workers.
 

### 2015-05-24 docs/report/background.tex

@@
 All modern versions of major browsers (Internet Explorer, Mozilla Firefox, Google Chrome, Safari and Opera) allow the execution of JavaScript code in a single threaded event loop runtime \cite{event-loop}.
-User actions add elements to the event loop queue and these are processesed sequentially.
+User actions add elements to the event loop queue and these are processed sequentially.
 All process input/output is asynchronous to prevent the UI thread from blocking, thus keeping it responsive.
@@
 Web Workers allow the creation of ``thread like'' constructs in a browser environment, but they don't allow shared memory and instead communicate via message passing.
-The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
+The message passing overhead is very big for common objects and Web Workers ``have a high start-up performance cost'' \cite{w3c-ww-startup}, so they were commonly considered for long running tasks and operating on generally the same set of data during a single execution, making them unfit for a thread pool  model \cite{thread-pool}.\\
 However, a spec is being developed for version 7 of the EcmaScript standard that changes this, by introducing shared memory through \tsabuffer{}s \cite{sab}.

### 2015-05-24 docs/report/function_caching.tex

@@
 
-Before comparing the sequential implementation of \code{map} against the parallel one we believed that the cost of dynamically instantiating a \tfunction{} from a \tstring{} in the Web Workers could be avoided by introducing a cache for cases where the worker has already seen the code for a particular function.
+Before comparing the sequential implementation of \code{map} against the parallel one, we believed that the cost of dynamically instantiating a \tfunction{} from a \tstring{} in the Web Workers could be avoided by introducing a cache for cases where the Web Worker has already seen the code for a particular function.
 
-We decided to introduce a simple key/value map acting as a cache where the keys are the \tstring{}s with the function's code and the values are the \tfunction{} objects.
+We decided to introduce a simple key/value map acting as a cache where the keys are \tstring{}s with the function's code and the values are the \tfunction{} objects.
 
@@
 \subsection{Conclusion}
-As the benchmark shows versions 40 and 41 of Google Chrome have almost no difference between the implementations with and without the cache.
-To determine whether introducing the cache was worth it or not we considered the result from Firefox Nightly, which shows a big increase in performance (near 25\%), which is why we decided to make the \tfunction{} cache a part of our implementation.
+As the benchmark shows, there is almost no difference in versions 40 and 41 of Google Chrome between the implementations with and without the cache.
+To determine whether introducing the cache was worth it or not we considered the result from Firefox Nightly, which shows a big increase in performance (near 25\%).
+That is why we decided to make the \tfunction{} cache a part of our implementation.
 

### 2015-05-24 docs/report/micro_benchmarks.tex

@@
 \begin{description}
-\item[First] Be able to evalute the time it takes to perform the operation multiple times (a single execution could either be too small or an outlier and the measurement could have too much variance/error).
+\item[First] Be able to evaluate the time it takes to perform the operation multiple times (a single execution could either be too small or an outlier and the measurement could have too much variance/error).
 \item[Second] Allow JIT optimizations to kick in.
@@
 function between(a, b, c){
-  return a <= b <= c;
+  return a <= b && b <= c;
 }
@@
   \item SSD 128 GB
-  \item Chrome 43.0.2357
+  \item Chrome 40.0.2214
+  \item Firefox Nightly 41.0a1
 \end{itemize}

### 2015-05-24 spikes/6.2/inlinedVsGlobalInlined/jspBenchmark.js
Fixes benchmark image name.

@@ -46,12 +46,12 @@ var __finish = function (result) {
   deferred.resolve();
 };
 
-// Test case 1 - map with inlining
+// Test case 1 - map with function instance
 wrappedXs.map(inlinedMapper).seq(function (result) {
   __finish();
 });
 
-// Test case 2 - map with global inlining
+// Test case 2 - map with function key
 wrappedXs.map('x').seq(function (result) {
   __finish();
 });

### 2015-05-24 docs/report/initialization.tex
Added library overview section
updates report benchmarks links.
fixes initialization tex section.

@@
-\section{Initialization}
+\section{Initialization}\label{sec:initialization}
 
-As explained in the Web Workers standard \cite{w3c-ww} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
+As explained in the Web Workers standard \cite{w3c-ww-startup} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.
 
@@
 The following is the code for the benchmark:
-\bmcode{http://jsperf.com/worker-cold-start/5}{../../spikes/2.1.1/newVsPreExistent.js}
+\bmcode{http://jsperf.com/worker-cold-start/6}{../../spikes/2.1.1/newVsPreExistent.js}
 
@@
 The second factor makes it important for users to be able to control/limit the amount of web workers that the library instantiates.
-For that reason the library allows users to specify a maximum amount of workers to be instantiated, and also provide through its API a way to terminate all workers in existance.
+For that reason the library allows users to specify a maximum amount of workers to be instantiated, and also provide through its API a way to terminate all workers in existence.
 

### 2015-05-24 docs/report/function_transfer_time.tex

@@
 Each Web Worker has a cache to avoid using different instances of the same function (as explained in Section~\ref{sec:func-caching}).
-Nevertheless, if the same function is going to be used multiple times in a program as the callback for steps, it might make sense to send them to the Web Workers once and then simply make use of them.
+Nevertheless, if the same function is going to be used multiple times in a program as the callback for steps, it might make sense to send it to the Web Workers once and then simply make a reference to it.
 
-This can be achieved by taking advantage of the global context object (detailed in Section~\ref{sec:context}).
+This can be achieved by taking advantage of the global context object (introduced in Section~\ref{sec:context}).
 
-The first thing that needs to be done is to call \code{pjs.updateContext} by providing a \tfunction{} as a property of the context parameter. An example is shown in the following code snippet:
+The first thing that needs to be done is to call \code{pjs.updateContext} by providing a \tfunction{} as a property of the context object. An example is shown in the following code snippet:
 \begin{lstlisting}[caption=Sending function in global context]
@@
 An example is shown in the following code snippet:
-\begin{lstlisting}[caption=Passing a function name from the global context instead of a funciton as a callback]
+\begin{lstlisting}[caption=Passing a function name from the global context instead of a \tfunction{} as a callback]
 var promise = pjs.updateContext({
@@
 The results for the benchmark are displayed in Figure~\ref{fig:func-key-vs-instance}.
-\img{Comparing sending function key in global context against function instance}{../../spikes/6.2/inlinedVsGlobalInlined/result}{fig:func-key-vs-instance}
+\img{Comparing sending function key in global context against \tfunction{} instance}{../../spikes/6.2/inlinedVsGlobalInlined/result}{fig:func-key-vs-instance}
 
-\note{The improvement shown in Figure~\ref{fig:func-key-vs-instance} will vary if the difference in character length between the key and the function's code does.}
+\subsection{Conclusion}
+While the performance improvement obtained is not remarkable, it is still noticeable and does not require a lot of effort from a developer's standpoint.
+It is worth pointing out that the improvement shown in Figure~\ref{fig:func-key-vs-instance} will vary if the difference in character length between the key and the function's code does.
 

### 2015-05-24 docs/report/overview.tex
Fix array partition section and bib links

@@
 JavaScript developers are familiar with the concept of map, filter and reduce skeletons as language constructs.
-In JavaScript, \code{Array}s have \code{map}, \code{filter}, and \code{reduce} functions and version 6 of the EcmaScript standard adds those same methods to \ttarray{}s \cite{es6-ta}.
+In JavaScript, \code{Array}s have \code{map}, \code{filter}, and \code{reduce} functions and version 6 of the EcmaScript standard adds those same methods to \ttarray{}s \cite{es6}.
 For example, to multiply all the elements of an \code{Array} by two:

### 2015-05-24 docs/report/context.tex
updates report benchmarks links.
Changes for context

@@
 The reason why there are two kinds of context is because there will usually be data that does not need to be passed to the Web Workers with each chain, so having a global context allows this.
-Additionally, the local context allows users to avoid polluting the global one with keys that are not necessary for all operations.
+Additionally, the local context provides a way for developers to avoid polluting the global one with keys that are not necessary for all operations.
 
@@
 \subsection{Implementation}
-Internally the context is sent as a message to all workers to make its properties available. Two possible implementatios were possible:
+Internally the context is sent as a message to all Web Workers to make its properties available. Two possible implementatios were possible:
 \begin{itemize}
-  \item Send the object directly and rely on structured cloning
-  \item Serialize the object using \code{JSON.stringify} and deserialize it in the workers
+  \item Sending the object directly and rely on structured cloning
+  \item Serializing the object using \code{JSON.stringify} and deserialize it in the workers
 \end{itemize}
@@
 
-We created to separate benchmarks, one that sends small objects to a Web Worker and another one that sends large objects to a Web Worker. Both benchmarks compare the approaches against each other.
+We created two separate benchmarks, one that sends small objects to a Web Worker and another one that sends large objects to a Web Worker. Both benchmarks compare the approaches against each other.
 
@@
 \subsection{Conclusion}
-Based on the benchmark results we decided to use \code{JSON.stringify} to serialize \code{context} objects.
+Based on the benchmark results, we decided to use \code{JSON.stringify} to serialize \code{context} objects.
 This involves a tradeoff as, unlike structured cloning, it does not support cyclic references.

### 2015-05-25 docs/report/array_merging.tex

@@
 
-Once all the workers have executed the function provided to them on each element of their portion of the original array, they are to notify the main thread and provide the resulting \tabuffer{} or \tsabuffer{}.
+Once all the workers have executed the \tfunction{} provided to them on each element of their portion of the original array, they are to notify the UI thread and provide the resulting \tabuffer{} or \tsabuffer{}.
 
@@
 
-For the \tsabuffer{} case there are two different options:
+For the \tsabuffer{} case there are two possible situations:
 \begin{enumerate}
@@
 
-The benchmark presented modifies \ttarray{} merge benchmarks to support \tstarray{}s. The implementation removes \dataview{} test case as it is not supported by \tstarray{}:
+The benchmark modifies the \ttarray{} merge benchmark to support instances of the \tstarray{} type.
+The implementation does not include the \dataview{} alternative as it is not supported by \tstarray{}s:
 \bmcode{http://jsperf.com/sharedtypedarray-merge}{../../spikes/2.4.1/sharedMerge.js}
@@
 \subsection{Conclusion}
-The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of eighter \ttarray or \tstarray.
+The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of either \ttarray{} or \tstarray{}.
 

### 2015-05-25 docs/report/array_partitioning.tex

@@
 
-As explained in section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance.
-Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned.}
+As explained in Section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from Web Workers requires the use of \code{Transferable}s to achieve acceptable performance.
+Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it spends as little time as possible partitioning the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned.}
 
 If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s.
-For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but \ttarray{}s do not support shared memory and element ownership is tranferred to the Web Workers, the elements must be copied to the sub-arrays.
+For ideal performance it would be nice to be able to pass a \textit{pointer} into the array and an amount of elements to process, but because \ttarray{}s do not support shared memory, and element ownership is tranferred to the Web Workers, the elements must be copied to the sub-arrays.
 
@@
 \begin{itemize}
-  \item Implementing a non-native function that copies the array elements one by one
+  \item Implementing a non-native function that copies the array elements one by one.
   \item Using a \code{Blob} to store the contents of a \ttarray{} and reading the \code{Blob}'s contents into a new \ttarray{}.
@@
 
-Based on those results we can discard the \code{Blob} approach for splitting \ttarray{}s.
+Based on those results we can discard the \code{Blob} approach for partitioning \ttarray{}s.
 
@@
 
-\subsubsection{\ttarray{} split benchmark}
-Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array into parts:
+\subsubsection{\ttarray{} partition benchmark}
+Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array:
 \bmcode{http://jsperf.com/arraybuffer-split/3}{../../spikes/2.3.1/splitComparison.js}
@@
 The results for it are displayed in Figure~\ref{fig:typed-array-split}.
-\img{Comparing \ttarray{} split approaches}{../../spikes/2.3.1/splitComparison}{fig:typed-array-split}
+\img{Comparing \ttarray{} partition approaches}{../../spikes/2.3.1/splitComparison}{fig:typed-array-split}
 
@@
 The benchmark shows that there are three very similar alternatives to partition a \ttarray{}.
-We are going to use \textit{Buffer subarray} as it yields better performance in a newer version of Chrome.
+We decided to use \textit{Buffer subarray}.
 The implementation of the slice function could look something like this:

### 2015-05-25 docs/report/code_sharing.tex

@@
 Function's code not only needs to be serialized but also transferred.
-For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and tranfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.
+For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and transfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.
 
@@
 
-\subsubsection{\tstring{} serialization and tranfer benchmarks}
+\subsubsection{\tstring{} serialization and transfer benchmarks}
 We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time.
@@
 \begin{enumerate}
-  \item Encoding the \tstring{} into an \tabuffer{}, sending it as a \code{Tranferable} and decoding it on the Web Worker.
+  \item Encoding the \tstring{} into an \tabuffer{}, sending it as a \code{Transferable} and decoding it on the Web Worker.
   \item Sending the string directly
@@
 
-\subsubsection{\tstring{} serialization and shared tranfer benchmark}
+\subsubsection{\tstring{} serialization and shared transfer benchmark}
 
@@
 
-\subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}
-We created a benchmark that tranfers a \ttarray{} and a \tstring{} that are properties of the same object back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.
+\subsubsection{\tstring{} and \ttarray{} serialization and transfer benchmarks}
+We created a benchmark that transfers a \ttarray{} and a \tstring{} that are properties of the same object back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just transferring a \tstring{}.
 
@@
 \subsubsection{Conclusion}
-As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by sending it directly.
+As the benchmarks show, the fastest way to transfer the \tstring{} between the UI thread and workers is by sending it directly.
 

### 2015-05-25 docs/report/inlining.tex
Benchmark link fixes

@@
 
-Which was implemented as shown in the following listing for the parallel case:
+It was implemented as shown in the following listing for the parallel case:
 \begin{lstlisting}[caption=Sepia tone initial parallel implementation]
@@
 \subsection{The problem}
-Taking the experience from previous experiments into account we considered that the performance of the two implementations with a source \code{Uint32Array} of about \(8E7\) (eight million) elements should be similar, even favoring the parallel one.
+Taking the experience from previous experiments into account we considered that the performance of the two implementations with a source \code{Uint32Array} of about 8,000,000 (eight million) elements should be similar, even favoring the parallel one.
 The initial measurements showed that the linear implementation fared approximately four times better than the parallel one.
@@
 Specifically we believed that the internal functions \code{noise}, \code{clamp} and \code{colorDistance} were not being inlined.
-Inlining could be the cause of both benefits\cite{inline-pro} and drawbacks\cite{inline-cons} but the belief was that performance was being affected because of lack of it
+Inlining could be the cause of both benefits\cite{inline-pro} and drawbacks\cite{inline-cons} but the belief was that performance was being affected because of lack of it.
 
@@
 \subsubsection{Parallel implementation generated code}
-Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the serial implementation. As it can be seen from them, only \code{Math.random} is being inlined.
+Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the parallel implementation. As it can be seen from them, only \code{Math.random} is being inlined.
 \img{\code{clamp} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-clamp}{fig:parallel-clamp}
@@
 We created a benchmark to understand the difference between the initial version (the one in which functions were not inlined) and the final one (with inlining in charge of the developer):
-\bmcode{http://jsperf.com/pjs-map-inlining}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}
+\bmcode{http://jsperf.com/pjs-map-inlining/6}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}
 
 The results for it are displayed in Figure~\ref{fig:pjs-map-inlining}.
-\img{Comparing string encoding/decoding approaches}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining}{fig:pjs-map-inlining}
+\img{Comparing a version of \code{map} without inlining against a developer inlined one}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining}{fig:pjs-map-inlining}
 

### 2015-05-25 docs/report/modules.tex

@@
 
-The two most known module formats/API in JavaScript are known as:
+The two most known module formats/API in JavaScript are:
 \begin{description}
@@
 
-Instead of making the choice for users, our library will support both systems in a way that makes it easier for users to consume either of them.
+Instead of making the choice for users, our library will support both systems in a way that makes it easy for users to consume either of them.
 
@@
 \subsection{AMD support}
-This example shows how to use the library with require.js. It assumes a folder structure like the one shown below:
+This example shows how to use the library with require.js. It assumes a folder structure like the one in the following listing:
 \begin{lstlisting}[language=bash, caption=AMD folder structure]
@@
 \subsection{Conclusion}
-As the previous examples show, using our library with either module system is really simple and is done in a way that should feel familiar for users of those module formats.
+As the previous examples show, using our library with either module system is really simple and is done in a way that should feel familiar for users of those systems.
 

### 2015-05-25 docs/report/serialization_transference.tex

@@
 The following is the code for the benchmark:
-\bmcode{http://jsperf.com/transferrable-vs-cloning}{../../spikes/1.2/transferrableVsCloning.js}
+\bmcode{http://jsperf.com/transferrable-vs-cloning/3}{../../spikes/1.2/transferrableVsCloning.js}
 
@@
 The following is the code for the benchmark:
-\bmcode{http://jsperf.com/longer-transferrable-vs-cloning}{../../spikes/1.2/longerTransferrableVsCloning.js}
+\bmcode{http://jsperf.com/longer-transferrable-vs-cloning/3}{../../spikes/1.2/longerTransferrableVsCloning.js}
 

### 2015-05-25 docs/report/code_sharing.tex
fixes code sharing subtitles

@@
 
-\subsubsection{\tstring{} serialization benchmarks}
+\subsubsection{\tstring{} encoding benchmarks}
 One possible way of passing serialized \tfunction{}s to workers would be to encode them as binary and transfer them as an \tabuffer{}. To encode the strings there are two possible approaches:
@@
 
-\subsubsection{\tstring{} serialization and transfer benchmarks}
+\subsubsection{\tstring{} encoding and transfer benchmarks}
 We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time.

### 2015-05-25 docs/report/context.tex
fixes context benchmark image.

@@
 The code for the benchmark is in the following listing:
-\bmcode{http://jsperf.com/additional-sendingdata-comparison/2}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark.js}
+\bmcode{http://jsperf.com/additional-sendingdata-comparison/3}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark.js}
 
 The results for the benchmark are displayed in Figure~\ref{fig:context-send}.
-\img{Comparing small context message passing}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark-all-browsers}{fig:context-send}
+\img{Comparing small context message passing}{../../spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark}{fig:context-send}
 

### 2015-05-25 spikes/6.1/sendingAdditionalsSeqVsPar/jspBenchmark.js

@@ -1,4 +1,4 @@
-// http://jsperf.com/additional-sendingdata-comparison/2
+// http://jsperf.com/additional-sendingdata-comparison/3
 
 // HTML setup
 <script>

### 2015-05-25 spikes/3.1.3/pjsMapInlining/pjsMapInlining.js
updates inlining and cache benchmarks.

@@ -1,4 +1,4 @@
-// http://jsperf.com/pjs-map-inlining/6
+// http://jsperf.com/pjs-map-inlining/7
 
 // HTML setup
 <script src="http://rawgit.com/pjsteam/pjs/v1.0.0-beta/dist/p-j-s.min.js"></script>

### 2015-05-25 docs/report/inlining.tex
fixes inlining code sample from inlining report section.

@@
   return (pixel & 0xFF000000) + (new_b << 16) + (new_g << 8) + (new_r & 0xFF);
-}, function(result){
+}).seq(function(result){
   // work on result
@@
 We created a benchmark to understand the difference between the initial version (the one in which functions were not inlined) and the final one (with inlining in charge of the developer):
-\bmcode{http://jsperf.com/pjs-map-inlining/6}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}
+\bmcode{http://jsperf.com/pjs-map-inlining/7}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}
 
